---
title: "MDD Integrated Module GeneShot Annotation"
author: "Daniel Derrick"
date: "4/24/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    df_print: paged
  html_notebook:
    code_folding: hide
    toc: true
    toc_float: true
---


This script creates volcano plots from MDD integrated module genes,
annotated with enrichment scores from geneShot literature searches
for terms related to the modules.

```{r setupAndPaths, include=FALSE}
library(DESeq2)
library(biomaRt)
library(tidyverse)
library(plotly)
library(cowplot)
theme_set(theme_cowplot())

if(!grepl("R$", getwd())) {setwd("R")}

gSDir     <- "misc/geneShot"
moduleDir <- "misc/Ligand_tables"
resFile   <- "RNAseq/MCF10A_RNAseq_DESeq2Results48.Rdata"

outDir <- "../plots/allModule_geneShot"
outPathWidget <- "/Users/derrickd/workspace/MDD/plots/allModule_geneShot/MDD_RNAseq_geneshot_scatterplot.html"
```

### Geneshot searches 

Reading in geneshot files for cell cycle, EMT, migration, and interferon gamma-related
literature seraches.

|GeneShot Search|Search terms|Module|Ligand|
|---|---|---|---|
|cell cycle|cell cycle, proliferation| Module 3 | PBS |
|EMT|epithelial-to-mesenchymal transition, EMT| Modules 7,8 | TGFB |
|migration|motility, migration | Module 10 | OSM |
|IFNG|Interferon gamma, immune, immunity| Module 12 | IFNG |


```{r}
gsFiles <- grep("README", 
                list.files(gSDir, full.names = TRUE),
                invert = TRUE, value = TRUE)

gsList <- mapply(function(x, x_nm) {
  x <- read.table(x, 
                  sep = "\t", stringsAsFactor = FALSE, header = TRUE)
  colnames(x)[-2] <- sprintf("%s_%s", colnames(x)[-2], x_nm)
  x
  }, 
  x = gsFiles, 
  x_nm = str_remove(str_remove(str_extract(gsFiles, "_[:alnum:]+.tsv"), ".tsv"), "_"),
  SIMPLIFY = FALSE)

names(gsList) <-
  str_remove(str_remove(str_extract(gsFiles, "_[:alnum:]+.tsv"), ".tsv"), "_")

gsLong <-
  Reduce(full_join, gsList) %>% 
  dplyr::rename(hgnc_symbol = Gene)
```

### Reading in module features, selecting RNAseq features

```{r}
mFiles        <- list.files(moduleDir, full.names = TRUE, pattern = "*.csv")
names(mFiles) <- str_replace(str_extract(mFiles, "cluster_[0-9]+"), 
                              "cluster", "module")

fT <- lapply(mFiles, read.csv, header = FALSE, stringsAsFactors = FALSE)

fT <-
  lapply(fT, function(x) {
    colnames(x) <- c("feature", sprintf("V%s", seq(2, 10, 1)), "type", "symbol")
    x 
    })

moduleRNAseqFeatures <- lapply(fT, function(x) {
  x <- x %>%
    filter(type == "RNA") %>%
    distinct(symbol) %>%
    pull(symbol)
  })
moduleRNAseqFeatures <- moduleRNAseqFeatures[c(1, 5:12, 2:4)]
```

### Reading in RNAseq differential expression results

```{r}
load(resFile)
```


### Replacing adjusted p-values of 0

Some genes in the DESeq2 results tables have adjusted p-values
of 0, which causes issues when the data are log-transformed.
Adjusted p-values of 0 will be replaced by the minimum
adjusted p-value for the evaluated condition.

```{r}
RNAseqRT <-
  mapply(function(x, x_nm) {
    if(min(x$padj, na.rm = TRUE) == 0) {
      xmin <- min(sort(unique(x$padj))[-1], na.rm = TRUE)
      x$padj[x$padj == 0] <- xmin
    } else {
      x <- x
    }
    temp <- sprintf("%s_%s", colnames(x[-c(1, 7)]), x_nm)
    colnames(x)[-c(1,2,7)] <- sprintf("%s_%s", colnames(x)[-c(1,2, 7)], x_nm)
    x
  }, 
  x = RNAseq_resultsTables, x_nm = names(RNAseq_resultsTables), SIMPLIFY = FALSE)

RNAseqLong <- Reduce(left_join, RNAseqRT)
```

### Joining modules to RNAseq results


```{r}
GS_RNA <- 
  RNAseqLong %>%
  left_join(gsLong)

combined <- GS_RNA %>% 
  mutate(module1 = hgnc_symbol %in% moduleRNAseqFeatures$module_1,
         module2 = hgnc_symbol %in% moduleRNAseqFeatures$module_2,
         module3 = hgnc_symbol %in% moduleRNAseqFeatures$module_3,
         module4 = hgnc_symbol %in% moduleRNAseqFeatures$module_4,
         module5 = hgnc_symbol %in% moduleRNAseqFeatures$module_5,
         module6 = hgnc_symbol %in% moduleRNAseqFeatures$module_6,
         module7 = hgnc_symbol %in% moduleRNAseqFeatures$module_7,
         module8 = hgnc_symbol %in% moduleRNAseqFeatures$module_8,
         module9 = hgnc_symbol %in% moduleRNAseqFeatures$module_9,
         module10 = hgnc_symbol %in% moduleRNAseqFeatures$module_10,
         module11 = hgnc_symbol %in% moduleRNAseqFeatures$module_11,
         module12 = hgnc_symbol %in% moduleRNAseqFeatures$module_12)
```

## Plots

### GeneShot: "Cell Cycle"; RNAseq changes: PBS_48

The volcano plots below show RNA-seq changes in PBS_48 annotated with a 
GeneShot search for "cell cycle" and "proliferation", in modules 3, 7, 8, 10, and 12.

#### Module: Module 3 (PBS Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module3) %>% 
  ggplot(aes(x = log2FoldChange_PBS_48,
             y = -log10(padj_PBS_48),
             col = Rank_cellCycle,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("PBS_48 log2FC") +
  ylab("PBS_48 -log10(padj)")+
  ggtitle("Module 3 Genes",
          "geneshot: 'cell cycle', 'proliferation'")
```

#### Module: Module 7 (TGFB Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module7) %>% 
  ggplot(aes(x = log2FoldChange_PBS_48,
             y = -log10(padj_PBS_48),
             col = Rank_cellCycle,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("PBS_48 log2FC") +
  ylab("PBS_48 -log10(padj)")+
  ggtitle("Module 7 Genes",
          "geneshot: 'cell cycle', 'proliferation'")
```

#### Module: Module 8 (TGFB Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module8) %>% 
  ggplot(aes(x = log2FoldChange_PBS_48,
             y = -log10(padj_PBS_48),
             col = Rank_cellCycle,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("PBS_48 log2FC") +
  ylab("PBS_48 -log10(padj)")+
  ggtitle("Module 8 Genes",
          "geneshot: 'cell cycle', 'proliferation'")
```

#### Module: Module 10 (OSM Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module10) %>% 
  ggplot(aes(x = log2FoldChange_PBS_48,
             y = -log10(padj_PBS_48),
             col = Rank_cellCycle,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("PBS_48 log2FC") +
  ylab("PBS_48 -log10(padj)")+
  ggtitle("Module 10 Genes",
          "geneshot: 'cell cycle', 'proliferation'")
```

#### Module: Module 12 (IFNG Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module12) %>% 
  ggplot(aes(x = log2FoldChange_PBS_48,
             y = -log10(padj_PBS_48),
             col = Rank_cellCycle,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("PBS_48 log2FC") +
  ylab("PBS_48 -log10(padj)")+
  ggtitle("Module 12 Genes",
          "geneshot: 'cell cycle', 'proliferation'")
```

### GeneShot: "EMT" + RNAseq changes: TGFB_48

The volcano plots below show RNA-seq changes in TGFB_48 annotated with a 
GeneShot search for "epithelial-to-mesenchymal transition" and "EMT", in modules 3, 7, 8, 10, and 12.

#### Module: Module 3 (PBS Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module3) %>% 
  ggplot(aes(x = log2FoldChange_TGFB_48,
             y = -log10(padj_TGFB_48),
             col = Rank_EMT,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("TGFB_48 log2FC") +
  ylab("TGFB_48 -log10(padj)")+
  ggtitle("Module 3 Genes",
          "geneshot: 'epithelial-to-mesenchymal transition', 'EMT'")
```

#### Module: Module 7 (TGFB Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module7) %>% 
  ggplot(aes(x = log2FoldChange_TGFB_48,
             y = -log10(padj_TGFB_48),
             col = Rank_EMT,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("TGFB_48 log2FC") +
  ylab("TGFB_48 -log10(padj)")+
  ggtitle("Module 7 Genes",
          "geneshot: 'epithelial-to-mesenchymal transition', 'EMT'")
```

#### Module: Module 8 (TGFB Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module8) %>% 
  ggplot(aes(x = log2FoldChange_TGFB_48,
             y = -log10(padj_TGFB_48),
             col = Rank_EMT,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("TGFB_48 log2FC") +
  ylab("TGFB_48 -log10(padj)")+
  ggtitle("Module 8 Genes",
          "geneshot: 'epithelial-to-mesenchymal transition', 'EMT'")
```

#### Module: Module 10 (OSM Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module10) %>% 
  ggplot(aes(x = log2FoldChange_TGFB_48,
             y = -log10(padj_TGFB_48),
             col = Rank_EMT,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("TGFB_48 log2FC") +
  ylab("TGFB_48 -log10(padj)")+
  ggtitle("Module 10 Genes",
          "geneshot: 'epithelial-to-mesenchymal transition', 'EMT'")
```

#### Module: Module 12 (IFNG Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module12) %>% 
  ggplot(aes(x = log2FoldChange_TGFB_48,
             y = -log10(padj_TGFB_48),
             col = Rank_EMT,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("TGFB_48 log2FC") +
  ylab("TGFB_48 -log10(padj)")+
  ggtitle("Module 12 Genes",
          "geneshot: 'epithelial-to-mesenchymal transition', 'EMT'")
```

### GeneShot: "Migration" + RNAseq changes: OSM_48

The volcano plots below show RNA-seq changes in OSM_48 annotated with a 
GeneShot search for "migration" and "motility", in modules 3, 7, 8, 10, and 12.

#### Module: Module 3 (PBS Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module3) %>% 
  ggplot(aes(x = log2FoldChange_OSM_48,
             y = -log10(padj_OSM_48),
             col = Rank_migration,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("OSM_48 log2FC") +
  ylab("OSM_48 -log10(padj)")+
  ggtitle("Module 3 Genes",
          "geneshot: 'migration', 'motility'")
```

#### Module: Module 7 (TGFB Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module7) %>% 
  ggplot(aes(x = log2FoldChange_OSM_48,
             y = -log10(padj_OSM_48),
             col = Rank_migration,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("OSM_48 log2FC") +
  ylab("OSM_48 -log10(padj)")+
  ggtitle("Module 7 Genes",
          "geneshot: 'migration', 'motility'")
```

#### Module: Module 8 (TGFB Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module8) %>% 
  ggplot(aes(x = log2FoldChange_OSM_48,
             y = -log10(padj_OSM_48),
             col = Rank_migration,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("OSM_48 log2FC") +
  ylab("OSM_48 -log10(padj)")+
  ggtitle("Module 8 Genes",
          "geneshot: 'migration', 'motility'")
```

#### Module: Module 10 (OSM Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module10) %>% 
  ggplot(aes(x = log2FoldChange_OSM_48,
             y = -log10(padj_OSM_48),
             col = Rank_migration,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("OSM_48 log2FC") +
  ylab("OSM_48 -log10(padj)")+
  ggtitle("Module 10 Genes",
          "geneshot: 'migration', 'motility'")
```

#### Module: Module 12 (IFNG Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module12) %>% 
  ggplot(aes(x = log2FoldChange_OSM_48,
             y = -log10(padj_OSM_48),
             col = Rank_migration,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("OSM_48 log2FC") +
  ylab("OSM_48 -log10(padj)")+
  ggtitle("Module 12 Genes",
          "geneshot: 'migration', 'motility'")
```



### GeneShot: "immune" + RNAseq changes: IFNG_48

The volcano plots below show RNA-seq changes in IFNG_48 annotated with a 
GeneShot search for "immune", "immunity" and "interferon gamma", in
modules 3, 7, 8, 10, and 12.

#### Module: Module 3 (PBS Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module3) %>% 
  ggplot(aes(x = log2FoldChange_IFNG_48,
             y = -log10(padj_IFNG_48),
             col = Rank_IFNG,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("IFNG_48 log2FC") +
  ylab("IFNG_48 -log10(padj)")+
  ggtitle("Module 3 Genes",
          "geneshot: 'immune', 'immunity', 'interferon gamma'")
```

#### Module: Module 7 (TGFB Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module7) %>% 
  ggplot(aes(x = log2FoldChange_IFNG_48,
             y = -log10(padj_IFNG_48),
             col = Rank_IFNG,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("IFNG_48 log2FC") +
  ylab("IFNG_48 -log10(padj)")+
  ggtitle("Module 7 Genes",
          "geneshot: 'immune', 'immunity', 'interferon gamma'")
```

#### Module: Module 8 (TGFB Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module8) %>% 
  ggplot(aes(x = log2FoldChange_IFNG_48,
             y = -log10(padj_IFNG_48),
             col = Rank_IFNG,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("IFNG_48 log2FC") +
  ylab("IFNG_48 -log10(padj)")+
  ggtitle("Module 8 Genes",
          "geneshot: 'immune', 'immunity', 'interferon gamma'")
```

#### Module: Module 10 (OSM Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module10) %>% 
  ggplot(aes(x = log2FoldChange_IFNG_48,
             y = -log10(padj_IFNG_48),
             col = Rank_IFNG,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("IFNG_48 log2FC") +
  ylab("IFNG_48 -log10(padj)")+
  ggtitle("Module 10 Genes",
          "geneshot: 'immune', 'immunity', 'interferon gamma'")
```

#### Module: Module 12 (IFNG Module)

```{r, fig.height = 12, fig.width = 12}
combined %>% 
  filter(module12) %>% 
  ggplot(aes(x = log2FoldChange_IFNG_48,
             y = -log10(padj_IFNG_48),
             col = Rank_IFNG,
             label = hgnc_symbol)) +
  geom_text(alpha = .85) +
  scale_color_gradient(low = "pink", high = "red") +
  xlab("IFNG_48 log2FC") +
  ylab("IFNG_48 -log10(padj)")+
  ggtitle("Module 12 Genes",
          "geneshot: 'immune', 'immunity', 'interferon gamma'")
```

