---
title: "MDD IF batch analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE,  message=FALSE, warning = FALSE, fig.height=10, fig.width=10)

suppressMessages(library(tidyverse))
suppressMessages(library(ComplexHeatmap))
library(iheatmapr)
library(DT)
```


## Overview{.tabset .tabset-fade}
These figures are derived from the data downloaded from the LINCS Molecular Deep Dive files on Synapse https://www.synapse.org/#!Synapse:syn2862345/wiki/588244 and from segmenting the same images using CellProfiler.  



```{r globalVariables}

ligand_cols <- c("ctrl" = "#7A4A2A",
                 "PBS" = "#8dd3c7",
                  "HGF" = "#80b1d3",
                  "OSM" = "#fdb462",
                  "EGF" = "#fb8072",
                  "BMP2" = "#b3de69",
                  "IFNG" = "#bebada",
                  "TGFB" = "#ffd92f")

ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")

```

```{r read_data_and_metadata}

md <- read_csv("metadata/MDD_sample_annotations.csv")

IF_CP <- read_csv("IF/Data/MCF10A_IF_Ilastik_Well_File.csv") %>%
  mutate(ligand = factor(ligand, levels = c("ctrl","PBS","BMP2","IFNG","TGFB","HGF","OSM","EGF")),
         PairedWithEGF = FALSE,
         media = "")
IF_CP$PairedWithEGF[IF_CP$ligand %in% c("BMP2", "IFNG","TGFB")] <- TRUE
IF_CP$media[IF_CP$ligand %in% c("BMP2", "IFNG","TGFB")] <- "+EGF"

IF_CP$ligand_media <- paste0(IF_CP$ligand, IF_CP$media)
IF_CP$ligand_media <- factor(IF_CP$ligand_media, levels = c("ctrl","PBS","BMP2+EGF","IFNG+EGF","TGFB+EGF","HGF","OSM","EGF"))

IF_SC_Level1 <- read_csv("IF/Data/MDD_IF_Level1.csv")

IF_CP_SC <- IF_SC_Level1 %>%
  group_by(barcode, WellIndex) %>%
  summarise(Well_Cell_Count_SC = n(),
            MeanIntensity_KRT5 = mean(MeanIntensity_KRT5),
            MeanIntensity_DAPI = mean(MeanIntensity_DAPI),
            MeanIntensity_CellMask = mean(MeanIntensity_CellMask)) %>%
  ungroup() %>%
  right_join(IF_CP)

IF_C3 <- read_csv("cycIF/Data/CyclicIF_cell_counts_by_well.csv") %>%
  mutate(collection = "C3",
         condition = paste(ligand, time, sep = "_"),
         specimanName = paste(condition, collection, replicate, sep = "_"),
         Plate = as.character(Plate)) %>%
  rename(barcode = Plate,
         Well_Cell_Count = Counts) %>%
  mutate(ligand = factor(ligand, levels = c("ctrl","PBS","BMP2","IFNG","TGFB","HGF","OSM","EGF")),
         PairedWithEGF = FALSE,
         media = "")

IF_C3$PairedWithEGF[IF_C3$ligand %in% c("BMP2", "IFNG","TGFB")] <- TRUE
IF_C3$media[IF_C3$ligand %in% c("BMP2", "IFNG","TGFB")] <- "+EGF"

IF_C3$ligand_media <- paste0(IF_C3$ligand, IF_C3$media)
IF_C3$ligand_media <- factor(IF_C3$ligand_media, levels = c("ctrl","PBS","BMP2+EGF","IFNG+EGF","TGFB+EGF","HGF","OSM","EGF"))

  IF_C1C2C3 <- bind_rows(IF_CP, IF_C3) %>%
    filter(time %in% c(0,24,48))
  
```


```{r compare_CP_ScanR, fig.width=4, fig.height=4,eval = TRUE}

cor_pear <- cor(IF_CP_SC$Well_Cell_Count, IF_CP_SC$Well_Cell_Count_SC)

p <- ggplot(IF_CP_SC, aes(Well_Cell_Count_SC, Well_Cell_Count)) +
  geom_point(alpha = .8) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_continuous(breaks = c(0,2000, 4000, 6000)) +
  scale_y_continuous(breaks = c(0,2000, 4000, 6000)) +
  labs(title = "CP vs. ScanR cell counts are highly correlated",
       subtitle = sprintf("Pearson correlation: %3.2f", cor_pear),
       x = "ScanR well cell counts",
       y = "CellProfiler well cell counts") +
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 8))
p

cor_pear <- cor(IF_CP_SC$MeanIntensity_KRT5, IF_CP_SC$Intensity_MeanIntensity_KRT5)

p <- ggplot(IF_CP_SC, aes(MeanIntensity_KRT5, Intensity_MeanIntensity_KRT5)) +
  geom_point(alpha = .8) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_continuous(limits = c(400, 1700), breaks = seq(400, 1600, 200))+
  scale_y_continuous(limits = c(400, 1700), breaks = seq(400, 1600, 200)) +
  labs(title = "CP vs. ScanR KRT5 intensities are highly correlated",
       subtitle = sprintf("Pearson correlation: %3.2f", cor_pear),
       x = "ScanR KRT5 mean intensity",
       y = "CellProfiler KRT5 mean intensity") +
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 8))
p

```

$~$    
     
The following figures are based on the CellProfiler data for IF collections 1 and 2 and HMS cycIF collection 3.  

The cell counts in collections 1 and 2 are the sum of 12 images per well. Collection 3 cell counts are from 9 images per well.  
$~$ 
  

```{r count_and_intensity_boxplots, fig.width= 8, fig.height=4}

p <- ggplot(IF_C1C2C3, aes(x=ligand_media, y=Well_Cell_Count, fill=factor(PairedWithEGF)))+
  geom_boxplot(position = position_dodge(width=0.9), outlier.colour = NA)+
  labs(x="Ligand",
       y="Well Cell Count",
       fill="Media with EGF",
       title="Cell Count by ligand and time point")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)),
        plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(1)),
        legend.title=element_text(size = rel(1)))+
  facet_wrap(~time)
p <- p +  geom_point(aes(colour=collection), position=position_jitterdodge(dodge.width=0.9), size = .7,  alpha=.8)+
  scale_color_manual(values=c("C1"="blue", "C2"="purple", "C3"="red"))
print(p)

```


```{r count_and_intensity_batches_boxplots, fig.width= 8, fig.height=8}

p <- ggplot(IF_C1C2C3, aes(x=ligand_media, y=Well_Cell_Count, colour=collection, fill=factor(PairedWithEGF)))+
  geom_boxplot(position = position_dodge(width=0.9), outlier.colour = NA)+
  labs(x="Ligand",
       y="Well Cell Count",
       fill="Media with EGF",
       title="Cell Count by ligand and time point")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)),
        plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.7)),
        legend.title=element_text(size = rel(1)))+
  facet_wrap(~time, ncol = 1)
p <- p +  geom_point(aes(colour=collection), position=position_jitterdodge(dodge.width=0.9), size = .7, alpha=.6)+
  scale_color_manual(values=c("C1"="blue", "C2"="purple", "C3"="red"))
p


```

$~$

The following are for direct comparison of CP-based figures to previously created ScanR-based figures.



```{r count_and_intensity_boxplots_C1C2, fig.width= 8, fig.height=4}
p <- ggplot(IF_CP, aes(x=ligand_media, y=Well_Cell_Count, fill=factor(PairedWithEGF)))+
  geom_boxplot(position = position_dodge(width=0.9), outlier.colour = NA)+
  labs(x="Ligand",
       y="Well Cell Count",
       fill="Media with EGF",
       title="Cell Count by ligand and time point")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)),
        plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(1)),
        legend.title=element_text(size = rel(1)))+
  facet_wrap(~time)
p <- p +  geom_point(aes(colour=collection), position=position_jitterdodge(dodge.width=0.9), alpha=.8)+
  scale_color_manual(values=c("C1"="blue", "C2"="purple", "C3"="burgundy"))
print(p)

p <- ggplot(IF_CP, aes(x=factor(ligand_media), y=Intensity_MeanIntensity_KRT5, fill=factor(PairedWithEGF)))+
  geom_boxplot(position = position_dodge(width=0.9), outlier.colour = NA)+
  labs(x="Ligand",
       y="KRT5 Mean Intensity",
       fill="Media with EGF",
       title="KRT5 Mean Intensity by ligand and time point")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)),
        plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(1)),
        legend.title=element_text(size = rel(1)))+
  facet_wrap(~time)
p <- p +  geom_point(aes(colour=collection), position=position_jitterdodge(dodge.width=0.9), alpha=.8)+
  scale_color_manual(values=c("C1"="blue", "C2"="purple", "C3"="burgundy"))
print(p)

```


```{r count_and_intensity_batches_boxplots_C1C2, fig.width= 8, fig.height=4}
p <- ggplot(IF_CP, aes(x=ligand_media, y=Well_Cell_Count, colour=collection, fill=factor(PairedWithEGF)))+
  geom_boxplot(position = position_dodge(width=0.9), outlier.colour = NA)+
  labs(x="Ligand",
       y="Well Cell Count",
       fill="Media with EGF",
       title="Cell Count by ligand and time point")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)),
        plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.7)),
        legend.title=element_text(size = rel(1)))+
  facet_wrap(~time)
p <- p +  geom_point(aes(colour=collection), position=position_jitterdodge(dodge.width=0.9), size = 1, alpha=.6)+
  scale_color_manual(values=c("C1"="blue", "C2"="purple", "C3"="red"))
print(p)

p <- ggplot(IF_CP, aes(x=factor(ligand_media), y=Intensity_MeanIntensity_KRT5, colour=collection, fill=factor(PairedWithEGF)))+
  geom_boxplot(position = position_dodge(width=0.9), outlier.colour = NA)+
  labs(x="Ligand",
       y="KRT5 Mean Intensity",
       fill="Media with EGF",
       title="KRT5 Mean Intensity by ligand and time point")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)),
        plot.title = element_text(size = rel(1)), legend.text=element_text(size = rel(.7)),
        legend.title=element_text(size = rel(1)))+
  facet_wrap(~time)
p <- p +  geom_point(aes(colour=collection), position=position_jitterdodge(dodge.width=0.9), size = 1, alpha=.6)+
  scale_color_manual(values=c("C1"="blue", "C2"="purple", "C3"="red"))
print(p)

```


```{r PCA}

IF_CP_nna <- IF_CP %>%
  select(matches("specimenName|Intensity_|AreaShape_|Neighbor")) %>%
  drop_na()

IF_CP_PCA <- IF_CP_nna %>%
  select(-specimenName) %>%
  as.matrix() %>%
  prcomp()

```

```{r visualize_PCA}

df <- IF_CP_PCA[["x"]] %>%
  as_tibble() %>%
  mutate(specimenName = IF_CP_nna$specimenName,
         condition = str_remove(specimenName, "_C[123]_."),
         replicate = str_remove(specimenName, ".*_"),
         ligand = str_remove(condition, "_.*"),
         time = str_remove(condition, ".*_"),
         time = as.numeric(time))

p <- ggplot(df, aes(PC1, PC2, colour = ligand, size = time)) +
  geom_point(alpha = .7) +
  theme_bw()
p

p <- ggplot(df, aes(PC3, PC4, colour = ligand, size = time)) +
  geom_point(alpha = .7) +
  theme_bw()
p

p <- ggplot(df, aes(PC5, PC6, colour = ligand, size = time)) +
  geom_point(alpha = .7) +
  theme_bw()
p

p <- ggplot(df, aes(PC7, PC8, colour = ligand, size = time)) +
  geom_point(alpha = .7) +
  theme_bw()
p

plot(IF_CP_PCA)
```