---
title: "MDD interactive heatmaps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE,  message=FALSE, warning = FALSE, fig.height=10, fig.width=10)

suppressMessages(library(tidyverse))
suppressMessages(library(ComplexHeatmap))
library(iheatmapr)
suppressMessages(library(circlize))
library(umap)
```


## Overview{.tabset .tabset-fade}
These figures are derived from the data downloaded from the LINCS Molecular Deep Dive files on Synapse https://www.synapse.org/#!Synapse:syn2862345/wiki/588244.  


```{r standardFigureFunctions}

#' Read an assay's csv or tsv data file 
load_data <- function(assay_name, level = 4){
  file_name <- dir(pattern = paste0("MDD_", assay_name,"_","Level",level,".csv"),recursive = TRUE, full.names = TRUE)
  if(!length(file_name) == 1) stop("There was not one ", assay_name, " data file found")
    
  if(str_detect(file_name, "csv")) {
    df <- read_csv(file_name)
  } else if (str_detect(file_name, "tsv")){
    df <- read_tsv(file_name)
  } else stop("The data file ",file_name, " must be a csv or tsv file type")
} #end of function

get_win_probs <- function(mat, low = .02, hi = .98){
  windsorize_probs <- mat %>%
    quantile(probs = c(low, hi), na.rm = TRUE)
}
  
prep_data <- function(df, md){
    df <- df %>%
    gather(specimenID, value, -feature) %>%
    inner_join(md, by = "specimenID") %>%
    dplyr::select(feature, specimenName, value) %>%
    spread(specimenName, value) %>%
    mutate(feature = factor(feature, levels = unique(df$feature))) %>%
      arrange(feature,)
  #Set the order of the columns
  condition_order <-  paste(rep(c("ctrl","PBS", "HGF", "OSM", "EGF","BMP2", "IFNG", "TGFB"), each = 30, times = 3), rep(c(0, 1, 4 ,8, 24, 48), each = 5, times = 3), rep(c("C1", "C2", "C3"), each = 240), rep(c("A", "B", "C", "D", "E"), times = 109),  sep = "_") %>%
    intersect(colnames(df))
  
  #Arrange columns by standard condition order
  df <- df %>%
    dplyr::select(feature, condition_order)
  return(df)
}

prep_annotations <- function(df, md){
      df_ann <- df %>%
      gather("specimenName", "value", -feature) %>%
      inner_join(md, by = "specimenName") %>%
      dplyr::select(ligand, secondLigand, experimentalTimePoint, replicate, collection) %>%
      distinct() %>%
      rename(Time = experimentalTimePoint,
             Ligand = ligand,
             Ligand2 = secondLigand,
             Replicate = replicate,
             Collection = collection) %>%
      as.data.frame()
      return(df_ann)
}

# prep_annotations_conditions <- function(df, md){
#       df_ann <- df %>%
#       gather("condition", "value", -feature) %>%
#       inner_join(md, by = "specimenName") %>%
#       dplyr::select(ligand, secondLigand, experimentalTimePoint, replicate, collection) %>%
#       distinct() %>%
#       rename(Time = experimentalTimePoint,
#              Ligand = ligand,
#              Ligand2 = secondLigand,
#              Replicate = replicate,
#              Collection = collection) %>%
#       as.data.frame()
#       return(df_ann)
# }


heatmap_order <- function(df, md, EGF_normed = TRUE){
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md)
  if(EGF_normed){
    df <- df %>%
    dplyr::select(-matches("EGF"))
  }

df_as_matrix <- dplyr::select(df, -feature) %>%
  as.matrix() %>%
  t %>%
  scale() %>%
  t
  rownames(df_as_matrix) <- df$feature
   #Create the heatmap
  hm <- Heatmap(matrix = df_as_matrix,
          name = "abundance",
          show_column_names = FALSE,
          cluster_columns = FALSE,
          cluster_rows = TRUE,
          na_col = "grey")
  rowOrder <- row_order(hm)
  rowOrder <- rowOrder[length(rowOrder):1]
  return(rowOrder)
}

plot_iheatmap <- function(df, md, assay_name, EGF_normed = TRUE, var_thresh_percentile = 0, win_prob_low = .02, win_prob_up = .98){
  
    df_vars <- df %>%
    dplyr::select(-feature) %>%
    as.matrix() %>%
    apply(1, var, na.rm = TRUE)
  
  var_thresh <- quantile(df_vars, probs = var_thresh_percentile)
  df <- df[df_vars >= var_thresh,]
  
  #Get a dendrogram from the complex heatmap function
  rowOrder <- heatmap_order(df, md = md, EGF_normed = EGF_normed)

  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md)
  
  #Create annotation values
  df_ann <-prep_annotations(df, md) %>%
    dplyr::select(Collection, Replicate, Time, Ligand2, Ligand)
  
  title_suffix <- ""
  if(EGF_normed){
    title_suffix = ", EGF timecourse normalized"
    df <- df %>%
      dplyr::select(-matches("EGF"))
    df_ann <- df_ann %>%
      filter(!Ligand == "EGF")
  }
 
  df_as_matrix <- dplyr::select(df, -feature) %>%
    as.matrix() %>%
    t %>%
  scale() %>%
  t
  rownames(df_as_matrix) <- df$feature
  
windsorize_probs <- get_win_probs(mat = df_as_matrix, win_prob_low, win_prob_up)

  #Create the heatmap
  hm <- main_heatmap(data = df_as_matrix,
                     name = "abundance",
                     zmid = 0,
                     zmin = windsorize_probs[1],  
                     zmax = windsorize_probs[2],
                     row_order = rowOrder) %>%
    add_row_clustering() %>%
    add_col_annotation(df_ann,
                       colors = list(Ligand = c( "ctrl" = "#7A4A2A",
                                                 "PBS" = "#8dd3c7",
                                                 "HGF" = "#80b1d3",
                                                 "OSM" = "#fdb462",
                                                 "EGF" = "#fb8072",
                                                 "BMP2" = "#b3de69",
                                                 "IFNG" = "#bebada",
                                                 "TGFB" = "#ffd92f"),
                                     Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                     Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                     Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                     Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")),
                       size = 0.02)
  hm
}
plot_iheatmap_conditions <- function(df, assay_name){
  
  #reorder
  
    #Set the order of the columns
  condition_order <-  paste(rep(c("ctrl", "HGF", "OSM", "EGF","BMP2", "IFNG", "TGFB"), each = 2), c(24, 48),  sep = "_") %>%
    intersect(colnames(df))
  
  #Arrange columns by standard condition order
  df <- df %>%
    dplyr::select(feature, condition_order)

  # #Create annotation values
  # df_ann <-prep_annotations_conditions(df, md) %>%
  #   dplyr::select(Collection, Replicate, Time, Ligand2, Ligand)
  # 
 
 
  df_as_matrix <- dplyr::select(df, -feature) %>%
    as.matrix() %>%
    t %>%
  scale() %>%
  t
  rownames(df_as_matrix) <- df$feature
  
#windsorize_probs <- get_win_probs(mat = df_as_matrix, win_prob_low, win_prob_up)

  #Create the heatmap
  hm <- main_heatmap(data = df_as_matrix,
                     name = "abundance",
                     show_col_names = TRUE) %>%
    add_row_clustering() %>%
     add_col_annotation(df_ann,
                       colors = list(Ligand = c( "ctrl" = "#7A4A2A",
                                                 "PBS" = "#8dd3c7",
                                                 "HGF" = "#80b1d3",
                                                 "OSM" = "#fdb462",
                                                 "EGF" = "#fb8072",
                                                 "BMP2" = "#b3de69",
                                                 "IFNG" = "#bebada",
                                                 "TGFB" = "#ffd92f"),
                                     Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black")),
                       size = 0.02)
  hm
}

plot_line_graphs <- function(df, md, ligand_cols, EGF_normed = TRUE, fc_thresh = 1, assay_name){
  
  windsorize_probs <- get_win_probs(df, .01, .99)
  
  df_sum <- df %>%
    prep_data(md = md) %>%
    gather("specimenName", "value", -feature) %>%
    inner_join(md, by = "specimenName") %>%
    dplyr::select(ligand, secondLigand, experimentalTimePoint, replicate, collection, feature, value) %>%
    rename(Time = experimentalTimePoint,
           Ligand = ligand,
           Ligand2 = secondLigand,
           Replicate = replicate,
           Collection = collection) %>%
    group_by(feature, Collection, Time, Ligand) %>%
    summarise(value = median(value)) %>%
    ungroup() 
  
  if(0 %in% unique(df_sum$Time)){
    df2 <- df_sum %>%
      filter(Time == 0) %>%
      data.frame(Ligand = rep(unique(df_sum$Ligand), each = length(unique(df_sum$feature))),
                 stringsAsFactors = FALSE) %>%
      dplyr::select(feature, Ligand.1, value, Collection, Time) %>%
      rename(Ligand = Ligand.1) %>%
      bind_rows(df_sum) %>%
      filter(!Ligand == "ctrl") 
  } else {
    df2 <- df_sum
  }
  
  
  fc_set <- df2 %>%
    group_by(Time, Ligand) %>%
    filter(abs(value) > fc_thresh) %>%
    ungroup() %>%
    dplyr::select(feature) %>%
    distinct() %>%
    inner_join(df2, by = "feature") %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
  title_suffix <- ""
  if(EGF_normed){
    title_suffix <- ", EGF normalized"
    filter(fc_set, !ligand == "EGF")
  }
  p <- ggplot(fc_set, aes(x=Time, y=value, colour=Ligand))+
    geom_line()+
    labs(title=paste0("Line Graphs",title_suffix),
         x = "Time (hours)",
         y="Intensity (AU)") +
    scale_x_continuous(breaks = c(0,8,24,48)) +
    scale_color_manual(values = ligand_cols) +
    theme(axis.text.x = element_text(angle=90),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          strip.text = element_text(size=10, hjust=0),
            panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "gray95"))+
    facet_wrap(~feature, ncol = 10,scales = "free_y")
  p
  
  }
  
plot_umap <- function(df, md, ligand_cols = ligand_cols, assay_name,
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000"), select_timepoint = FALSE, ...){
  #Generate a 2d scatterplot of the UMAP dimensions
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  #windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md) %>%
    dplyr::select(-matches("EGF"))
  
    if(select_timepoint){
    df <- df %>%
    dplyr::select(matches(paste0("_",select_timepoint,"_","|feature")))
  }

   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    dplyr::select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature

  df_UMAP <- umap(df_mat, na.rm = TRUE, ...)$layout %>%
  data.frame() %>%
  rename(UMAP_1 = X1,
         UMAP_2 = X2) %>%
    cbind(df_ann) %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
p <- ggplot(df_UMAP, aes(x = UMAP_1,
                         y = UMAP_2,
                         size = Time,
                         colour = Ligand,
                         fill = Ligand2)) +
    geom_point(alpha = .8) +
  geom_point(shape = 21, stroke = 4, alpha = .8) +
  scale_color_manual( values = ligand_cols) +
  scale_fill_manual( values = ligand_2_cols) +
  scale_size(breaks = c(1, 4,8,24,48)) +
  labs(title = paste("UMAP embedding of ", assay_name, " data")) +
  theme_bw()
p
  
}

plot_pca <- function(df, md, ligand_cols = ligand_cols, assay_name,
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #Generate a 2d scatterplot of the first two princpal components
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  #windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimenName and reorder
  df <- prep_data(df, md) %>%
    dplyr::select(-matches("EGF"))
  
   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    dplyr::select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature

    df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann)
    
  df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann) %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
p <- ggplot(df_pca, aes(x = PC1,
                         y = PC2,
                         size = Time,
                         colour = Ligand,
                         fill = Ligand2)) +
  geom_point(alpha = .8) +
  geom_point(shape = 21, stroke = 4, alpha = .8) +
  scale_color_manual( values = ligand_cols) +
  scale_fill_manual( values = ligand_2_cols) +
  scale_size_area(max_size = 10, breaks = c(1,4,8,24,48)) +
  labs(title = paste("PCA of ", assay_name, " data")) + 
  theme_bw()
p
}

 #Create the heatmap column annotations
create_top_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md)
  haTop <- HeatmapAnnotation(df = df_ann,
                             col = list(Ligand = c("ctrl" = "#7A4A2A",
                                                   "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072",
                                                    "BMP2" = "#b3de69",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#ffd92f"),
                                        Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haTop)
}

prep_top_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md)
  haTop <- HeatmapAnnotation(df = df_ann,
                             col = list(Ligand = c("ctrl" = "#7A4A2A",
                                                   "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072",
                                                    "BMP2" = "#b3de69",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#ffd92f"),
                                        Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haTop)
}


plot_correlation <- function(df, md, assay_name, EGF_normed = TRUE,
                      ligand_cols = c( "PBS" = "#8dd3c7",
                                      "HGF" = "#80b1d3",
                                      "OSM" = "#fdb462",
                                      "EGF" = "#fb8072",
                                      "BMP2" = "#b3de69",
                                      "IFNG" = "#bebada",
                                      "TGFB" = "#ffd92f"),
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #calculate correlations across the conditions and show in a heatmap

  #convert specimenID column names to specimanName, reorder delete EGF samples
  df <- prep_data(df, md) 
  
  title_suffix <- ""
  if(EGF_normed){
    title_suffix <- ", EGF Normalized"
    df <- df %>%
    dplyr::select(-matches("EGF"))
  }
  #Create annotation values
  #df_ann <-prep_annotations(df, md)
  
  haRow <- create_row_annotations(df, md)
  
  #Create the heatmap
  Heatmap(matrix = dplyr::select(df, -feature) %>%
            as.matrix() %>%
            t %>%
            scale(scale = FALSE) %>%
            t %>%
            cor(use = "complete.obs",method = "spearman"),
          name = "correlation",
          column_title = paste0("Correlation of ",assay_name, title_suffix),
          top_annotation = create_top_annotations(df, md),
          left_annotation = haRow,
          show_row_names = FALSE,
          show_column_names = FALSE,
          cluster_rows = FALSE,
          cluster_columns = FALSE,
          col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")
}

create_row_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md) %>%
    dplyr::select(Collection, Replicate, Time, Ligand2, Ligand)
  haRow <- rowAnnotation(df = df_ann, show_legend = FALSE,
                             col = list(Ligand = c( "ctrl" = "#7A4A2A",
                                                    "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072",
                                                    "BMP2" = "#b3de69",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#ffd92f"),
                                        Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haRow)
}

  

```

```{r globalVariables}
ligand_cols <- c("ctrl" = "#7A4A2A",
                 "PBS" = "#8dd3c7",
                  "HGF" = "#80b1d3",
                  "OSM" = "#fdb462",
                  "EGF" = "#fb8072",
                  "BMP2" = "#b3de69",
                  "IFNG" = "#bebada",
                  "TGFB" = "#ffd92f")


ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")


md <- read_csv("metadata/MDD_sample_annotations.csv")
```

### RPPA{.tabset .tabset-fade}

#### Heatmap

```{r rppa_figures}

RPPA_l1 <- load_data(assay_name = "RPPA", level = 1) %>%
  rename(feature = antibody) %>%
  gather(key = specimenID, value = value, -feature) %>%
  group_by(feature) %>%
  mutate(value = value - median(value, na.rm = TRUE)) %>%
  ungroup() %>%
  spread(key = specimenID, value = value) %>%
  dplyr::select(-sid64)

```


```{r}
var_thresh_percentile <- 0

plot_iheatmap(RPPA_l1, md, assay_name = "RPPA", EGF_normed = FALSE, var_thresh_percentile = var_thresh_percentile, win_prob_low = .02, win_prob_up = .98)

```

Antibodies with variances above the `r var_thresh_percentile*100` percentile are shown in the heatmaps.  


### cycIF{.tabset .tabset-fade}

#### Heatmap 

```{r cycIF_figures, eval = TRUE}

cycIF <- load_data(assay_name = "cycIF", level = 3) %>%
  filter(str_detect(feature, "_int_")) %>%
  mutate_if(is.numeric, log2)

cycIF_vars <- cycIF %>%
  dplyr::select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh <- quantile(cycIF_vars, probs = var_thresh_percentile)
cycIF_hi_vars <- cycIF[cycIF_vars > var_thresh,]

plot_iheatmap(cycIF_hi_vars, md, assay_name = "cycIF")

```


Intensity features included in the heatmap have a variance above the `r var_thresh_percentile` percentile . 


### RNAseq{.tabset .tabset-fade}

#### Heatmap  

```{r rnaSeq_figures, eval = TRUE}
RNAseq <- load_data(assay_name = "RNAseq", level = 2) %>%
  rename(feature = ensembl_gene_id) %>%
  dplyr::select(-sid64)

RNAseq_vars <- RNAseq %>%
  dplyr::select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh_percentile <- .95
var_thresh <- quantile(RNAseq_vars, probs = var_thresh_percentile)
RNAseq_hi_vars <- RNAseq[RNAseq_vars > var_thresh,]
  
plot_iheatmap(RNAseq_hi_vars, md, assay_name = "RNAseq")

```


Genes included in the heatmap have a variance above the `r var_thresh_percentile` percentile . 


### TF_Pathways{.tabset .tabset-fade}

#### Heatmap  

```{r TF_pathways_figures}
reverseTFScore <- function(x, max_TF_mean_rank){
  x <- max_TF_mean_rank - x
}
  TFs <- read_csv("RNAseq/Data/ChEA3_TFs.csv") %>%
    mutate(condition = str_remove(specimenName, "h$")) %>%
    dplyr::select(-specimenName)
  
#reverse the TF values so that higher values represent higher TF scores
  max_TF_mean_rank <- TFs %>%
    dplyr::select(-condition) %>%
    max
  TFs <- TFs %>%
    mutate_if(is.numeric, reverseTFScore, max_TF_mean_rank = max_TF_mean_rank) %>%
    gather(key = "feature", value="value",-condition)

  RPPA_l1_pathways <- load_data("RPPA", "1_Pathways_Score") %>%
  rename(specimenID = X1) %>%
  gather(key = feature, value = Score, -specimenID) %>%
  spread(key = specimenID, value = Score)  %>%
  dplyr::select(-sid64) %>%
    gather(specimenID, value, -feature) %>%
  inner_join(md, by = "specimenID") %>%
   dplyr::select(specimenName, feature, value) %>%
    mutate(condition = str_remove(specimenName, "_C1.*")) %>%
    group_by(condition, feature) %>%
    summarise_at("value", mean) %>%
  ungroup
  
TF_pathways <-  TFs %>%
  bind_rows(RPPA_l1_pathways) %>%
  spread(key = condition, value = value) %>%
  select(matches("_24|_48|feature"), -matches("PBS"))
  
TF_pathways_cor <- TF_pathways %>%
  gather(key = "condition", value = value, -feature) %>%
  spread(key = feature, value = value) %>%
  select(-condition) %>%
  cor(method = "spearman")

```



```{r TFxPathwayCorrelations}

df <- TF_pathways_cor %>%
  as_tibble(rownames="TF") %>%
  dplyr::select(matches("_Score|^TF$")) %>%
  filter(!str_detect(TF,"_Score"))

df_as_matrix <- df %>%
  select(-TF) %>%
  as.matrix()

Heatmap(matrix = df_as_matrix,
          name = "correlations",
          column_title = paste0("Correlation of all TFs and pathways"),
          show_row_names = FALSE,
          show_column_names = TRUE,
          cluster_rows = TRUE,
          cluster_columns = TRUE,
          col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")

```


Method Summary  

Start with DCIC's preprocessed RNAseq values  
Filter to genes with logFC >1  
For each specimen, feed selected into ChEA3  
Reverse ChEa3 TF mean rankings so that high values represent highly scored TFs  
Mean summarize replicate RPPA pathway scores  
Find the correlations betwwen all TFs and pathways across all 24 and 48 hour conditions and display as a heatmap  


#### Top TFs correlations  

```{r top_TF_pathways_figures, eval = TRUE}

  #Select top TFs of each ligand
  top_TF_names <- TFs %>%
  spread(key = feature, value = value) %>%
    mutate(Ligand = str_remove(condition, "_.*")) %>%
    group_by(Ligand) %>%
    summarise_if(is.numeric, max) %>%
    ungroup %>%
    gather(key = TF, value = Rank, -Ligand) %>%
    group_by(Ligand) %>%
    top_n(25, Rank) %>%
    ungroup() %>%
    spread(key = TF, value = Rank) %>%
    dplyr::select(-Ligand)
  
  top_TFs <- TFs %>%
  spread(key = feature, value = value)  %>%
    dplyr::select(condition, colnames(top_TF_names))
  
  top_TF_pathways <-  top_TFs %>%
    gather(key = feature,value = value, -condition) %>%
  bind_rows(RPPA_l1_pathways) %>%
  spread(key = condition, value = value) %>%
  select(matches("_24|_48|feature"), -matches("PBS"))
  
    
top_TF_pathways_cor <- top_TF_pathways %>%
  gather(key = "condition", value = value, -feature) %>%
  spread(key = feature, value = value) %>%
  select(-condition) %>%
  cor(method = "spearman")

  
#plot_iheatmap(TF_pathways, md, assay_name = "TFs and RPPA pathways")

```

```{r topTFxPathwayCorrelations}

df <- top_TF_pathways_cor %>%
  as_tibble(rownames="TF") %>%
  dplyr::select(matches("_Score|^TF$")) %>%
  filter(!str_detect(TF,"_Score"))

df_as_matrix <- df %>%
  select(-TF) %>%
  as.matrix()

Heatmap(matrix = df_as_matrix,
          name = "correlations",
          column_title = paste0("Correlation of top TFs and pathways"),
          show_row_names = FALSE,
          show_column_names = TRUE,
          cluster_rows = TRUE,
          cluster_columns = TRUE,
          col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")

```

Method Summary  

Z score RNAseq genes across all conditions  
For each specimen, feed genes with Z score > 2.5 into ChEA3  
Reverse ChEa3 TF mean rankings so that high values represent highly scored TFs  
Group by ligand  
Select the top ranked 25 TFs in all timepoints of each ligand  
Combine the top ranked TFs into one set  
Filter the reversed TF rankings to only include the TFs in the top ranked set  
Combine RPPA pathway z score with TF Z scores  


#### Line graphs

```{r, fig.height=30, fig.width=15}

#pathway line graphs
  df_sum <- TF_pathways %>%
    prep_data(md = md) %>%
    gather("specimenName", "value", -feature) %>%
    inner_join(md, by = "specimenName") %>%
    dplyr::select(ligand, secondLigand, experimentalTimePoint, replicate, collection, feature, value) %>%
    rename(Time = experimentalTimePoint,
           Ligand = ligand,
           Ligand2 = secondLigand,
           Replicate = replicate,
           Collection = collection) %>%
    group_by(feature, Collection, Time, Ligand) %>%
    summarise(value = median(value)) %>%
    ungroup() 
  
   if(0 %in% unique(df_sum$Time)){
    df2 <- df_sum %>%
      filter(Time == 0) %>%
      data.frame(Ligand = rep(unique(df_sum$Ligand), each = length(unique(df_sum$feature))),
                 stringsAsFactors = FALSE) %>%
      dplyr::select(feature, Ligand.1, value, Collection, Time) %>%
      rename(Ligand = Ligand.1) %>%
      bind_rows(df_sum) %>%
      filter(!Ligand == "ctrl") 
  } else {
    df2 <- df_sum
  }
  
  
  fc_set <- df2 %>%
    dplyr::select(feature) %>%
    distinct() %>%
    inner_join(df2, by = "feature") %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  

  
p <- ggplot(fc_set, aes(x=Time, y=value, colour=Ligand))+
  geom_line()+
  labs(title=paste0("TF pathways"),
       x = "Time (hours)",
       y="Intensity (AU)") +
  scale_x_continuous(breaks = c(0,8,24,48)) +
  scale_color_manual(values = ligand_cols) +
  theme(axis.text.x = element_text(angle=90),
        # axis.text.y=element_blank(),
        # axis.ticks.y=element_blank(),
        strip.text = element_text(size=6, hjust=0),
        panel.background = element_rect(fill = NA),
        panel.grid.major = element_line(colour = "gray95"))+
  facet_wrap(~feature, ncol = 8,scales = "free_y")
p
  
```

##