---
title: "MDD interactive heatmaps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE,  message=FALSE, warning = FALSE, fig.height=10, fig.width=10)

suppressMessages(library(tidyverse))
suppressMessages(library(ComplexHeatmap))
library(iheatmapr)
suppressMessages(library(circlize))
library(umap)
```


## Overview{.tabset .tabset-fade}
These figures are derived from the data downloaded from the LINCS Molecular Deep Dive files on Synapse https://www.synapse.org/#!Synapse:syn2862345/wiki/588244.  


```{r standardFigureFunctions}

#' Read an assay's csv or tsv data file 
load_data <- function(assay_name, level = 4){
  file_name <- dir(pattern = paste0("MDD_", assay_name,"_","Level",level,".csv"),recursive = TRUE, full.names = TRUE)
  if(!length(file_name) == 1) stop("There was not one ", assay_name, " data file found")
    
  if(str_detect(file_name, "csv")) {
    df <- read_csv(file_name)
  } else if (str_detect(file_name, "tsv")){
    df <- read_tsv(file_name)
  } else stop("The data file ",file_name, " must be a csv or tsv file type")
} #end of function

get_win_probs <- function(mat, low = .02, hi = .98){
  windsorize_probs <- mat %>%
    quantile(probs = c(low, hi), na.rm = TRUE)
}
  
prep_data <- function(df, md){
    df <- df %>%
    gather(specimenID, value, -feature) %>%
    inner_join(md, by = "specimenID") %>%
    dplyr::select(feature, specimenName, value) %>%
    spread(specimenName, value) %>%
    mutate(feature = factor(feature, levels = unique(df$feature))) %>%
      arrange(feature,)
  #Set the order of the columns
  condition_order <-  paste(rep(c("ctrl","PBS", "HGF", "OSM", "EGF","BMP2", "IFNG", "TGFB"), each = 30, times = 3), rep(c(0, 1, 4 ,8, 24, 48), each = 5, times = 3), rep(c("C1", "C2", "C3"), each = 240), rep(c("A", "B", "C", "D", "E"), times = 109),  sep = "_") %>%
    intersect(colnames(df))
  
  #Arrange columns by standard condition order
  df <- df %>%
    dplyr::select(feature, condition_order)
  return(df)
}

prep_annotations <- function(df, md){
      df_ann <- df %>%
      gather("specimenName", "value", -feature) %>%
      inner_join(md, by = "specimenName") %>%
      dplyr::select(ligand, secondLigand, experimentalTimePoint, replicate, collection) %>%
      distinct() %>%
      rename(Time = experimentalTimePoint,
             Ligand = ligand,
             Ligand2 = secondLigand,
             Replicate = replicate,
             Collection = collection) %>%
      as.data.frame()
      return(df_ann)
}

prep_annotations_conditions <- function(df){
      df_ann <- df %>%
      gather("condition", "value", -feature) %>%
        mutate(Ligand = str_remove(condition, "_.*"),
               Time = str_remove(condition,".*_")) %>%
      dplyr::select(Ligand, Time) %>%
      distinct() %>%
      as.data.frame()
      return(df_ann)
}


plot_heatmap <- function(df, md, assay_name, EGF_normed = TRUE, var_thresh_percentile = 0, show_row_names = FALSE, mark_rows = NULL, cluster_rows = TRUE, select_timepoint = FALSE, zscore_rows = TRUE,...) {
  
  df_vars <- df %>%
  dplyr::select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(df_vars, probs = var_thresh_percentile, na.rm = TRUE)
df <- df[df_vars >= var_thresh,]
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md)
title_suffix <- ""
  if(EGF_normed){
    title_suffix = ", EGF timecourse normalized"
    df <- df %>%
    dplyr::select(-matches("EGF"))
  }

  if(select_timepoint){
    df <- df %>%
    dplyr::select(matches(paste0("_",select_timepoint,"_","|feature")))
  }

  if(!is.null(mark_rows)) {
    row_nums <- which(df$feature %in% mark_rows)
    ra <- rowAnnotation(foo = anno_mark(at = c(row_nums), labels = df$feature[row_nums]))
  } else {
    ra <- NULL
  }
if(zscore_rows) {
  df_as_matrix <- dplyr::select(df, -feature) %>%
    as.matrix() %>%
    t %>%
    scale() %>%
    t
} else {
  df_as_matrix <- dplyr::select(df, -feature) %>%
    as.matrix()
}
  rownames(df_as_matrix) <- df$feature

 windsorize_probs <- df_as_matrix %>%
      unlist %>% 
   quantile(probs = c(.02, .98), na.rm = TRUE)
 
   #Create the heatmap
  hm <- Heatmap(matrix = df_as_matrix,
          name = "abundance",
          column_title = paste0(assay_name, title_suffix),
          top_annotation = create_top_annotations(df, md),
          right_annotation = ra,
          show_row_names = show_row_names,
          show_column_names = FALSE,
          cluster_columns = FALSE,
          cluster_rows = cluster_rows,
          col = circlize::colorRamp2(c(windsorize_probs[1], 0, windsorize_probs[2]), c("blue", "white", "red")),
          na_col = "grey",...)
  hm
}


heatmap_order <- function(df, md, EGF_normed = TRUE){
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md)
  if(EGF_normed){
    df <- df %>%
    dplyr::select(-matches("EGF"))
  }

df_as_matrix <- dplyr::select(df, -feature) %>%
  as.matrix() %>%
  t %>%
  scale() %>%
  t
  rownames(df_as_matrix) <- df$feature
   #Create the heatmap
  hm <- Heatmap(matrix = df_as_matrix,
          name = "abundance",
          show_column_names = FALSE,
          cluster_columns = FALSE,
          cluster_rows = TRUE,
          na_col = "grey")
  rowOrder <- row_order(hm)
  rowOrder <- rowOrder[length(rowOrder):1]
  return(rowOrder)
}

plot_iheatmap <- function(df, md, assay_name, EGF_normed = TRUE, var_thresh_percentile = 0, win_prob_low = .02, win_prob_up = .98){
  
    df_vars <- df %>%
    dplyr::select(-feature) %>%
    as.matrix() %>%
    apply(1, var, na.rm = TRUE)
  
  var_thresh <- quantile(df_vars, probs = var_thresh_percentile, na.rm = TRUE)
  df <- df[df_vars >= var_thresh,]
  
  #Get a dendrogram from the complex heatmap function
  rowOrder <- heatmap_order(df, md = md, EGF_normed = EGF_normed)

  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md)
  
  #Create annotation values
  df_ann <-prep_annotations(df, md) %>%
    dplyr::select(Collection, Replicate, Time, Ligand2, Ligand)
  
  title_suffix <- ""
  if(EGF_normed){
    title_suffix = ", EGF timecourse normalized"
    df <- df %>%
      dplyr::select(-matches("EGF"))
    df_ann <- df_ann %>%
      filter(!Ligand == "EGF")
  }
 
  df_as_matrix <- dplyr::select(df, -feature) %>%
    as.matrix() %>%
    t %>%
  scale() %>%
  t
  rownames(df_as_matrix) <- df$feature
  
windsorize_probs <- get_win_probs(mat = df_as_matrix, win_prob_low, win_prob_up)

  #Create the heatmap
  hm <- main_heatmap(data = df_as_matrix,
                     name = "abundance",
                     zmid = 0,
                     zmin = windsorize_probs[1],  
                     zmax = windsorize_probs[2],
                     row_order = rowOrder) %>%
    add_row_clustering() %>%
    add_col_annotation(df_ann,
                       colors = list(Ligand = c( "ctrl" = "#7A4A2A",
                                                 "PBS" = "#8dd3c7",
                                                 "HGF" = "#80b1d3",
                                                 "OSM" = "#fdb462",
                                                 "EGF" = "#fb8072",
                                                 "BMP2" = "#b3de69",
                                                 "IFNG" = "#bebada",
                                                 "TGFB" = "#ffd92f"),
                                     Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                     Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                     Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                     Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")),
                       size = 0.02)
  hm
}
plot_iheatmap_conditions <- function(df, assay_name){
  
  #reorder
  
    #Set the order of the columns
  condition_order <-  paste(rep(c("HGF", "OSM", "EGF","BMP2", "IFNG", "TGFB"), each = 2), c(24, 48),  sep = "_") %>%
    intersect(colnames(df))
  
  #Arrange columns by standard condition order
  df <- df %>%
    dplyr::select(feature, condition_order) %>%
    drop_na()

  #Create annotation values
  df_ann <-prep_annotations_conditions(df)

  df_as_matrix <- dplyr::select(df, -feature) %>%
    as.matrix() %>%
    t %>%
  scale() %>%
  t
  rownames(df_as_matrix) <- df$feature
  
#windsorize_probs <- get_win_probs(mat = df_as_matrix, win_prob_low, win_prob_up)

  #Create the heatmap
  hm <- main_heatmap(data = df_as_matrix,
                     name = "abundance") %>%
    add_row_clustering() %>%
     add_col_annotation(df_ann,
                       size = 0.02)
  hm
}

plot_line_graphs <- function(df, md, ligand_cols, EGF_normed = TRUE, fc_thresh = 1, assay_name){
  
  windsorize_probs <- get_win_probs(df, .01, .99)
  
  df_sum <- df %>%
    prep_data(md = md) %>%
    gather("specimenName", "value", -feature) %>%
    inner_join(md, by = "specimenName") %>%
    dplyr::select(ligand, secondLigand, experimentalTimePoint, replicate, collection, feature, value) %>%
    rename(Time = experimentalTimePoint,
           Ligand = ligand,
           Ligand2 = secondLigand,
           Replicate = replicate,
           Collection = collection) %>%
    group_by(feature, Collection, Time, Ligand) %>%
    summarise(value = median(value)) %>%
    ungroup() 
  
  if(0 %in% unique(df_sum$Time)){
    df2 <- df_sum %>%
      filter(Time == 0) %>%
      data.frame(Ligand = rep(unique(df_sum$Ligand), each = length(unique(df_sum$feature))),
                 stringsAsFactors = FALSE) %>%
      dplyr::select(feature, Ligand.1, value, Collection, Time) %>%
      rename(Ligand = Ligand.1) %>%
      bind_rows(df_sum) %>%
      filter(!Ligand == "ctrl") 
  } else {
    df2 <- df_sum
  }
  
  
  fc_set <- df2 %>%
    group_by(Time, Ligand) %>%
    filter(abs(value) > fc_thresh) %>%
    ungroup() %>%
    dplyr::select(feature) %>%
    distinct() %>%
    inner_join(df2, by = "feature") %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
  title_suffix <- ""
  if(EGF_normed){
    title_suffix <- ", EGF normalized"
    filter(fc_set, !ligand == "EGF")
  }
  p <- ggplot(fc_set, aes(x=Time, y=value, colour=Ligand))+
    geom_line()+
    labs(title=paste0("Line Graphs",title_suffix),
         x = "Time (hours)",
         y="Intensity (AU)") +
    scale_x_continuous(breaks = c(0,8,24,48)) +
    scale_color_manual(values = ligand_cols) +
    theme(axis.text.x = element_text(angle=90),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          strip.text = element_text(size=10, hjust=0),
            panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "gray95"))+
    facet_wrap(~feature, ncol = 10,scales = "free_y")
  p
  
  }
  
plot_umap <- function(df, md, ligand_cols = ligand_cols, assay_name,
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000"), select_timepoint = FALSE, ...){
  #Generate a 2d scatterplot of the UMAP dimensions
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  #windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md) %>%
    dplyr::select(-matches("EGF"))
  
    if(select_timepoint){
    df <- df %>%
    dplyr::select(matches(paste0("_",select_timepoint,"_","|feature")))
  }

   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    dplyr::select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature

  df_UMAP <- umap(df_mat, na.rm = TRUE, ...)$layout %>%
  data.frame() %>%
  rename(UMAP_1 = X1,
         UMAP_2 = X2) %>%
    cbind(df_ann) %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
p <- ggplot(df_UMAP, aes(x = UMAP_1,
                         y = UMAP_2,
                         size = Time,
                         colour = Ligand,
                         fill = Ligand2)) +
    geom_point(alpha = .8) +
  geom_point(shape = 21, stroke = 4, alpha = .8) +
  scale_color_manual( values = ligand_cols) +
  scale_fill_manual( values = ligand_2_cols) +
  scale_size(breaks = c(1, 4,8,24,48)) +
  labs(title = paste("UMAP embedding of ", assay_name, " data")) +
  theme_bw()
p
  
}

plot_pca <- function(df, md, ligand_cols = ligand_cols, assay_name,
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #Generate a 2d scatterplot of the first two princpal components
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  #windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimenName and reorder
  df <- prep_data(df, md) %>%
    dplyr::select(-matches("EGF"))
  
   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    dplyr::select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature

    df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann)
    
  df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann) %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
p <- ggplot(df_pca, aes(x = PC1,
                         y = PC2,
                         size = Time,
                         colour = Ligand,
                         fill = Ligand2)) +
  geom_point(alpha = .8) +
  geom_point(shape = 21, stroke = 4, alpha = .8) +
  scale_color_manual( values = ligand_cols) +
  scale_fill_manual( values = ligand_2_cols) +
  scale_size_area(max_size = 10, breaks = c(1,4,8,24,48)) +
  labs(title = paste("PCA of ", assay_name, " data")) + 
  theme_bw()
p
}

 #Create the heatmap column annotations
create_top_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md)
  haTop <- HeatmapAnnotation(df = df_ann,
                             col = list(Ligand = c("ctrl" = "#7A4A2A",
                                                   "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072",
                                                    "BMP2" = "#b3de69",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#ffd92f"),
                                        Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haTop)
}

prep_top_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md)
  haTop <- HeatmapAnnotation(df = df_ann,
                             col = list(Ligand = c("ctrl" = "#7A4A2A",
                                                   "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072",
                                                    "BMP2" = "#b3de69",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#ffd92f"),
                                        Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haTop)
}


plot_correlation <- function(df, md, assay_name, EGF_normed = TRUE,
                      ligand_cols = c( "PBS" = "#8dd3c7",
                                      "HGF" = "#80b1d3",
                                      "OSM" = "#fdb462",
                                      "EGF" = "#fb8072",
                                      "BMP2" = "#b3de69",
                                      "IFNG" = "#bebada",
                                      "TGFB" = "#ffd92f"),
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #calculate correlations across the conditions and show in a heatmap

  #convert specimenID column names to specimanName, reorder delete EGF samples
  df <- prep_data(df, md) 
  
  title_suffix <- ""
  if(EGF_normed){
    title_suffix <- ", EGF Normalized"
    df <- df %>%
    dplyr::select(-matches("EGF"))
  }
  #Create annotation values
  #df_ann <-prep_annotations(df, md)
  
  haRow <- create_row_annotations(df, md)
  
  #Create the heatmap
  Heatmap(matrix = dplyr::select(df, -feature) %>%
            as.matrix() %>%
            t %>%
            scale(scale = FALSE) %>%
            t %>%
            cor(use = "complete.obs",method = "spearman"),
          name = "correlation",
          column_title = paste0("Correlation of ",assay_name, title_suffix),
          top_annotation = create_top_annotations(df, md),
          left_annotation = haRow,
          show_row_names = FALSE,
          show_column_names = FALSE,
          cluster_rows = FALSE,
          cluster_columns = FALSE,
          col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")
}

create_row_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md) %>%
    dplyr::select(Collection, Replicate, Time, Ligand2, Ligand)
  haRow <- rowAnnotation(df = df_ann, show_legend = FALSE,
                             col = list(Ligand = c( "ctrl" = "#7A4A2A",
                                                    "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072",
                                                    "BMP2" = "#b3de69",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#ffd92f"),
                                        Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haRow)
}

  

```

```{r globalVariables}
ligand_cols <- c("ctrl" = "#7A4A2A",
                 "PBS" = "#8dd3c7",
                  "HGF" = "#80b1d3",
                  "OSM" = "#fdb462",
                  "EGF" = "#fb8072",
                  "BMP2" = "#b3de69",
                  "IFNG" = "#bebada",
                  "TGFB" = "#ffd92f")


ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")


md <- read_csv("metadata/MDD_sample_annotations.csv")
```

### RPPA{.tabset .tabset-fade}

#### Heatmap

```{r rppa_figures}

RPPA_l1 <- load_data(assay_name = "RPPA", level = 1) %>%
  rename(feature = antibody) %>%
  gather(key = specimenID, value = value, -feature) %>%
  group_by(feature) %>%
  mutate(value = value - median(value, na.rm = TRUE)) %>%
  ungroup() %>%
  spread(key = specimenID, value = value) %>%
  dplyr::select(-sid64)

#T0 normalize within each replicate batch
RPPA_l1_all <- load_data(assay_name = "RPPA", level = 1) %>%
  rename(feature = antibody) %>%
  gather("specimenID", "value", -feature) %>%
  inner_join(md) 

RPPA_T0Norm <- RPPA_l1_all %>%
  filter(experimentalTimePoint == 0) %>%
  select(feature, replicate, value) %>%
  rename(value_T0 = value) %>%
  right_join(RPPA_l1_all) %>%
  group_by(feature, replicate) %>%
  mutate(value_T0Norm = value - value_T0) %>%
  ungroup %>%
  select(feature, value_T0Norm, specimenName, specimenID) %>%
  filter(!str_detect(specimenName, "ctrl|PBS")) %>%
  rename(value = value_T0Norm) %>%
  select(-specimenID)

```


```{r, eval = FALSE}
var_thresh_percentile <- 0

plot_iheatmap_conditions(RPPA_T0Norm, assay_name = "RPPA")

```


### RNAseq{.tabset .tabset-fade}

#### Heatmap  

```{r rnaSeq_figures, eval = FALSE}
fns <- dir("RNAseq/Data", pattern = "h.txt", full.names = TRUE)

RNAseq_data <- fns %>%
  map(read_tsv)
names(RNAseq_data) <- fns %>%
  str_remove_all(".*/Data/|h.txt")

RNAseq <- bind_rows(RNAseq_data,.id = "condition") %>%
  select(condition, feature = gene_symbol, value = logFC) %>%
  spread(condition, value)

RNAseq_vars <- RNAseq %>%
  dplyr::select(-feature) %>%
  #drop_na %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh_percentile <- .95
var_thresh <- quantile(RNAseq_vars, probs = var_thresh_percentile, na.rm = TRUE)
RNAseq_hi_vars <- RNAseq[RNAseq_vars > var_thresh,]
  
plot_iheatmap_conditions(RNAseq_hi_vars, assay_name = "RNAseq")

```



### TF_Pathways{.tabset .tabset-fade}

 

```{r TF_pathways_figures}
reverseTFScore <- function(x, max_TF_mean_rank){
  x <- max_TF_mean_rank - x
}

#read in and combine MDD TF scores, reverse ranks
TFs_input <- dir("RNAseq/Data/ChEA3_results_MD_OHSU/",
           pattern = "ctrl_vs_.*xlsx",
           full.names = TRUE) %>% 
  map(readxl::read_excel, col_types = c("text","numeric", "text","numeric", "text", "text")) %>%
  bind_rows() %>%
  as_tibble(.name_repair = "universal") %>%
  mutate(experimentalTime = str_extract(Query.Name, "[24][48]"),
         experimentalTime = as.numeric(experimentalTime),
         ligand = str_remove(Query.Name, "ctrl vs "),
         ligand = str_remove(ligand, ",.*"),
         condition = paste0(ligand, "_", experimentalTime),
         value = 1633 - Rank,
         feature = TF)

  TFs <- TFs_input %>%
    select(condition, value, feature)

#Read in RPPA pathways scores, mean summarise within condition
  RPPA_l1_pathways <- read_csv("RPPA/Data/MDD_Pathways_Score.csv") %>%
  rename(specimenID = X1) %>%
  gather(key = feature, value = Score, -specimenID) %>%
  spread(key = specimenID, value = Score)  %>%
  dplyr::select(-sid64) %>%
    gather(specimenID, value, -feature) %>%
  inner_join(md, by = "specimenID") %>%
   dplyr::select(specimenName, feature, value) %>%
    mutate(condition = str_remove(specimenName, "_C1.*")) %>%
    group_by(condition, feature) %>%
    summarise_at("value", mean) %>%
  ungroup
  
TF_pathways <-  TFs %>%
  bind_rows(RPPA_l1_pathways) %>%
  spread(key = condition, value = value) %>%
  select(matches("_24|_48|feature"), -matches("PBS"))
  
TF_pathways_cor <- TF_pathways %>%
  gather(key = "condition", value = value, -feature) %>%
  spread(key = feature, value = value) %>%
  select(-condition) %>%
  cor(method = "spearman")

RPPA_metadata <- read_csv("RPPA/Metadata/MDD_RPPA_antibodyAnnotations.csv") %>%
  mutate(feature = str_remove(MDD, "-.-.$"))

common_TFs <- intersect(TFs_input$TF, RPPA_metadata$Symbols)

RPPA_l1_conditions <- load_data("RPPA", 1) %>%
  gather(key = specimenID, value = value, -antibody) %>%
  inner_join(md, by = "specimenID") %>%
  rename(feature = antibody) %>%
  dplyr::select(specimenName, feature, value) %>%
  mutate(condition = str_remove(specimenName, "_C1.*")) %>%
  group_by(condition, feature) %>%
  summarise_at("value", mean) %>%
  ungroup %>%
  rename(RPPA_value = value) %>%
  left_join(RPPA_metadata, by = c("feature")) %>%
  filter(Symbols %in% common_TFs,
         !str_detect(feature, "FoxO3a_p|ER-a_p|Stat3_p")) %>%
  select(condition, Symbols, RPPA_value)

RPPA_ChEA3_values <- TFs %>%
  filter(feature %in% common_TFs) %>%
  rename(ChEA3_rank = value) %>%
  inner_join(RPPA_l1_conditions, by= c("feature" = "Symbols", "condition"))

if(FALSE){
  RPPA_metadata <- read_csv("RPPA/Metadata/MDD_RPPA_antibodyAnnotations.csv")
  
  RPPA_genes_HGNC_direct <- RPPA_metadata$Symbols
  
  RPPA_genes <- RPPA_l1$feature %>%
    str_remove("_.*") %>%
    toupper() %>%
    unique
  RPPA_genes_HGNC <- RPPA_genes %>%
    str_replace("C-MYC", "MYC") %>%
    str_replace("ER", "ESR1") %>%
    str_replace("PR", "PGR") %>%
    str_replace("FOXO3A", "FOXO3") %>%
    str_replace("TWIST", "TWIST1") %>%
    str_replace("CREB", "CREB1") %>%
    str_replace("SNAIL", "SNAI1") %>%
    str_replace("HIF-1-ALPHA", "HIF1A") %>%
    str_remove("-")
  
  ChEA3_TFs <- TFs_input$TF %>%
    unique
  ChEA3_TFs[str_detect(ChEA3_TFs, "XBP")]
  RPPA_genes_HGNC_direct[str_detect(RPPA_genes_HGNC_direct, "XBP")]
  common_TFs <- intersect(RPPA_genes_HGNC, ChEA3_TFs)
  common_TFs_direct <- intersect(RPPA_genes_HGNC_direct, ChEA3_TFs)
  set_diff <- setdiff(common_TFs_direct, common_TFs)
  set_diff2 <- setdiff(common_TFs, common_TFs_direct)

}
```


```{r top_TF_pathways_figures, eval = TRUE}

  #Select top TFs of each ligand
  top_TF_names <- TFs %>%
  spread(key = feature, value = value) %>%
    mutate(Ligand = str_remove(condition, "_.*")) %>%
    group_by(Ligand) %>%
    summarise_if(is.numeric, max) %>%
    ungroup %>%
    gather(key = TF, value = Rank, -Ligand) %>%
    group_by(Ligand) %>%
    top_n(25, Rank) %>%
    ungroup() %>%
    spread(key = TF, value = Rank) %>%
    dplyr::select(-Ligand)
  
  top_TFs <- TFs %>%
  spread(key = feature, value = value)  %>%
    dplyr::select(condition, colnames(top_TF_names))
  
  top_TF_pathways <-  top_TFs %>%
    gather(key = feature,value = value, -condition) %>%
  bind_rows(RPPA_l1_pathways) %>%
  spread(key = condition, value = value) %>%
  select(matches("_24|_48|feature"), -matches("PBS"))
  
    
top_TF_pathways_cor <- top_TF_pathways %>%
  gather(key = "condition", value = value, -feature) %>%
  spread(key = feature, value = value) %>%
  select(-condition) %>%
  cor(method = "spearman")

RPPA_l1_pathways_null <- RPPA_l1_pathways %>%
  mutate(value = value[sample(1:nrow(RPPA_l1_pathways), replace = FALSE)])

top_TF_pathways_null <-  top_TFs %>%
  gather(key = feature,value = value, -condition) %>%
  bind_rows(RPPA_l1_pathways_null) %>%
  spread(key = condition, value = value) %>%
  select(matches("_24|_48|feature"), -matches("PBS"))

top_TF_pathways_cor_null <- top_TF_pathways_null %>%
  gather(key = "condition", value = value, -feature) %>%
  spread(key = feature, value = value) %>%
  select(-condition) %>%
  cor(method = "spearman")

```


#### Pathways heatmap 

```{r plotTFsvsPathways}

RPPA_l1_pathways <- read_csv("RPPA/Data/MDD_Pathways_Score.csv") %>%
  rename(specimenID = X1) %>%
  gather(key = feature, value = Score, -specimenID) %>%
  spread(key = specimenID, value = Score)

plot_heatmap(RPPA_l1_pathways, md, EGF_normed = FALSE, assay_name = "RPPA",show_row_names = TRUE,
                       row_names_gp = gpar(fontsize = 10))
```

#### Top TF heatmap

```{r}
df <- top_TF_pathways %>%
  dplyr::select(-matches("_Score")) %>%
  filter(!str_detect(feature,"_Score"))

df_as_matrix <- df %>%
  select(-feature) %>%
  as.matrix() %>%
  t %>%
    scale() %>%
    t
colnames(df_as_matrix) <- str_remove(colnames(df_as_matrix),"_Pathway_Score")
rownames(df_as_matrix) <- df$feature

Heatmap(matrix = df_as_matrix,
          name = "ranks",
          column_title = paste0("Ranks of top TFs"),
          show_row_names = TRUE,
        row_names_gp = gpar(fontsize = 8),
          show_column_names = TRUE,
          cluster_rows = TRUE,
          cluster_columns = FALSE,
          #col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")
```


#### RPPA vs ChEA3 values

```{r}

p <- ggplot(RPPA_ChEA3_values, aes(ChEA3_rank, RPPA_value, colour=condition)) +
  geom_point(alpha = .8) +
  facet_wrap(~feature)
p

p <- ggplot(RPPA_ChEA3_values, aes(ChEA3_rank, RPPA_value, colour=condition)) +
  geom_point(alpha = .8) +
  facet_wrap(~feature, scales = "free_y", ncol = 5)
p
```


#### Correlations heatmap


```{r TFxPathwayCorrelations}

df <- TF_pathways_cor %>%
  as_tibble(rownames="TF") %>%
  dplyr::select(matches("_Score|^TF$")) %>%
  filter(!str_detect(TF,"_Score"))

df_as_matrix <- df %>%
  select(-TF) %>%
  as.matrix()
colnames(df_as_matrix) <- str_remove(colnames(df_as_matrix),"_Pathway_Score")

Heatmap(matrix = df_as_matrix,
          name = "correlations",
          column_title = paste0("Correlation of all TFs and pathways"),
          show_row_names = FALSE,
          show_column_names = TRUE,
          cluster_rows = TRUE,
          cluster_columns = TRUE,
          col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")

```


Method Summary  

Start with DCIC's top ranked TFs from ChEA3 analysis  
Reverse ChEA3 TF rankings so that high values represent highly scored TFs  
Mean summarize replicate RPPA pathway scores  
Combine TF rankings and RPPA pathways scores
Find the correlations between all TFs and pathways across all 24 and 48 hour conditions and display as a heatmap  


#### Top TFs correlations  

```{r topTFxPathwayCorrelations, fig.height=14}

df <- top_TF_pathways_cor %>%
  as_tibble(rownames="TF") %>%
  dplyr::select(matches("_Score|^TF$")) %>%
  filter(!str_detect(TF,"_Score"))

df_as_matrix <- df %>%
  select(-TF) %>%
  as.matrix()
colnames(df_as_matrix) <- str_remove(colnames(df_as_matrix),"_Pathway_Score")
rownames(df_as_matrix) <- df$TF

Heatmap(matrix = df_as_matrix,
          name = "correlations",
          column_title = paste0("Correlation of top TFs and pathways"),
          show_row_names = TRUE,
        row_names_gp = gpar(fontsize = 6),
          show_column_names = TRUE,
          cluster_rows = TRUE,
          cluster_columns = TRUE,
          col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")

df2 <- df_as_matrix %>%
  as_tibble() %>%
  gather(key = "feature")

p <- ggplot(df2, aes(value)) +
  geom_density()+
  facet_wrap(~feature)
p

top_TF_pathways_cor %>%
  as_tibble(rownames="TF") %>%
  dplyr::select(matches("_Score|^TF$")) %>%
  filter(!str_detect(TF,"_Score")) %>%
  gather(key = "Pathway", value = "Correlation", -TF) %>%
  distinct() %>%
  mutate(Correlation = signif(Correlation, 2),
         Pathway = str_remove(Pathway, "_Pathway_Score")) %>%
  DT::datatable()
```

Method Summary  

Start with DCIC's top ranked TFs from ChEA3 analysis  
Reverse ChEA3 TF rankings so that high values represent highly scored TFs  
Group by ligand  
Select the top ranked 25 TFs in both timepoints of each ligand  
Mean summarize replicate RPPA pathway scores  
Combine TF rankings and RPPA pathways scores  
Find the correlations between the top TFs and the RPPA pathways across all 24 and 48 hour  conditions and display as a heatmap  

```{r correlationScatterplots, fig.width= 12, fig.height=20}

top_TF_df <- top_TF_pathways %>%
  gather(key = "condition", value = "TF", -feature) %>%
  filter(!str_detect(feature,"Pathway")) 

pathway <- "EMT_Pathway_Score"
df <- top_TF_pathways %>%
  gather(key = "condition", value = "value", -feature) %>%
  filter(str_detect(feature,pathway)) %>%
  spread(key = feature, value = value) %>%
  right_join(top_TF_df)
  
p <- ggplot(df, aes_string(pathway, "TF")) +
  stat_smooth(aes_string(pathway, "TF"), method = "lm", se = FALSE) +
  geom_point(aes_string(pathway, "TF", colour = "condition")) +
  facet_wrap(~feature)

print(p)

```


#### Top TFs correlation nulls

```{r topTFxPathwayCorrelationNulls, fig.height=14}

df <- top_TF_pathways_cor_null %>%
  as_tibble(rownames="TF") %>%
  dplyr::select(matches("_Score|^TF$")) %>%
  filter(!str_detect(TF,"_Score"))

df_as_matrix <- df %>%
  select(-TF) %>%
  as.matrix()
colnames(df_as_matrix) <- str_remove(colnames(df_as_matrix),"_Pathway_Score")
rownames(df_as_matrix) <- df$TF

Heatmap(matrix = df_as_matrix,
          name = "correlations",
          column_title = paste0("Correlation of top TFs and randomized pathways"),
          show_row_names = TRUE,
        row_names_gp = gpar(fontsize = 8),
          show_column_names = TRUE,
          cluster_rows = TRUE,
          cluster_columns = TRUE,
          col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")


df2 <- df_as_matrix %>%
  t %>%
  as_tibble() %>%
  gather(key = "feature")

p <- ggplot(df2, aes(value)) +
  geom_density()+
  facet_wrap(~feature)
p

```

```{r nullCorrelationScatterplots, fig.width= 12, fig.height=20}

top_TF_df_null <- top_TF_pathways_null %>%
  gather(key = "condition", value = "TF", -feature) %>%
  filter(!str_detect(feature,"Pathway")) 

df_null <- top_TF_pathways_null %>%
  gather(key = "condition", value = "value", -feature) %>%
  filter(str_detect(feature,pathway)) %>%
  spread(key = feature, value = value) %>%
  right_join(top_TF_df_null)
  
p <- ggplot(df_null, aes_string(pathway, "TF", colour = "condition")) +
  geom_point() +
  labs(title = paste("Randomized pathway scores for", pathway)) +
  facet_wrap(~feature)
print(p)

```

##