---
title: "MDD batch analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, cache.lazy = FALSE, message=FALSE, warning = FALSE, fig.height=10, fig.width=10)

suppressMessages(library(tidyverse))

```


```{r read_data_and_metadata}

md <- read_csv("metadata/MDD_sample_annotations.csv")


#Read in RPPA values and merge with metadata
RPPA_values <- read_csv("RPPA/Data/MDD_RPPA_Level3.csv") %>%
  gather(specimenID, value = value,  -antibody) %>%
  left_join(md, by = "specimenID") %>%
  mutate(value = 2^value) %>%
  rename(condition = experimentalCondition,
         time = experimentalTimePoint)
```

Start the batch analysis by comparing RPPA replicates. Each dot in the scatterplots is an RPPA antibody at one condition.  

```{r fig.height=3.5, fig.width=4}

df <- RPPA_values %>%
  filter(replicate %in% c("A","B")) %>%
  dplyr::select(condition, time, ligand, antibody, value, replicate) %>%
  spread(key = replicate, value = value)

p <- ggplot(df, aes(x = A, y = B, color = ligand)) +
  geom_point(alpha = .3) + 
  coord_cartesian(xlim = c(0,5),
                  ylim = c(0,5)) +
  labs(title = "RPPA replicate B vs replicate A")
 
p 

p <- ggplot(df, aes(x = A, y = B, color = time)) +
  geom_point(alpha = .3) + 
  coord_cartesian(xlim = c(0,5),
                  ylim = c(0,5)) +
  labs(title = "RPPA replicate B vs replicate A")
 
p 

p <- ggplot(df, aes(x = A, y = B)) +
  geom_hex(binwidth = .02, show.legend = FALSE) + 
  coord_cartesian(xlim = c(0,5),
                  ylim = c(0,5)) +
  labs(title = "RPPA replicate B vs replicate A",
       subtitle = "colored by density of values")
 
p 
```

```{r fig.height=3.5, fig.width=4}

df <- RPPA_values %>%
  filter(replicate %in% c("B","C")) %>%
  dplyr::select(condition, time, ligand, antibody, value, replicate) %>%
  spread(key = replicate, value = value)

p <- ggplot(df, aes(x = B, y = C, color = ligand)) +
  geom_point(alpha = .3) + 
  coord_cartesian(xlim = c(0,5),
                  ylim = c(0,5)) +
  labs(title = "RPPA replicate C vs replicate B")
 
p 

p <- ggplot(df, aes(x = B, y = C, color = time)) +
  geom_point(alpha = .3) + 
  coord_cartesian(xlim = c(0,5),
                  ylim = c(0,5)) +
  labs(title = "RPPA replicate C vs replicate B")
 
p 

p <- ggplot(df, aes(x = B, y = C)) +
  geom_hex(binwidth = .02, show.legend = FALSE) + 
  coord_cartesian(xlim = c(0,5),
                  ylim = c(0,5)) +
  labs(title = "RPPA replicate C vs replicate B",
       subtitle = "colored by density of values")
 
p 
```


```{r fig.height=3.5, fig.width=4}

df <- RPPA_values %>%
  filter(replicate %in% c("A","C")) %>%
  dplyr::select(condition, time, ligand, antibody, value, replicate) %>%
  spread(key = replicate, value = value)

p <- ggplot(df, aes(x = A, y = C, color = ligand)) +
  geom_point(alpha = .3) + 
  coord_cartesian(xlim = c(0,5),
                  ylim = c(0,5)) +
  labs(title = "RPPA replicate C vs replicate A")
 
p 

p <- ggplot(df, aes(x = A, y = C, color = time)) +
  geom_point(alpha = .3) + 
  coord_cartesian(xlim = c(0,5),
                  ylim = c(0,5)) +
  labs(title = "RPPA replicate C vs replicate A")
 
p 

p <- ggplot(df, aes(x = A, y = C,)) +
  geom_hex(binwidth = .02, show.legend = FALSE) + 
  coord_cartesian(xlim = c(0,5),
                  ylim = c(0,5)) +
  labs(title = "RPPA replicate C vs replicate A",
       subtitle = "colored by density of values")
 
p 
```

Display correlations between `r length(unique(RPPA_values$condition))` conditions in each set of replicates for each RPPA antibody.  

```{r fig.height=3.5, fig.width=4}

df_A_B <- RPPA_values %>%
  dplyr::select(specimenName, value, antibody) %>%
    filter(str_detect(specimenName, "_A|_B")) %>%
  spread(key = specimenName, value = value)

df_mat_A_B <- df_A_B %>%
  dplyr::select_if(is.numeric) %>%
  as.matrix()

df_cor_A_B <- df_mat_A_B %>%
            t %>%
            cor(use = "complete.obs",method = "spearman") %>%
  as.tibble() %>%
  gather(key = "antibody", value = value) %>%
  mutate(correlation = "A_B")

df_A_C <- RPPA_values %>%
  dplyr::select(specimenName, value, antibody) %>%
    filter(str_detect(specimenName, "_A|_C")) %>%
  spread(key = specimenName, value = value)

df_mat_A_C <- df_A_C %>%
  dplyr::select_if(is.numeric) %>%
  as.matrix()

df_cor_A_C <- df_mat_A_C %>%
            t %>%
            cor(use = "complete.obs",method = "spearman") %>%
  as.tibble() %>%
  gather(key = "antibody", value = value) %>%
  mutate(correlation = "A_C")

df_B_C <- RPPA_values %>%
  dplyr::select(specimenName, value, antibody) %>%
    filter(str_detect(specimenName, "_B|_C")) %>%
  spread(key = specimenName, value = value)

df_mat_B_C <- df_B_C %>%
  dplyr::select_if(is.numeric) %>%
  as.matrix()

df_cor_all <- df_mat_B_C %>%
            t %>%
            cor(use = "complete.obs",method = "spearman") %>%
  as.tibble() %>%
  gather(key = "antibody", value = value) %>%
  mutate(correlation = "B_C") %>%
  bind_rows(df_cor_A_B, df_cor_A_C)

p <- ggplot(df_cor_all, aes(x = value, fill = correlation)) +
  geom_density(alpha = .3) 
 
p 

p <- ggplot(df_cor_all, aes(x = value, fill = correlation)) +
  geom_density(alpha = .3) +
  facet_wrap(~correlation)
 
p 
```



```{r RPPApairwiseCorrelations}

  df_A <- RPPA_values %>%
  dplyr::select(specimenName, value, antibody) %>%
    filter(str_detect(specimenName, "_A")) %>%
  spread(key = specimenName, value = value)

df_mat_A <- df_A %>%
  dplyr::select_if(is.numeric) %>%
  as.matrix()
rownames(df_mat_A) <- df_A$antibody

 df_B <- RPPA_values %>%
  dplyr::select(specimenName, value, antibody) %>%
    filter(str_detect(specimenName, "_B")) %>%
  spread(key = specimenName, value = value)

df_mat_B <- df_B %>%
  dplyr::select_if(is.numeric) %>%
  as.matrix()
rownames(df_mat_B) <- df_B$antibody

 df_C <- RPPA_values %>%
  dplyr::select(specimenName, value, antibody) %>%
    filter(str_detect(specimenName, "_C")) %>%
  spread(key = specimenName, value = value)

df_mat_C <- df_C %>%
  dplyr::select_if(is.numeric) %>%
  as.matrix()
rownames(df_mat_C) <- df_C$antibody

alpha_order <- function(x,y){
  foo <- sapply(seq_along(x), function(i) {
    paste0(sort(c(x[i],y[i])),collapse = "_") %>%
      as.vector
})
}

#Within a replciates set, calculate pairwise correlations between 
#each features values

#given a matrix of values with specimenName columns
#and features in the rows, calculate the spearman 
#correlations between the rows
#Create an alphabetically ordered name for each correlation
#Assign a replicate label and filter to a dataframe of unique rows
calcPairwiseCors <- function(mat, replicate = NULL){
  mat %>%
  t %>%
  cor(use = "complete.obs",method = "spearman") %>%
  as.tibble() %>%
  mutate(antibody_2 = rownames(mat)) %>%
  gather(key = "antibody", value = correlation, -antibody_2) %>%
  mutate(replicate = replicate,
         antibody_ordered = alpha_order(antibody, antibody_2)) %>%
    filter(!antibody == antibody_2) %>%
  select(antibody_ordered, correlation, replicate) %>%
  distinct()
}

pairwiseCors <- bind_rows(calcPairwiseCors(df_mat_A, "A"),
                 calcPairwiseCors(df_mat_B, "B"),
                 calcPairwiseCors(df_mat_C, "C"))

#Create a null set of correlations 
#For each feature in replicates set 1, calculate its correlation to replicate set 2

#given a matrix of values with specimenName columns
#and features in the rows, calculate the spearman 
#correlations between the rows
#Create an alphabetically ordered name for each correlation
#Assign a replicate label and filter to a dataframe of unique rows
calcNullPairwiseCors <- function(mat_1, mat_2, replicate = NULL){
  
  #Calculate the correlation between each row in mat_1 to each row in mat_2
  calcNullCors <- function(mat_1_row){
    apply(mat_2, 1, calcCor, m1r = mat_1_row)
  }
  
  calcCor <- function(x, m1r){
    cor(x, m1r)
  }
  foo <- apply(mat_1, 1, calcNullCors)
}

null_cor_mat <- calcNullPairwiseCors(df_mat_A, df_mat_B)

null_cors <- lapply(2:nrow(null_cor_mat), function(x){
  res <- null_cor_mat[x:nrow(null_cor_mat), x-1]
}) %>%
  unlist

df_cor_A <- df_mat_A %>%
  t %>%
  cor(use = "complete.obs",method = "spearman") %>%
  as.tibble() %>%
  mutate(antibody_2 = df_A$antibody) %>%
  gather(key = "antibody", value = correlation, -antibody_2) %>%
  mutate(replicate = "A",
         antibody_ordered = alpha_order(antibody, antibody_2)) %>%
    filter(!antibody == antibody_2) %>%
  select(antibody_ordered, correlation, replicate) %>%
  distinct()

p <- ggplot(df_cor_A, aes(x = correlation)) +
  geom_density() +
  labs(title = "Pairwise correlations between RPPA replicate A antibodies")
#p

p <- ggplot(pairwiseCors, aes(x = correlation, fill = replicate)) +
  geom_density(alpha= .4)
#p

df <- bind_rows(pairwiseCors,
                tibble(correlation = null_cors,
                       replicate = "null",
                       antibody_ordered = names(null_cors)))

p <- ggplot(filter(df, !replicate=="null"),  aes(x = correlation, fill = replicate)) +
  geom_density(alpha = .4)
p <- p + geom_density(data=filter(df, replicate=="null"), aes(x = correlation),fill = "transparent", size = 2) +
    labs(title = "Pairwise correlations within RPPA replicate sets")

p

```

These density plots show the distributions of pairwise correlations between the vector of values from each RPPA antibody within each of the three replicate sets. The heavy line shows the distribition of a null set of correlations created by comparing each feature in replicate set A to the rest of the features in replicate set B.  

