---
title: "MDD integrated data analysis by time"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, cache.lazy = FALSE, message=FALSE, warning = FALSE, fig.height=10, fig.width=12)

suppressMessages(library(tidyverse))
library(iheatmapr)
library(umap)

source("R/MDD_functions.R")
```


```{r read_data_and_metadata}
  
data_path <- "Data/selected_assay_pk_data_rr.rda"
if(!file.exists(data_path)) stop("cannot load ",data_path)
load(data_path)

```

## {.tabset .tabset-fade}

### heatmap

```{r interactiveHeatmaps, fig.height = 8}

df_p <- selected_assay_pk_data_rr %>%
  mutate(experimentalCondition = paste0(ligand,"_",experimentalTimePoint)) %>%
  filter(experimentalTimePoint %in% c(24,48))

df_p_as_matrix <-  prep_hm_matrix(df_p, columns = "time") %>%
  t %>%
  as.matrix() %>%
  rrscale::winsor() %>%
  scale(scale = FALSE) %>%
  t
ann_df <- prep_hm_annotations(df_p, columns = "time")

clust_num <- 5
cluster_method = "kmeans"
#setup for common cluster colors used in heatmap and ggplot
cluster_cols <- RColorBrewer::brewer.pal(clust_num, "Paired")
assay_cols <- RColorBrewer::brewer.pal(length(unique(df_p$Type)), "Paired")
names(assay_cols) <- unique(df_p$Type)
                            
hm <- get_iheatmap(df_p_as_matrix, assay_name = "Selected assay data, rrscaled") %>%
  format_hm(k = clust_num, cluster_method = cluster_method)
hm

hm_data <- tibble(feature = names(hm@plots@listData$`Cluster(kmeans)`@data),
                  Cluster = hm@plots@listData$`Cluster(kmeans)`@data,
                  Type = hm@plots@listData$Type@text)

df_p_sp <- df_p_as_matrix %>%
  data.frame() %>%
  rownames_to_column("feature") %>%
  left_join(hm_data, by = "feature")

```



###


### PCA  

```{r PCAAnalysis}

pca_obj <- prcomp(df_p_as_matrix)
plot(pca_obj)

pca_df <- as_tibble(pca_obj$x) %>%
  mutate(feature = rownames(pca_obj$x),
         Type = str_extract(feature,"_(RNA|RPPA|cycIF|ATAC|GCP|IF)"),
         Type = str_remove_all(Type, "_")) %>%
  left_join(hm_data, by = c("feature", "Type")) %>%
  mutate(feature = str_remove(feature, "_.*"))
  

pca_diagnostics_df <- as_tibble(pca_obj$rotation) %>%
  mutate(condition = rownames(pca_obj$rotation)) 
```

```{r pcaFigures}


p <- ggplot(pca_df, aes(x = PC1, y = PC2, colour = factor(Type))) +
              geom_point(alpha = .5)
p
# 
# p <- ggplot(pca_df, aes(x = PC3, y = PC4, colour = factor(Type))) +
#               geom_point(alpha = .5)
# p

p <- ggplot(pca_df, aes(x = PC1, y = PC2, colour = factor(Cluster))) +
              geom_point(alpha = .5)
p

df <- pca_diagnostics_df %>%
  gather(PC, value, -condition)

p <- ggplot(df, aes(x = PC, y = value, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge")
p

p <- ggplot(df, aes(x = condition, y = value, fill = PC)) +
  geom_bar(stat = "identity", position = "dodge")
p

```

###


### UMAPs

```{r UMAPAnalysis, fig.width=5.75, fig.height=4.75}

#Generate a 2d scatterplot of the UMAP dimensions and color by clusters, assay type

#Create annotation values
hm_data_matrix <- df_p_sp %>%
  select(-feature, -Cluster, -Type) %>%
  as.matrix()
rownames(hm_data_matrix) <- df_p_sp$feature

df_UMAP <- umap(hm_data_matrix, na.rm = TRUE)$layout %>%
  data.frame(feature = rownames(hm_data_matrix)) %>%
  rename(UMAP_1 = X1,
         UMAP_2 = X2) %>%
 left_join(df_p_sp, by = "feature") 

p <- ggplot(df_UMAP, aes(x = UMAP_1,
                         y = UMAP_2,
                         colour = factor(Cluster))) +
  geom_point(size = .8, alpha = .8) +
  scale_color_manual(values = cluster_cols) +
  labs(title = paste("UMAP embedding colored by cluster"),
       colour = "Cluster") +
    guides(colour = guide_legend(override.aes = list(alpha = 1, size = 4))) +
  theme_bw() +
  theme(axis.title = element_blank())
p
```


```{r UMAPAnalysis2, fig.width=6.5, fig.height=4.5}
p <- ggplot(df_UMAP, aes(x = UMAP_1,
                         y = UMAP_2,
                         colour = factor(Type))) +
  geom_point(size = .8, alpha = .8) +
  scale_colour_brewer(palette = "Paired")+
  labs(title = paste("UMAP embedding colored by assay"),
       colour = "Assay") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 4))) +
  theme_bw() +
  theme(axis.title = element_blank())
p

```

```{r highlightAssay, fig.width=2.5, fig.height=2.5}

for(assay in unique(df_p$Type)){
  regex_pat <- paste0("^",assay,"$")
  df <- df_UMAP %>%
    mutate(Highlight = str_extract(Type, regex_pat),
           Highlight = str_replace_na(Highlight),
           Highlight = str_replace(Highlight, "^NA$", ".1"),
           Highlight = str_replace(Highlight, assay, "3"),
           Highlight = as.numeric(Highlight))
  p <- ggplot(df, aes(x = UMAP_1,
                      y = UMAP_2,
                      colour = factor(Highlight),
                      size = Highlight)) +
    geom_point(alpha = .6) +
    annotate("text",x = -4, y = -9, label = assay)+
    scale_color_manual(values=c("grey", "slateblue1"))+
    scale_size_area(limits = c(0, 30)) +
    labs()+
    guides(colour = FALSE,
           size = FALSE)+
    theme_bw()+
    theme(text = element_text(size = 1),
          axis.title = element_blank())
  print(p)
}

```

```{r UMAPbyCondition, fig.width=2.5, fig.height=2.5}

df_UMAP_times <- colnames(df_UMAP)[str_detect(colnames(df_UMAP), "T4|T8|T24|T24")]
for(expTime in df_UMAP_times){
  p <- ggplot(df_UMAP, aes_string(x = "UMAP_1",
                                  y = "UMAP_2",
                                  colour = expTime)) +
    geom_point(size = .4, alpha = .6) +
    annotate("text",x = -4, y = -9, label = expTime)+
    scale_color_gradient2(low = "blue4", mid = "azure",
                          high = "red", midpoint = 0,) +
    labs()+
    guides(colour = FALSE)+
    theme_bw()+
    theme(text = element_text(size = 1),
          axis.title = element_blank())
  print(p)
}

```

###

### method  

This dataset combines MCF10A Molecular Deep Dive (MDD) RNAseq, Reverse Phase Protein Array(RPPA), Global Chromatin Profiling (GCP) histone modifications, immunofluorescence (IF), cyclic immunofluorescence (cycIF) and ATACseq chromatin configuration data. All data at all timepoints is used. Most of the data is median summarized from three replicate experiments run in sequential weeks. It is then rrscale transformed and then zscored by feature.   

RRscale  
Use rrscale https://CRAN.R-project.org/package=rrscale to transform features in the heatmap before clustering.  This method performs a box-cox gaussianizing transformation on each feature, zscores the results then eliminates outliers with absolute values greater than `r zscore_cutoff`. All values from each assay was rrscaled as a matrix with features in the columns and conditions in the rows. Because of the differences in scales in the morphology, spatial and statistical values in the IF dataset, only the intensity values are included.      

This analysis is based on the LINCS Molecular Deep Dive data files on Synapse at  https://www.synapse.org/#!Synapse:syn2862345/wiki/588244.  


ATACseq Values  
Median summarize the values across the replicates  
Using a differential analysis between each ligand and ctrl, select a fixed number of high variance gened per ligand.

cycIF Values  
Select `r length(unique(cycIF_values_rr$feature))` cycIF features based on their biological relevance  
Median summarize the values across the replicates  

GCP Values  
Median summarize the values across the replicates  

IF Values  
Filter out the CellMask intensities and use the remaining  intensity IF features  
Median summarize the values across the replicates  

RNAseq Values  
A gene list representative of the largest gene expression changes across all ligand treatments was produced from differential expression analyses of each experimental condition compared to the control condition. For each experimental condition, a differential expression analysis was performed using the R package DESeq2 (1.24.0), with shrunken log2 fold change estimates calculated using the apeglm method. Each list was filtered to include only genes with an adjusted p value < 0.05, then ranked by the absolute value of their log2 fold change estimate. The 50 most differentially expressed genes per condition were selected. Redundant genes were removed.

RPPA Values  
Median summarize the values across the replicates  

Clustering  
kmeans clustering is used with k = `r clust_num` based on gap statistic analysis. The scatter plot above the heatmap shows the mean value for each cluster within each condition.  


```{r featureTypes, fig.width=8, fig.height=4}
#create a bar chart of the feature types
df <- df_p %>%
  drop_na() %>%
  select(feature, Type, experimentalTimePoint) %>%
  distinct()
p <- ggplot(df, aes(x = Type, fill = Type)) +
  geom_bar() +
  labs(title = paste("Type distribution of the",length(unique(paste0(df_p$feature, df_p$experimentalTimePoint))), "features in the MDD integrated dataset"),
       x ="Feature type") +
  theme(axis.text = element_text(angle = 90))
p

```

###

### cluster analysis

```{r gapStatistics, fig.width=4, fig.height=4, eval = TRUE}
#implement gap statistic analysis

ligand_order <-  c("PBS", "HGF", "OSM", "EGF","BMP2", "IFNG", "TGFB")
ligand_order <- ligand_order[ligand_order %in% selected_assay_pk_data_rr$ligand]

df_as_matrix <- df_p %>% 
  filter(ligand %in% ligand_order) %>%
spread(ligand, value) %>%
  drop_na() %>%
  gather(ligand, value, -Type, -feature, -experimentalTimePoint) %>%
  prep_hm_matrix(columns = "ligand")
if(file.exists(paste0("R/gss_",cluster_method,"_assay_data.rda"))){
  load(paste0("R/gss_",cluster_method,"_assay_data.rda"))
} else {
  gss <- clusGap(df_as_matrix, FUN = hafun, K.max = 20, B = 100, verbose = FALSE)
  save(gss, file = paste0("R/gss_",cluster_method,"_assay_data.rda"))
}

plot_gap(gss)

p <- ggplot(hm_data, aes(x = as.factor(Cluster), fill = Type)) +
  geom_bar() +
  labs(x = "Cluster",
       fill = "Assay")
p
```

```{r assayClusterdistribution, fig.width=8, fig.height=4, eval = TRUE}
p <- ggplot(hm_data, aes(x = as.factor(Cluster), fill = Type)) +
  geom_bar(position = "dodge") +
  labs(x = "Cluster",
       fill = "Assay") +
  facet_wrap(~Type, scales = "free_y")
p
```


This analysis can be interpreted to select the number of clusters in the dataset. It is based on a method described in Modern Statistics for Modern Biology by Susan Holmes and Wolfgang Huber http://web.stanford.edu/class/bios221/book/Chap-Clustering.html and is excerpted below.  

Taking the logarithm of the within-sum-of-squares (log(WSSk)) and comparing it to averages from simulated data with less structure can be a good way of choosing k. This is the basic idea of the gap statistic introduced by Tibshirani, Walther, and Hastie (2001). We compute log(WSSk) for a range of values of k, the number of clusters, and compare it to that obtained on reference data of similar dimensions with various possible ‘non-clustered’ distributions. We can use uniformly distributed data as we did above or data simulated with the same covariance structure as our original data.

The default choice for the number of clusters, k1, is the first value of k for which the gap is not larger than the first local maximum minus a standard error (see the manual page of the clusGap function). ... the choice recommended by Tibshirani, Walther, and Hastie (2001) is the smallest k such that gap(k)≥gap(k+1)−s′k+1...

The first local maximum is k = 3 which does not yield informative clusters. The second local maximum is k = `r clust_num-1`. This analysis is choosing to use k = `r clust_num` which is is the first value where the gap value minus its standard error is less than this local maximum at `r clust_num-1`.  


<bksp>

<bksp>

###

