---
title: "Upset analysis of the MDD RPPA_24 dataset"
output: html_document
---

```{r setup, include=FALSE}
####ToDOos
#change distance calculation in clustering
#do cluster analysis to set cluster number
#fix Type colors and use in bar chart

knitr::opts_chunk$set(echo = FALSE, cache=FALSE, cache.lazy = FALSE, message=FALSE, warning = FALSE, fig.height=6, fig.width=12)

suppressMessages(library(tidyverse))
library(UpSetR)
library(DT)
```


```{r read_data_and_metadata}

load("Data/selected_assay_pk_data.rda")

df_sp <- RPPA_selected %>%
  mutate(experimentalCondition = paste0(ligand, "_", experimentalTimePoint)) %>%
  select(experimentalCondition, feature, Type, value) %>%
  spread(experimentalCondition, value) 

cor_mat <- df_sp %>%
  select_if(is.numeric) %>%
            as.matrix() %>%
            t %>%
            cor(use = "complete.obs", method = "spearman")

rownames(cor_mat) <- df_sp$feature
colnames(cor_mat) <- df_sp$feature

cor_df_all <- cor_mat %>%
  as_tibble() %>%
  mutate(feature = df_sp$feature) %>%
  select(feature, everything()) %>%
  gather(feature2, value, -feature)

```


<bksp>  


## {.tabset .tabset-fade}


### RPPA_24 Values{.tabset .tabset-fade}

#### Up/Down


```{r RPPA_24HligandSets_updown}

zscore_thresh <- 0.25
cor_thresh <- 0.7

# define members of ligand sets as having any reading above a threshold
#then do upset analysis on those ligand sets
set_membership <- function(x, zscore_thresh){
  res <- numeric(length(x))
  res[x > zscore_thresh] <- 1
  res[x < -zscore_thresh] <- -1
  return(res)
}

df <- RPPA_selected %>%
  mutate(experimentalCondition = paste0(ligand, "_", experimentalTimePoint)) %>%
  select(experimentalCondition, feature, Type, value) %>%
  filter(str_detect(experimentalCondition, "_24")) %>%
  group_by(feature) %>%
  mutate(value = value/median(value)) %>%
  ungroup() %>%
  mutate(value = log2(value),
         direction = set_membership(value, zscore_thresh = zscore_thresh),
         direction = as.character(direction),
         direction = str_replace(direction, "-1", "down"),
         direction = str_replace(direction, "1", "up"),
         direction = factor(direction, ordered = TRUE, levels = c("up","down")),
         experimentalCondition = factor(experimentalCondition, levels = c("PBS_24",  "TGFB_24", "IFNG_24", "OSM_24",  "EGF_24","HGF_24", "BMP2_24"), ordered = TRUE)) %>%
  filter(!direction == 0) %>%
  arrange(direction)

p <- ggplot(df, aes(x = experimentalCondition, fill = factor(direction))) +
  scale_fill_manual(values = c("firebrick4","dodgerblue4")) +
  geom_bar() +
  labs(title = "PBS response dominates the RPPA 24 hour feature sets",
       fill = "Direction",
       x = "Condition",
       y = ("Number of features")) +
  theme_bw()
p

pdf(file = "RPPA_24_set_bargraph.pdf", width = 6, height = 4)
p
dev.off()

df_mc <- RPPA_selected %>%
  mutate(experimentalCondition = paste0(ligand, "_", experimentalTimePoint)) %>%
  select(experimentalCondition, feature, Type, value) %>%
  filter(str_detect(experimentalCondition, "_24")) %>%
  group_by(feature) %>%
  mutate(value = value/median(value)) %>%
  ungroup() %>%
  mutate(value = log2(value),
         direction = set_membership(value, zscore_thresh = zscore_thresh),
         direction = abs(direction)) %>%
  select(-value) %>%
  spread(experimentalCondition, direction, fill = 0)  %>%
  as.data.frame()

upset(df_mc, nsets = 25, number.angles = 30, point.size = 3.5, line.size = 2, nintersects = 40, order = "freq",
      mainbar.y.label = "Set Intersections", sets.x.label = "Total Features Per Condition", 
      text.scale = c(1.3, 1.3, 1, 1, 1, 1),
      mb.ratio = c(0.5, 0.5))

#datatable(df)

df_mc <- RPPA_selected %>%
  mutate(experimentalCondition = paste0(ligand, "_", experimentalTimePoint)) %>%
  select(experimentalCondition, feature, Type, value) %>%
  filter(str_detect(experimentalCondition, "_24")) %>%
  group_by(feature) %>%
  mutate(value = value/median(value)) %>%
  ungroup() %>%
  mutate(value = log2(value),
         direction = set_membership(value, zscore_thresh = zscore_thresh),
         direction = abs(direction))
datatable(df_mc)
```


Features are a member of a condition set in the RPPA 24 hour data if their absolute value is greater than `r zscore_thresh`.  

#### Up


```{r RPPA_24HligandSets_up}

# define members of ligad sets as having any reading above a threshold
#then do upset analysis on those ligand sets
set_membership_up <- function(x, zscore_thresh = 1.5){
  res <- numeric(length(x))
  res[x > zscore_thresh] <- 1
  return(res)
}

df_mc <- RPPA_selected %>%
      mutate(experimentalCondition = paste0(ligand, "_", experimentalTimePoint)) %>%
  select(experimentalCondition, feature, Type, value) %>%
  filter(str_detect(experimentalCondition, "_24")) %>%
   group_by(feature) %>%
  mutate(value = value/median(value)) %>%
  ungroup() %>%
  mutate(value = log2(value),
         direction = set_membership_up(value, zscore_thresh = zscore_thresh)) %>%
    select(-value) %>%
  spread(experimentalCondition, direction, fill = 0)  %>%
  as.data.frame()
  
upset(df_mc, nsets = 25, number.angles = 30, point.size = 3.5, line.size = 2, nintersects = 40, order = "freq",
    mainbar.y.label = "Condition sets Intersections", sets.x.label = "Total Features Per Condition", 
    text.scale = c(1.3, 1.3, 1, 1, 1, 1),
    mb.ratio = c(0.5, 0.5))

pdf("RPPA_24_upset_up.pdf")
upset(df_mc, nsets = 25, number.angles = 30, point.size = 3.5, line.size = 2, nintersects = 40, order = "freq",
    mainbar.y.label = "Condition sets Intersections", sets.x.label = "Total Features Per Condition", 
    text.scale = c(1.3, 1.3, 1, 1, 1, 1),
    mb.ratio = c(0.5, 0.5))
dev.off()
df_mc <- RPPA_selected %>%
      mutate(experimentalCondition = paste0(ligand, "_", experimentalTimePoint)) %>%
  select(experimentalCondition, feature, Type, value) %>%
  filter(str_detect(experimentalCondition, "_24")) %>%
  group_by(feature) %>%
  mutate(value = value/median(value)) %>%
  ungroup() %>%
  mutate(value = log2(value),
         direction = set_membership_up(value, zscore_thresh = zscore_thresh))
datatable(df_mc)
```

Features are a member of a condition set in the RPPA 24 hour data if the value is greater than `r zscore_thresh`.  

#### Down

```{r RPPA_24HligandSets_down}

# define members of ligad sets as having any reading above a threshold
#then do upset analysis on those ligand sets
set_membership_down <- function(x, zscore_thresh = 1.5){
  res <- numeric(length(x))
  res[x < -zscore_thresh] <- 1
  return(res)
}
df_mc <- RPPA_selected %>%
        mutate(experimentalCondition = paste0(ligand, "_", experimentalTimePoint)) %>%
  select(experimentalCondition, feature, Type, value) %>%
   filter(str_detect(experimentalCondition, "_24")) %>%
   group_by(feature) %>%
  mutate(value = value/median(value)) %>%
  ungroup() %>%
  mutate(value = log2(value),
         direction = set_membership_down(value, zscore_thresh = zscore_thresh)) %>%  
  select(-value) %>%
  spread(experimentalCondition, direction, fill = 0)  %>%
  as.data.frame()
  
upset(df_mc, nsets = 25, number.angles = 30, point.size = 3.5, line.size = 2, nintersects = 40, order = "freq",
    mainbar.y.label = "Condition sets Intersections", sets.x.label = "Total Features Per Condition", 
    text.scale = c(1.3, 1.3, 1, 1, 1, 1),
    mb.ratio = c(0.5, 0.5))

pdf("RPPA_24_upset_down.pdf")
upset(df_mc, nsets = 25, number.angles = 30, point.size = 3.5, line.size = 2, nintersects = 40, order = "freq",
    mainbar.y.label = "Condition sets Intersections", sets.x.label = "Total Features Per Condition", 
    text.scale = c(1.3, 1.3, 1, 1, 1, 1),
    mb.ratio = c(0.5, 0.5))
dev.off()

df_mc <- RPPA_selected %>%
        mutate(experimentalCondition = paste0(ligand, "_", experimentalTimePoint)) %>%
  select(experimentalCondition, feature, Type, value) %>%
  filter(str_detect(experimentalCondition, "_24")) %>%
  group_by(feature) %>%
  mutate(value = value/median(value)) %>%
  ungroup() %>%
  mutate(value = log2(value),
         direction = set_membership_down(value, zscore_thresh = zscore_thresh))
datatable(df_mc)

```


Features are a member of a condition set in the RPPA 24 hour data if the value is less than -`r zscore_thresh`.  


###

### Correlations{.tabset .tabset-fade}


#### Correlation Table

```{r}
df <- cor_df_all %>%
  filter(value > cor_thresh) %>%
  mutate(value = signif(value, digits = 3))
datatable(df)

```

####

###

### Method  

This analysis is based on the LINCS Molecular Deep Dive data files on Synapse at  https://www.synapse.org/#!Synapse:syn2862345/wiki/588244.  



###

##