---
title: "MDD standard figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE,  message=FALSE, warning = FALSE, fig.height=10, fig.width=10)

suppressMessages(library(tidyverse))
suppressMessages(library(ComplexHeatmap))
suppressMessages(library(circlize))
library(umap)
library(pheatmap)
library(circlize)
#library(synapser)

```


## Overview{.tabset .tabset-fade}
These figures are derived from data downloaded from the LINCS Molecular Deep Dive files on Synapse https://www.synapse.org/#!Synapse:syn2862345/wiki/588244.  

The data have been filtered for QC issues and normalized to the EGF timepoints.  

```{r circosFunctions}

parseCorMatrix <- function(x){
  alpha_factor <- 255
  from_condition <- x$Condition
  to_condition <- names(x)[c(-1,-22,-23)]
  #remove metadata from correlations
  cors <- x[c(-1,-22,-23)] %>%
    unlist()
  #remove correlations within the ligand conditions
  cors[str_detect(names(cors), x[["Ligand"]])] <- 0
  #Remove negative correlations
  cors[cors<0] <- 0
  #scale alpha based on correlations
  alpha <- (cors * alpha_factor) %% 256
  alpha <- as.integer(alpha) %>%
    as.character.hexmode()
  col <- paste0(col_ligands[x$Ligand],alpha)
  df <- tibble(From_Condition = x$Condition,
               To_Condition =  names(x)[c(-1,-22,-23)],
               From_sector = str_remove(from_condition, "_.*"),
               To_sector = str_remove(to_condition, "_.*"),
               From_timepoint = str_remove(from_condition, ".*_"),
               To_timepoint = str_remove(to_condition, ".*_"),
               Col = col,
               Cors = as.numeric(cors)) %>%
    mutate(From_timepoint = str_replace(From_timepoint,"^4$",".5"),
           From_timepoint = str_replace(From_timepoint,"^8$","1.5"),
           From_timepoint = str_replace(From_timepoint,"^24$","2.5"),
           From_timepoint = str_replace(From_timepoint,"^48$","3.5"),
           From_timepoint = as.numeric(From_timepoint),
           To_timepoint = str_replace(To_timepoint,"^4$",".5"),
           To_timepoint = str_replace(To_timepoint,"^8$","1.5"),
           To_timepoint = str_replace(To_timepoint,"^24$","2.5"),
           To_timepoint = str_replace(To_timepoint,"^48$","3.5"),
           To_timepoint = as.numeric(To_timepoint),
           From_timepoint_st = From_timepoint - .45,
           From_timepoint_end = From_timepoint + .45,
           To_timepoint_st = To_timepoint - .45,
           To_timepoint_end = To_timepoint + .45)
}

circlize_plot <- function() {
  circos.initializeWithIdeogram(plotType = NULL)
  
  bed = generateRandomBed(nr = 300)
  bed = generateRandomBed(nr = 300, nc = 2)
  circos.genomicTrackPlotRegion(bed,
                                panel.fun = function(region, value, ...) {
                                  circos.genomicPoints(region, value, cex = 0.5, pch = 16, col = 2:3, ...)
                                })
  
  bed = generateRandomBed(nr = 500, nc = 2)
  circos.genomicTrackPlotRegion(bed,
                                panel.fun = function(region, value, ...) {
                                  circos.genomicLines(region, value, col = 4:5, ...)
                                })
  
  bed1 = generateRandomBed(nr = 100)
  bed1 = bed1[sample(nrow(bed1), 20), ]
  bed2 = generateRandomBed(nr = 100)
  bed2 = bed2[sample(nrow(bed2), 20), ]
  
  circos.genomicLink(bed1, bed2, col = col_fun(bed1[[4]]))
  
  circos.clear()
}
addLink <- function(x){
  cor_thresh <- .45
  if(x["Cors"] > cor_thresh) {
    lwd <- max(0, as.numeric(x["Cors"])-cor_thresh)*30
    #cat("lwd ",lwd, "col ", x["Col"])
    circos.link(x["From_sector"], as.numeric(x["From_timepoint"]), x["To_sector"], as.numeric(x["To_timepoint"]), h = 0.4, lwd = lwd, col =x["Col"])
  }
}

prepare_data <- function(x){
  df <- x %>%
    data.frame() %>%
    rownames_to_column(var ="Condition") %>%
    mutate(Time = str_remove(Condition, ".*_"),
           Time = str_replace_all(Time, "^4$","1"),
           Time = str_replace_all(Time, "^8$","2"),
           Time = str_replace_all(Time, "^24$","3"),
           Time = str_replace_all(Time, "^48$","4"),
           Ligand = str_remove(Condition, "_.*")
           ) %>%
    rename(Replicate = replicate)
}

draw_one_track <- function(title_text){
  circos.par("track.height" = 0.2,
             "points.overflow.warning" = FALSE)
  circos.initialize(factors = as.factor(df$Ligand), xlim =c(0,4), sector.width = 20)

  circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
    xlim = CELL_META$xlim
    ylim = CELL_META$ylim
    breaks = seq(xlim[1], xlim[2], by = 1)
    n_breaks = length(breaks)
    circos.rect(breaks[-n_breaks], rep(ylim[1], n_breaks - 1),
                breaks[-1], rep(ylim[2], n_breaks - 1),
                col = track_col_names,
                border = NA)
    circos.text(CELL_META$xcenter,
                CELL_META$cell.ylim[2] + uy(3, "mm"),
                CELL_META$sector.index,
                col = c("black","red","black", "red", "red"))
  }) 
  
  title(title_text,
        outer = TRUE,
        cex.main = 1,
        line = -1)
  
  legend(-1.1,
         -.9,
         "TGFB, BMP2 and IFNG\n are paired with EGF",
         bty = "n",
         cex = .6)
}

```

```{r standardFigureFunctions}

#' Read an assay's csv or tsv data file 
load_data <- function(assay_name){
  file_name <- dir("Data",pattern = paste0("MDD_", assay_name,"_","Level4"),recursive = TRUE, full.names = TRUE)
  if(!length(file_name) == 1) stop("There was not one ", assay_name, " data file found")
    
  if(str_detect(file_name, "csv")) {
    df <- read_csv(file_name)
  } else if (str_detect(file_name, "tsv")){
    df <- read_tsv(file_name)
  } else stop("The data file ",file_name, " must be a csv or tsv file type")
} #end of function

get_win_probs <- function(df, low = .02, hi = .98){
  windsorize_probs <- df %>%
    select(-feature) %>%
    unlist %>%
    quantile(probs = c(low, hi), na.rm = TRUE)
}
  
prep_data <- function(df, md){
    df <- df %>%
    gather(specimenID, value, -feature) %>%
    inner_join(md, by = "specimenID") %>%
    select(feature, specimenName, value) %>%
    spread(specimenName, value)
  #Set the order of the columns
  condition_order <-  paste(rep(c("BMP2", "IFNG", "TGFB","PBS", "HGF", "OSM", "EGF"), each = 25, times = 3), rep(c(1, 4 ,8, 24, 48), each = 5, times = 3), rep(c("C1", "C2", "C3"), each = 175), rep(c("A", "B", "C", "D", "E"), times = 105),  sep = "_") %>%
    intersect(colnames(df))
  
  #Arrange columns by standard condition order
  df <- df %>%
    select(feature, condition_order)
  return(df)
}

prep_annotations <- function(df, md){
      df_ann <- df %>%
      gather("specimenName", "value", -feature) %>%
      inner_join(md, by = "specimenName") %>%
      select(ligand, secondLigand, experimentalTimePoint, replicate, collection) %>%
      distinct() %>%
      rename(Time = experimentalTimePoint,
             Ligand = ligand,
             Ligand2 = secondLigand,
             Replicate = replicate,
             Collection = collection) %>%
      as.data.frame()
      return(df_ann)
}
 #Create the heatmap column annotations
create_top_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md)
  
  haTop <- HeatmapAnnotation(df = df_ann,
                             col = list(Ligand = c( "BMP2" = "#ffffb3",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#b3de69",
                                                    "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072"),
                                        Time = c("1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = "#a6cee3","B" = "#1f78b4", "C" = "#b2df8a", "D" = "#33a02c", "E" = 
"#fb9a99"),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haTop)
}

create_row_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md) %>%
    select(Collection, Replicate, Time, Ligand2, Ligand)
  ####ha = rowAnnotation(foo = anno_mark(at = c(1:4, 20, 60, 97:100), labels = month.name[1:10])
  ###
  haRow <- rowAnnotation(df = df_ann, show_legend = FALSE,
                             col = list(Ligand = c("BMP2" = "#ffffb3",
                                                   "IFNG" = "#bebada",
                                                   "TGFB" = "#b3de69",
                                                   "PBS" = "#8dd3c7",
                                                   "HGF" = "#80b1d3",
                                                   "OSM" = "#fdb462",
                                                   "EGF" = "#fb8072"),
                                        Time = c("1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = "#a6cee3","B" = "#1f78b4", "C" = "#b2df8a", "D" = "#33a02c", "E" = "#fb9a99"),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haRow)
}

  
#####
plot_heatmap <- function(df, md, assay_name, show_row_names = FALSE, mark_rows = NULL){
  
  windsorize_probs <- get_win_probs(df, .02, .98)
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md) %>%
    select(-matches("EGF"))
  
  if(!is.null(mark_rows)) {
    row_nums <- which(df$feature %in% mark_rows)
    ra <- rowAnnotation(foo = anno_mark(at = c(row_nums), labels = df$feature[row_nums]))
  } else {
    ra <- NULL
  }
  
   #Create the heatmap
  Heatmap(matrix = select(df, -feature) %>%
            as.matrix(),
          name = "abundance",
          column_title = paste(assay_name, " EGF timecourse normalized"),
          top_annotation = create_top_annotations(df, md),
          right_annotation = ra,
          show_row_names = show_row_names,
          show_column_names = FALSE,
          cluster_columns = FALSE,
          col = colorRamp2(c(windsorize_probs[1], 0, windsorize_probs[2]), c("blue", "white", "red")),
          na_col = "grey")
}

plot_umap <- function(df, md, assay_name,
                      ligand_cols = c(
                   "BMP2" = "#ffffb3",
                   "IFNG" = "#bebada",
                   "TGFB" = "#b3de69",
                   "PBS" = "#8dd3c7",
                   "HGF" = "#80b1d3",
                   "OSM" = "#fdb462",
                   "EGF" = "#fb8072"),
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #Generate a 2d scatterplot of the UMAP dimensions
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md) %>%
    select(-matches("EGF"))
  
   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature

  df_UMAP <- umap(df_mat, na.rm = TRUE)$layout %>%
  data.frame() %>%
  rename(UMAP_1 = X1,
         UMAP_2 = X2) %>%
    cbind(df_ann)
  
p <- ggplot(df_UMAP, aes(x = UMAP_1,
                         y = UMAP_2,
                         size = Time,
                         colour = Ligand,
                         fill = Ligand2)) +
    geom_point(alpha = .8) +
  geom_point(shape = 21, stroke = 4, alpha = .8) +
  scale_color_manual( values = ligand_cols) +
  scale_fill_manual( values = ligand_2_cols) +
  scale_size(breaks = c(1,4,8,24,48)) +
  labs(title = paste("UMAP embedding of ", assay_name, " data")) +
  theme_bw()
p
  
}

plot_pca <- function(df, md, assay_name,
                      ligand_cols = c(
                   "BMP2" = "#ffffb3",
                   "IFNG" = "#bebada",
                   "TGFB" = "#b3de69",
                   "PBS" = "#8dd3c7",
                   "HGF" = "#80b1d3",
                   "OSM" = "#fdb462",
                   "EGF" = "#fb8072"),
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #Generate a 2d scatterplot of the first two princpal components
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md) %>%
    select(-matches("EGF"))
  
   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature

    df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann)
    
  df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann)
  
p <- ggplot(df_pca, aes(x = PC1,
                         y = PC2,
                         size = Time,
                         colour = Ligand,
                         fill = Ligand2)) +
  geom_point(alpha = .8) +
  geom_point(shape = 21, stroke = 4, alpha = .8) +
  scale_color_manual( values = ligand_cols) +
  scale_fill_manual( values = ligand_2_cols) +
  scale_size(breaks = c(1,4,8,24,48)) +
  labs(title = paste("PCA of ", assay_name, " data")) + 
  theme_bw()
p
}


plot_introspect <- function(df, md, assay_name,
                      ligand_cols = c(
                   "BMP2" = "#ffffb3",
                   "IFNG" = "#bebada",
                   "TGFB" = "#b3de69",
                   "PBS" = "#8dd3c7",
                   "HGF" = "#80b1d3",
                   "OSM" = "#fdb462",
                   "EGF" = "#fb8072"),
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #calculate correlations across the conditions and show in a heatmap

  windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimanName, reorder delete EGF samples
  df <- prep_data(df, md) %>%
    select(-matches("EGF"))
  
  #Create annotation values
  #df_ann <-prep_annotations(df, md)
  
  haRow <- create_row_annotations(df, md)
  
  #Create the heatmap
  Heatmap(matrix = select(df, -feature) %>%
            as.matrix() %>%
            cor(use = "complete.obs"),
          name = "abundance",
          column_title = paste("Introspect of",assay_name, "EGF timecourse normalized"),
          top_annotation = create_top_annotations(df, md),
          left_annotation = haRow,
          show_row_names = FALSE,
          show_column_names = FALSE,
          cluster_rows = FALSE,
          cluster_columns = FALSE,
          col = colorRamp2(c(windsorize_probs[1], 0, windsorize_probs[2]), c("blue", "white", "red")),
          na_col = "grey")
}

# plot_circos <- function(df){
#  ####DEBUG refactor to match new data sturcture###### 
# #Assign colors to the ligands 
# col_ligands <- c("PBS" = "#8dd3c7", "BMP2" = "#ffffb3", "IFNG" = "#bebada", "TGFB" = "#b3de69", "HGF" = "#80b1d3", "OSM" = "#fdb462", "EGF" = "#fb8072", "ctrl" = "#ffffff66")
# #names(col_ligands) <- unique(rppa$Ligand)
# 
# #Assign colors to conditions in the track
# track_col_names <- c("1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black") 
# 
# res <- draw_one_track(title_text = "RPPA Correlations")
# df <- rppa
# for(i in 1:nrow(df)){
#   foo <- parseCorMatrix(df[i,])
#   res <- apply(foo, 1, addLink)
# }
# circos.clear()
# 
# }
# 
# load("Data/MCF10A_RPPA_GCP_CYCIF_condition_correlation.Rdata")
# 
# rppa <- prepare_data(cormats[["RPPA"]])
# #cycIF <- prepare_data(cormats[["CYCIF"]])
# #GCP <- prepare_data(cormats[["GCP"]])
# 
# #Assign ggplot colors to the ligands based on their order in the RPPA dataset
# col_ligands <- gg_color_hue(5)
# names(col_ligands) <- unique(rppa$Ligand)
# #Assign colors to conditions in the track
# track_col_names <- c("grey50", "grey25","red3","royalblue3")
# 
# pdf("CircosPlots.pdf",width = 7, height = 7)
# for(ligand in unique(rppa$Ligand)){
#   res <- draw_one_track(title_text = "RPPA Correlations")
#   df <- rppa %>%
#     filter(Ligand == ligand)
#   for(i in 1:nrow(df)){
#     foo <- parseCorMatrix(df[i,])
#     res <- apply(foo, 1, addLink)
#   }
#   circos.clear()
# }
# res <- draw_one_track(title_text = "RPPA Correlations")
# df <- rppa
# for(i in 1:nrow(df)){
#   foo <- parseCorMatrix(df[i,])
#   res <- apply(foo, 1, addLink)
# }
# circos.clear()

# for(ligand in unique(cycIF$Ligand)){
#   res <- draw_one_track(title_text = "cycIF Correlations")
#   df <- rppa %>%
#     filter(Ligand == ligand)
#   for(i in 1:nrow(df)){
#     foo <- parseCorMatrix(df[i,])
#     res <- apply(foo, 1, addLink)
#   }
#   circos.clear()
# }
# res <- draw_one_track(title_text = "cycIF Correlations")
# df <- cycIF
# for(i in 1:nrow(df)){
#   foo <- parseCorMatrix(df[i,])
#   res <- apply(foo, 1, addLink)
# }
# circos.clear()
# 
# res <- draw_one_track(title_text = "GCP Correlations")
# df <- GCP
# for(i in 1:nrow(df)){
#   foo <- parseCorMatrix(df[i,])
#   res <- apply(foo, 1, addLink)
# }
# circos.clear()

##########
```

### RPPA{.tabset .tabset-fade}

#### Heatmap

```{r rppa_figures, eval = TRUE}

md <- read_csv("metadata/MDD_sample_annotations.csv")

RPPA <- load_data("RPPA") %>%
  rename(feature = antibody)
```


```{r}

plot_heatmap(RPPA, md, assay_name = "RPPA", mark_rows = c("EGFR", "Smad3","HER2", "HER3", "Stat3", "Claudin-7", "YAP", "Notch1"))

```

All `r nrow(df)` proteins measured in the RPPA assay are shown in the heatmap.  

#### Line graphs

```{r, fig.height=15, fig.width=15}



plot_line_graphs <- function(df, md, assay_name){

     ligand_cols = c(
                   "BMP2" = "#ffffb3",
                   "IFNG" = "#bebada",
                   "TGFB" = "#b3de69",
                   "PBS" = "#8dd3c7",
                   "HGF" = "#80b1d3",
                   "OSM" = "#fdb462",
                   "EGF" = "#fb8072")
   
  windsorize_probs <- get_win_probs(df, .01, .99)
  
  df_med <- df %>%
    prep_data(md = md) %>%
    gather("specimenName", "value", -feature) %>%
    inner_join(md, by = "specimenName") %>%
    select(ligand, secondLigand, experimentalTimePoint, replicate, collection, feature, value) %>%
    filter(!ligand == "EGF") %>%
    rename(Time = experimentalTimePoint,
           Ligand = ligand,
           Ligand2 = secondLigand,
           Replicate = replicate,
           Collection = collection) %>%
    group_by(feature, Collection, Time, Ligand) %>%
    summarise(value = median(value)) %>%
    ungroup() 
  
  p <- ggplot(df_med, aes(x=Time, y=value, colour=Ligand))+
    geom_line()+
    labs(title="EGF Normalized Line Graphs",
         x = "Time (hours)",
         y="Intensity (AU)") +
    scale_x_continuous(breaks = c(0,8,24,48)) +
    scale_color_manual( values = ligand_cols) +
    theme(axis.text.x = element_text(angle=90),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          strip.text = element_text(size=6, hjust=0),
            panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "gray95"))+
    facet_wrap(~feature,scales = "free_y")
  
  p
  
}

plot_line_graphs(df = RPPA, md, assay_name = "RPPA")


```

#### UMAP

```{r}

plot_umap(RPPA, md, assay_name = "RPPA")

```

#### PCA 

```{r}

plot_pca(RPPA, md, assay_name = "RPPA")

```

#### Introspect

```{r}

plot_introspect(RPPA, md, assay_name = "RPPA")

```

###  

### IF{.tabset .tabset-fade}

#### Heatmap  

```{r IF_figures, eval = TRUE}

boundedLog2 <- function(x) log2(x + .0001)

immunofluoresence <- load_data(assay_name = "IF") %>%
  mutate_if(is.numeric, .funs = boundedLog2) %>%
  filter(!feature == "EdUPositiveProportion")
  
plot_heatmap(immunofluoresence, md, assay_name = "IF", mark_rows = immunofluoresence$feature)

```

The row features in the heatmap are `r knitr::combine_words(immunofluoresence$feature)`

#### UMAP

```{r}

plot_umap(immunofluoresence, md, assay_name = "IF")

```

#### PCA 

```{r}

plot_pca(immunofluoresence, md, assay_name = "IF")
```

#### Introspect

```{r}

plot_introspect(immunofluoresence, md, assay_name = "IF")

```

### cycIF{.tabset .tabset-fade}

#### Heatmap 

```{r cycIF_figures, eval = TRUE}

cycIF <- load_data(assay_name = "cycIF") %>%
  filter(str_detect(feature, "_int_")) %>%
  mutate_if(is.numeric, log2)

cycIF_vars <- cycIF %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh <- quantile(cycIF_vars, probs = .5)
cycIF_hi_vars <- cycIF[cycIF_vars > var_thresh,]

plot_heatmap(cycIF_hi_vars, md, assay_name = "cycIF", mark_rows = c("egfr_2_af488fitc_int_mean_nuc","stat3_4_af488fitc_int_med_cytoplasm","met_6_af488fitc_int_mean_nuc"))

```


Intensity features included in the heatmap have a variance above the `r names(var_thresh) %>% str_remove("%")` percentile . 

#### Line graphs

```{r, fig.height=15, fig.width=15}

plot_line_graphs(cycIF, md, assay_name = "cycIF")

```


#### UMAP

```{r}
plot_umap(cycIF, md, assay_name = "cycIF")

```

#### PCA 

```{r}

plot_pca(cycIF, md, assay_name = "cycIF")
```

#### Introspect

```{r}

plot_introspect(cycIF, md, assay_name = "cycIF")
```

### RNAseq{.tabset .tabset-fade}

#### Heatmap  

```{r rnaSeq_figures, eval = TRUE}
RNAseq <- load_data(assay_name = "RNAseq") %>%
  rename(feature = ensembl_gene_id)

RNAseq_vars <- RNAseq %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh <- quantile(RNAseq_vars, probs = .9)
RNAseq_hi_vars <- RNAseq[RNAseq_vars > var_thresh,]
  
plot_heatmap(RNAseq_hi_vars, md, assay_name = "RNAseq")

```


Genes included in the heatmap have a variance above the `r names(var_thresh) %>% str_remove("%")` percentile . 

#### UMAP

```{r}

plot_umap(RNAseq, md, assay_name = "RNAseq")

```

#### PCA 

```{r}

plot_pca(RNAseq, md, assay_name = "RNAseq")
```

#### Introspect

```{r}

plot_introspect(RNAseq_hi_vars, md, assay_name = "RNAseq")

```

### GCP{.tabset .tabset-fade}

#### Heatmap  

```{r GCP_figures, eval = TRUE}
### delete histone when data is missing
GCP <- load_data(assay_name = "GCP") %>%
  rename(feature = histone) %>%
  drop_na()
  
plot_heatmap(GCP, md, assay_name = "GCP")

```

Histone probes with NA values in any sample are dropped from the GCP dataset.

#### Line graphs

```{r, fig.height=15, fig.width=15}

plot_line_graphs(GCP, md, assay_name = "GCP")

```

#### UMAP

```{r}

plot_umap(GCP, md, assay_name = "GCP")

```

#### PCA 

```{r}

plot_pca(GCP, md, assay_name = "GCP")
```

#### Introspect

```{r}

plot_introspect(GCP, md, assay_name = "GCP")

```

### ATACseq{.tabset .tabset-fade}

#### Heatmap  

```{r ATACseq_figures, eval = TRUE}

ATACseq <- load_data(assay_name = "ATACseq") %>%
  rename(feature = peak)

ATACseq_vars <- ATACseq %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh <- quantile(ATACseq_vars, probs = .95)
ATACseq_hi_vars <- ATACseq[ATACseq_vars > var_thresh,]
  
plot_heatmap(ATACseq_hi_vars, md, assay_name = "ATACseq")

```


Peaks included in the heatmap have a variance above the `r names(var_thresh) %>% str_remove("%")` percentile . 


#### UMAP

```{r}

plot_umap(ATACseq, md, assay_name = "ATACseq")

```

#### PCA 

```{r}

plot_pca(ATACseq, md, assay_name = "ATACseq")
```

#### Introspect

```{r}

plot_introspect(ATACseq_hi_vars, md, assay_name = "ATACseq")


```
 
 ##  
 
```{r l1000_figures, eval = FALSE}


```



```{r eval = FALSE}
df <- data.frame(time = as.factor(c(0,1,4,8,24,48)),
                 assay = factor(rep(c("RPPA","cycIF","L1000", "GCP","IF","ATACseq","RNAseq"), each = 6), levels = c("ATACseq","RNAseq","IF", "GCP","L1000","cycIF","RPPA"), ordered = TRUE),
                 dataPresent = c(
                   rep(TRUE, 6),
                   rep(TRUE, 6),
                   rep(TRUE, 6),
                   c(TRUE, FALSE, TRUE, TRUE, TRUE, TRUE),
                   c(TRUE, FALSE, FALSE, FALSE, TRUE, TRUE),
                   c(TRUE, FALSE, FALSE, FALSE, TRUE, TRUE),
                   c(TRUE, FALSE, FALSE, FALSE, TRUE, TRUE)),
                 ligand = rep(c("BMP2", "IFNG", "TGFB","PBS", "HGF", "OSM", "EGF"), times = 36)
                 )
p <- ggplot(df, aes(x = time, y = assay,  fill = dataPresent)) +
  geom_tile(colour = "black") + 
  scale_fill_manual(values = c("TRUE" =  "#68D7F3", "FALSE" ="#193A43")) +
  theme_bw()
p  


```