---
title: "MDD integrated analysis heatmap"
output: html_document
---

```{r setup, include=FALSE}
####ToDOos
#change distance calculation in clustering
#do cluster analysis to set cluster number
#fix Type colors and use in bar chart

knitr::opts_chunk$set(echo = FALSE, cache=TRUE, cache.lazy = FALSE, message=FALSE, warning = FALSE, fig.height=10, fig.width=12)

suppressMessages(library(tidyverse))
suppressMessages(library(ComplexHeatmap))
library(iheatmapr)
library(DT)
library(rrscale)
library(ggbeeswarm)
library(umap)
library(plotly)
library(cluster)
library(fpc)

pk_preprocess_level3 <- function(df, type){
  #filter to time 0, 24 and 48
  #median summarise
  df_pp <- df %>%
    filter(str_detect(experimentalCondition,"0|24|48")) %>%
    mutate(Type = type) %>%
    group_by(experimentalCondition, feature, Type) %>%
    summarise(value = median(value, na.rm = TRUE)) %>%
    ungroup()
  return(df_pp)
}

select_features <- function(df, var_quantile = 0){
  df_small <- df %>%
    group_by(feature) %>%
    mutate(feature_variance = var(value, na.rm = TRUE)) %>%
    ungroup() %>%
    filter(feature_variance >= quantile(feature_variance,
                                       probs = var_quantile,
                                       na.rm = TRUE)) %>%
    dplyr::select(-feature_variance)
  return(df_small)
}

get_TFs <- function(sheet, dir_path, pattern) {
  #read in and combine MDD TF scores
TFs_input <- dir(path = dir_path,
                 pattern = pattern,
                 full.names = TRUE) %>%
  map(readxl::read_excel, col_types = c("text"), sheet = sheet, .name_repair = "universal") %>%
  bind_rows() %>%
  mutate(Rank = as.numeric(Rank),
         experimentalTime = str_extract(Query.Name, "[24][48]"),
         experimentalTime = as.numeric(experimentalTime),
         ligand = str_remove(Query.Name, "ctrl vs "),
         ligand = str_remove(ligand, ",.*"),
         condition = paste0(ligand, "_", experimentalTime),
         value =(1633-Rank)/1632,
         feature = TF,
         Library_only = str_remove(Library, ",.*"),
         Library_only = str_replace(Library_only, " ", "--"))
}

#' rrscale a numeric vector
rrscaleValues <- function(x, zeros = .001, ncores = 4){
  x_rr <- as.matrix(x) %>%
    rrscale(zeros = zeros, ncores = ncores)
  return(x_rr)
}

getrrDetails <- function(x){
  T_name <- map(x, function(xx) xx[["T_name"]]) %>%
    unlist
  par_hat <- map(x, function(xx) xx[["par_hat"]]) %>%
    unlist
  df <- tibble(feature = names(x),
               par_hat = par_hat,
               T_name = T_name)
  return(df)
}

#' df a dataframe with feature, value, condition and Type columns. The values must
#' be all numeric and transformed to common scales

#prepare a datamatrix from a dataframe of data and metadta
prep_hm_matrix <- function(df){
  
  condition_order <-  c("ctrl_0",paste(rep(c("PBS", "HGF", "OSM", "EGF","BMP2", "IFNG", "TGFB"), each = 2), rep(c(24, 48),  times = 14),   sep = "_") )
  
  condition_order <- condition_order[condition_order %in% df$experimentalCondition]
  df_sp <- df %>%
    dplyr::select(feature, value, experimentalCondition) %>%
    spread(key = experimentalCondition, value = value, fill = median(df$value, na.rm = TRUE)) %>%
    dplyr::select(feature, condition_order)
 #   drop_na()
  
  df_as_matrix <- df_sp %>%
    dplyr::select(-feature) %>%
    as.matrix()
  rownames(df_as_matrix) <- df_sp$feature
  return(df_as_matrix)
}

prep_hm_annotations <- function(df){
  df_as_matrix <- prep_hm_matrix(df)
  #create top annotations
  ann_nv_pairs <- df %>%
    dplyr::select(feature, Type) %>%
    distinct()
#    drop_na()
  
  ann_df <- tibble(feature = rownames(df_as_matrix)) %>%
    left_join(ann_nv_pairs, by = "feature") %>%
    dplyr::select(Type)
  return(ann_df)
}

get_iheatmap <- function(df_as_matrix, ...) {
  #Create the heatmap
  hm <- main_heatmap(data = df_as_matrix,
                     name = "score")
}

format_hm <- function(hm, k = 6, cluster_method = "hclust",...){
  hm_mod<- hm %>%
    add_row_labels(font = list(size = 4),
                   side = "right") %>%
    add_col_labels() %>%
    add_row_annotation(ann_df,
                       side = "left",
                       size = 0.05)  %>%
    modify_layout(list(margin = list(r = 120)))
  if(!k==0){
    hm_mod <- hm_mod %>%
      add_row_clustering(name = paste0("Cluster(",cluster_method,")"),
                         k = k,
                         method = cluster_method,
      ) %>%
      add_col_summary(size = .1,
                      groups = paste0("Cluster(",cluster_method,")"))
  }
  return(hm_mod)
}

#apply rrscale on all values of an assay
rrscale_assay <- function(df_long, zscore_cutoff = Inf){
  df_sp <- df_long %>%
    spread(feature, value)
  
  rr_objects <- df_sp %>%
    select(-Type, -experimentalCondition) %>%
    as.matrix() %>%
    rrscale(zeros = 0.01,z = zscore_cutoff)
  
  rr_mat <- rr_objects[["RR"]] %>%
    as_tibble()
  
  df_rr <- df_sp %>%
    select(experimentalCondition, Type) %>%
    bind_cols(rr_mat) %>%
    gather("feature", "value", -Type, -experimentalCondition)
  
  return(df_rr)
}

```


```{r read_data_and_metadata}

if(file.exists("R/all_pk_data.rda")){
    load("R/all_pk_data.rda")
} else {
  
md <- read_csv("metadata/MDD_sample_annotations.csv")
#Read and and process ATACseq metadata
ATACseq_metadata <- read_csv("ATACseq/Metadata/MDD_ATACseq_peakMetadata.csv") %>%
  select(peak, hgnc_symbol, annotation) %>%
  drop_na() %>%
  filter(str_detect(annotation, "Promoter|3' UTR|5' UTR|1st Exon")) %>%
  dplyr::select(-annotation)

#Read in and process ATACseq values
ATACseq_values <- read_csv("ATACseq/Data/MDD_ATACseq_Level3.csv") %>%
  rename(feature = peak) %>%
  gather(specimenID, value = value, -feature) %>%
  left_join(md, by = "specimenID") %>%
  inner_join(ATACseq_metadata, by = c("feature" = "peak")) %>%
    mutate(feature = paste0(hgnc_symbol, "_ATAC")) %>%
  pk_preprocess_level3(type = "ATACseq") %>%
  group_by(experimentalCondition, feature, Type) %>%
  summarise(value = max(value)) %>%
  ungroup()

#Read in and process ATACseq motif values
ATACseq_motif_values <- read_csv("ATACseq/Data/MDD_ATACseq_MotifZscores.csv") %>%
  rename(feature = motif) %>%
  gather(specimenID, value = value, -feature) %>%
  left_join(md, by = "specimenID") %>%
  mutate(feature = str_remove(feature, ".*_"),
         feature = paste0(feature, "_motif")) %>% 
  pk_preprocess_level3(type = "ATACseq_motif")

#Read in and process cycIF values
biomarkers <- c("p21waf1cip1_3_af647cy5_int_mean_nuc",
                "cyclind1_8_af488fitc_int_mean_nuc",
                "ki67_5_efluor570cy3_int_mean_nuc",
                "egfr_2_af488fitc_int_mean_nuc",
                "egfr_2_af488fitc_int_mean_cytoplasm",
                "met_6_af488fitc_int_mean_nuc",
                "met_6_af488fitc_int_mean_cytoplasm",
                "ndg1pt346_3_af488fitc_int_mean_nuc",
                "ndg1pt346_3_af488fitc_int_mean_cytoplasm",
                "s6_6_pecy3_int_mean_nuc",
                "s6_6_pecy3_int_mean_cytoplasm",
                "s6ps235s236_5_af647cy5_int_mean_nuc",
                "s6ps235s236_5_af647cy5_int_mean_cytoplasm",
                "s6ps240244_5_af488fitc_int_mean_nuc",
                "s6ps240244_5_af488fitc_int_mean_cytoplasm",
                "cateninbeta_4_af647cy5_int_mean_nuc",
                "cateninbeta_4_af647cy5_int_mean_cytoplasm",
                "nfkbp65_2_af647cy5_int_mean_nuc",
                "cjun_9_af488fitc_int_mean_nuc",
                "stat1ps727_2_pecy3_int_mean_nuc",
                "stat1alphaisoform_8_af647cy5_int_mean_nuc",
                "pdl1_6_af647cy5_int_mean_cytoplasm",
                "stat3_4_af488fitc_int_mean_nuc",
                "ecadherin_7_af647cy5_int_mean_nuc",
                "ecadherin_7_af647cy5_int_mean_cytoplasm",
                "vimentin_3_af555cy3_int_mean_nuc",
                "vimentin_3_af555cy3_int_mean_cytoplasm",
                "cytokeratin7human_4_af555cy3_int_mean_nuc",
                "cytokeratin7human_4_af555cy3_int_mean_cytoplasm",
                "cytokeratin18_7_af488fitc_int_mean_nuc",
                "cytokeratin18_7_af488fitc_int_mean_cytoplasm",
                "hes1_9_af647cy5_int_mean_nuc",
                "lc3ab_7_af555cy3_txt_standev_cytoplasm")

cycIF_values <- read_csv("cycIF/Data/MDD_cycIF_Level3.csv") %>%
  filter(feature %in% biomarkers) %>%
  gather(specimenID, value = value, -feature) %>%
  spread(key = feature, value = value) %>%
  mutate(egfr_2_af488fitc_int_mean_cell = (egfr_2_af488fitc_int_mean_nuc + egfr_2_af488fitc_int_mean_cytoplasm)/2,
         met_6_af488fitc_int_mean_cell = (met_6_af488fitc_int_mean_nuc + met_6_af488fitc_int_mean_cytoplasm)/2,
         ndg1pt346_3_af488fitc_int_mean_cell = (ndg1pt346_3_af488fitc_int_mean_nuc + ndg1pt346_3_af488fitc_int_mean_cytoplasm)/2,
         s6_6_pecy3_int_mean_cell = (s6_6_pecy3_int_mean_nuc + s6_6_pecy3_int_mean_cytoplasm)/2,
         s6ps235s236_5_af647cy5_cell = (s6ps235s236_5_af647cy5_int_mean_nuc + s6ps235s236_5_af647cy5_int_mean_cytoplasm)/2,
         s6_6_pecy3_int_mean_cell = (s6_6_pecy3_int_mean_nuc + s6_6_pecy3_int_mean_cytoplasm)/2,
         s6ps240244_5_af488fitc_int_mean_cell = (s6ps240244_5_af488fitc_int_mean_nuc + s6ps240244_5_af488fitc_int_mean_cytoplasm)/2,
         cateninbeta_4_af647cy5_int_mean_cell = (cateninbeta_4_af647cy5_int_mean_nuc + cateninbeta_4_af647cy5_int_mean_cytoplasm)/2,
         ecadherin_7_af647cy5_int_mean_cell = (ecadherin_7_af647cy5_int_mean_nuc + ecadherin_7_af647cy5_int_mean_cytoplasm)/2,
         vimentin_3_af555cy3_int_mean_cell = (vimentin_3_af555cy3_int_mean_nuc + vimentin_3_af555cy3_int_mean_cytoplasm)/2,
         cytokeratin7human_4_af555cy3_int_mean_cell = (cytokeratin7human_4_af555cy3_int_mean_nuc + cytokeratin7human_4_af555cy3_int_mean_cytoplasm)/2,
         cytokeratin18_7_af488fitc_int_mean_cell = (cytokeratin18_7_af488fitc_int_mean_nuc + cytokeratin18_7_af488fitc_int_mean_cytoplasm)/2) %>%
  select(-egfr_2_af488fitc_int_mean_nuc,
         -egfr_2_af488fitc_int_mean_cytoplasm,
         -met_6_af488fitc_int_mean_nuc,
         -met_6_af488fitc_int_mean_cytoplasm,
         -ndg1pt346_3_af488fitc_int_mean_nuc,
         -ndg1pt346_3_af488fitc_int_mean_cytoplasm,
         -s6_6_pecy3_int_mean_nuc,
         -s6_6_pecy3_int_mean_cytoplasm,
         -s6ps235s236_5_af647cy5_int_mean_nuc,
         -s6ps235s236_5_af647cy5_int_mean_cytoplasm,
         -s6ps240244_5_af488fitc_int_mean_nuc,
         -s6ps240244_5_af488fitc_int_mean_cytoplasm,
         -cateninbeta_4_af647cy5_int_mean_nuc,
         -cateninbeta_4_af647cy5_int_mean_cytoplasm,
         -ecadherin_7_af647cy5_int_mean_nuc,
         -ecadherin_7_af647cy5_int_mean_cytoplasm,
         -vimentin_3_af555cy3_int_mean_nuc,
         -vimentin_3_af555cy3_int_mean_cytoplasm,
         -cytokeratin7human_4_af555cy3_int_mean_nuc,
         -cytokeratin7human_4_af555cy3_int_mean_cytoplasm,
         -cytokeratin18_7_af488fitc_int_mean_nuc,
         -cytokeratin18_7_af488fitc_int_mean_cytoplasm) %>%
  gather(feature, value, -specimenID) %>%
left_join(md, by = "specimenID") %>%
  mutate(feature = str_remove(feature, "_.*"),
         feature = paste0(feature, "_cycIF")) %>%
  pk_preprocess_level3(type = "cycIF")

#Read in and process GCP values
GCP_values <- read_csv("GCP/Data/MDD_GCP_Level3.csv") %>%
  rename(feature = histone) %>%
  gather(specimenID, value = value, -feature) %>%
  left_join(md, by = "specimenID") %>%
  mutate(value = 2^value) %>%
  filter(!feature == "H3K27ac1K36me0") %>%
  pk_preprocess_level3(type = "GCP") %>%
  replace( is.na(.), .001)

#Read in Hallmark pathways that are based on RNAseq data
hallmark_pathways <- read_csv("RNAseq/Data/MDD_RNAseq_HallmarkNES.csv") %>%
  gather(condition, value, -variable) %>%
  rename(feature = variable) %>%
  rename(experimentalCondition = condition) %>%
  mutate(Type = "Hallmark") %>%
  pk_preprocess_level3(type =  "Hallmark")

#Read in and process IF values
IF_values <- read_csv("IF/Data/MCF10A_IF_Ilastik_Image_File.csv") %>%
  select(-ImageNumber, -barcode, -WellIndex, -collection, -ligand) %>%
  gather(feature, value, -specimenName, -time, -replicate) %>%
  mutate(experimentalTimePoint = time,
         experimentalCondition = str_remove(specimenName, "_C[12]_.")) %>%
  filter(!feature =="AreaShape_EulerNumber",
         !str_detect(feature,"CellMask")) %>%
  pk_preprocess_level3(type = "IF") %>%
replace( is.na(.), 1) %>%
  group_by(experimentalCondition, feature, Type) %>%
  summarise(value = median(value, na.rm = TRUE)) %>%
  ungroup()

#Read in RNAseq values
RNA_values <- read_csv("RNAseq/Data/MDD_RNAseq_med.csv") %>%
  drop_na() %>%
  mutate(feature = paste0(feature, "_RNA")) %>%
  rename(experimentalCondition = condition) %>%
  pk_preprocess_level3("RNAseq")

#Read in RPPA pathways scores, mean summarise within condition
RPPA_pathways <- read_csv("RPPA/Data/MDD_Pathways_Score.csv") %>%
  rename(specimenID = X1) %>%
  gather(key = feature, value = value, -specimenID) %>%
  inner_join(md, by = "specimenID") %>%
  dplyr::select(experimentalCondition, feature, value) %>%
  pk_preprocess_level3(type =  "RPPA pathway")

#Read in RPPA values and merge with metadata
RPPA_values <- read_csv("RPPA/Data/MDD_RPPA_Level3.csv") %>%
  mutate(feature = paste0(antibody,"_RPPA")) %>%
  gather(specimenID, value = value, -feature, -antibody) %>%
  left_join(md, by = "specimenID") %>%
  mutate(value = 2^value) %>%
  pk_preprocess_level3(type = "RPPA")


TFs <- get_TFs(dir_path = "RNAseq/Data/ChEA3_results_MD_OHSU", pattern = "ctrl_vs_.*xlsx", sheet =1)
TFs_details <- map(2:7, get_TFs, dir_path = "RNAseq/Data/ChEA3_results_MD_OHSU", pattern = "ctrl_vs_.*xlsx") %>%
  bind_rows()

TFs_input <- TFs_details %>%
  dplyr::select(Query.Name, TF, Odds.Ratio, FDR, Library_only) %>%
  right_join(TFs, by = c("Query.Name", "TF", "Library_only"))

TFs_values <- TFs_input %>%
  dplyr::select(condition, Odds.Ratio, feature) %>%
  rename(experimentalCondition = condition) %>%
  mutate(feature = paste0(feature, "_TF")) %>%
  rename(value = Odds.Ratio) %>%
  mutate(value = as.numeric(value),
         Type = "ChEA3 TF")

# TFs_down <- read_csv("RNAseq/Data/ChEA3_TFs_down.csv") %>%
#   gather(feature, value, -condition) %>%
#   mutate(Type = "ChEA3 TF Down")

pk_all <- bind_rows(ATACseq_values, ATACseq_motif_values, cycIF_values, GCP_values, hallmark_pathways, IF_values, RNA_values, RPPA_pathways, RPPA_values, TFs_values)

save(pk_all,
     md,
     file = "R/all_pk_data.rda") 

# phenotypes <- read_csv("cycIF/Data/MDD_cycIF_Level4.csv") %>%
#   filter(str_detect(feature, "neighbor"))
}
```


```{r prepare_datasets}
ATACseq_variance_probs_thresh <- .995
ATACseq_motif_variance_probs_thresh <- .5
GCP_variance_probs_thresh <- 0

RPPA_variance_probs_thresh <- 0
odds_ratio_thresh <- 5

if(file.exists("R/selected_pk_data.rda")){
    load("R/selected_pk_data.rda")
} else {
  
ATACseq_selected <- select_features(ATACseq_values, ATACseq_variance_probs_thresh)

ATACseq_motif_selected <- select_features(ATACseq_motif_values, ATACseq_motif_variance_probs_thresh)

# #Select top TFs of each ligand

TFs_selected_names <- TFs_values %>%
  spread(key = feature, value = value) %>%
  #mutate(Ligand = str_remove(experimentalCondition, "_.*")) %>%
  group_by(experimentalCondition) %>%
  summarise_if(is.numeric, min) %>%
  ungroup %>%
  gather(key = TF, value = Odds.Ratio, -experimentalCondition) %>%
  group_by(experimentalCondition) %>%
  #filter(Odds.Ratio >= odds_ratio_thresh)%>%
  arrange(Odds.Ratio) %>%
  top_n(25) %>%
  ungroup() %>%
  spread(key = TF, value = Odds.Ratio) %>%
  dplyr::select(-experimentalCondition)

TFs_selected <- TFs_values %>%
  spread(key = feature, value = value)  %>%
  dplyr::select(experimentalCondition, colnames(TFs_selected_names)) %>%
  gather(feature, value, -experimentalCondition) %>%
  mutate(Type = "ChEA3 TF") %>%
  drop_na()

#filter RPPA features on variance
RPPA_selected <- select_features(RPPA_values, RPPA_variance_probs_thresh)

#filter GCP features on variance
GCP_selected <- select_features(GCP_values, GCP_variance_probs_thresh)

#use RNAseq genes filterd on variance within each condition
RNAseq_variance_genes <- read_csv("RNAseq/Data/MDD_geneList200_1560genes.csv") %>%
  mutate(feature= paste0(hgnc_symbol, "_RNA")) %>%
  dplyr::select(-hgnc_symbol)

RNAseq_selected <- RNA_values %>%
  inner_join(RNAseq_variance_genes)

pk_selected <- bind_rows(ATACseq_selected, ATACseq_motif_selected, cycIF_values, GCP_selected, hallmark_pathways, IF_values, RNAseq_selected, RPPA_pathways, RPPA_selected, TFs_selected)

save(pk_selected,
     ATACseq_selected,
     ATACseq_motif_selected,
     cycIF_values,
     GCP_selected,
     hallmark_pathways,
     IF_values,
     RNAseq_selected,
     RPPA_selected,
     RPPA_pathways,
     TFs_selected,
     file = "R/selected_pk_data.rda")
}

zscore_cutoff <- Inf
if(file.exists("R/rrscale_assay.rda")){
  load("R/rrscale_assay.rda")
} else {
  ATACseq_selected_rr <- rrscale_assay(ATACseq_selected, zscore_cutoff = zscore_cutoff)
  ATACseq_motif_selected_rr <- rrscale_assay(ATACseq_motif_selected, zscore_cutoff = zscore_cutoff)
  cycIF_values_rr <- rrscale_assay(cycIF_values, zscore_cutoff = zscore_cutoff)
  GCP_selected_rr <- rrscale_assay(GCP_selected, zscore_cutoff = zscore_cutoff)
  hallmark_pathways_rr <- rrscale_assay(hallmark_pathways, zscore_cutoff = zscore_cutoff)
  IF_values_selected_rr <- IF_values %>%
    filter(str_detect(feature, "Intensity"),
           !str_detect(feature, "Std")) %>%
    rrscale_assay(zscore_cutoff = zscore_cutoff)
  RPPA_selected_rr <- rrscale_assay(RPPA_selected, zscore_cutoff = zscore_cutoff)
  RPPA_pathways_rr <- rrscale_assay(RPPA_pathways, zscore_cutoff = zscore_cutoff)
  RNAseq_selected_rr <- rrscale_assay(RNAseq_selected, zscore_cutoff = zscore_cutoff)
  TFs_selected_rr <- rrscale_assay(TFs_selected, zscore_cutoff = zscore_cutoff)
  
  pk_selected_rr <-  bind_rows(ATACseq_selected_rr, ATACseq_motif_selected_rr, cycIF_values_rr, GCP_selected_rr, hallmark_pathways_rr, IF_values_selected_rr, RNAseq_selected_rr, RPPA_pathways_rr, RPPA_selected_rr, TFs_selected_rr)
  save(pk_selected_rr, file = "R/rrscale_assay.rda")
}

if(file.exists("R/rrscale_assay_all.rda")){
  load("R/rrscale_assay_all.rda")
} else {
  ATACseq_rr <- rrscale_assay(ATACseq_values, zscore_cutoff = zscore_cutoff)
  ATACseq_motif_rr <- rrscale_assay(ATACseq_motif_values, zscore_cutoff = zscore_cutoff)
  cycIF_values_rr <- rrscale_assay(cycIF_values, zscore_cutoff = zscore_cutoff)
  GCP_rr <- rrscale_assay(GCP_values, zscore_cutoff = zscore_cutoff)
  hallmark_pathways_rr <- rrscale_assay(hallmark_pathways, zscore_cutoff = zscore_cutoff)
    IF_values_rr <- IF_values %>%
    filter(str_detect(feature, "Intensity"),
           !str_detect(feature, "Std")) %>%
    rrscale_assay(zscore_cutoff = zscore_cutoff)
  RPPA_rr <- rrscale_assay(RPPA_values, zscore_cutoff = zscore_cutoff)
  RPPA_pathways_rr <- rrscale_assay(RPPA_pathways, zscore_cutoff = zscore_cutoff)
  RNAseq_rr <- rrscale_assay(RNA_values, zscore_cutoff = zscore_cutoff)
  TFs_rr <- rrscale_assay(TFs_values, zscore_cutoff = zscore_cutoff)

  pk_rr <-  bind_rows(ATACseq_rr, ATACseq_motif_rr, cycIF_values_rr, GCP_rr, hallmark_pathways_rr, IF_values_rr, RNAseq_rr, RPPA_pathways_rr, RPPA_rr, TFs_rr)
  save(pk_rr, file = "R/rrscale_assay_all.rda")
}


```

## {.tabset .tabset-fade}

### heatmap

```{r interactiveHeatmaps, fig.height = 8}



df_as_matrix <-  prep_hm_matrix(pk_selected_rr) %>%
  t %>%
  as.matrix() %>%
  scale %>%
  t
ann_df <- prep_hm_annotations(pk_selected_rr)

#prep column anotation
phenotypes <- IF_values %>%
  filter(str_detect(feature,"Number|Fourth|Cell_Count")) %>%
  spread(feature, value) %>%
  mutate(experimentalCondition = factor(experimentalCondition, levels = colnames(df_as_matrix),ordered = TRUE)) %>%
  arrange(experimentalCondition) %>%
  select(matches("Neighbor|cluster|Cell_Count"))

clust_num <- 11
cluster_method = "kmeans"
hm <- get_iheatmap(df_as_matrix, assay_name = "Prior knowledge, rrscale-transformed") %>%
  format_hm(k = clust_num, cluster_method = cluster_method) %>%
  add_col_annotation(annotation = phenotypes, side = "top")
hm

# side_anots <-  prep_hm_matrix(pk_selected_rr) %>%
#   #replace( is.na(.), 0) %>%
#   t %>%
#   scale %>%
#   t %>%
#   as.data.frame %>%
#   dplyr::select(  PBS_48, PBS_24, ctrl_0)
# 
# score_pal <- colorRampPalette(hm@colorbars@listData$score@colors, interpolate = "spline")(10)
# hm_pbs <- add_row_annotation(hm,
#                              side_anots,
#                              side = "left",
#                              colors = list(ctrl_0 = score_pal, PBS_24 = score_pal, PBS_48 = score_pal),
#                              size = .07,
#                              inner_buffer = 0)
# hm_pbs
#   
```



### heatmap with callouts 

```{r staticHeatmap, fig.width=14}
#create a second, static heatmap to have larger labels
y_order <- hm@yaxes@listData[["y"]]@order %>%
  rev()

row_nums <- which(rownames(df_as_matrix) %in% c(hallmark_pathways$feature, RPPA_pathways$feature))
ra <- rowAnnotation(foo = anno_mark(at = c(row_nums), labels = rownames(df_as_matrix)[row_nums]))
#Create the heatmap
shm <- Heatmap(matrix = df_as_matrix,
               name = "",
               #         column_title = paste0("Correlation of ",assay_name, title_suffix),
               #         top_annotation = create_top_annotations(df, md),
               #         left_annotation = haRow,
               right_annotation = ra,
               show_row_names = FALSE,
               row_order = y_order,
               show_column_names = TRUE,
               cluster_rows = FALSE,
               cluster_columns = FALSE,
               #col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
               na_col = "grey")
shm

```


### heatmap no ctrl/PBS

```{r interactiveHeatmapsnoPBS, fig.height = 8}

df_no_ctrl_PBS <- prep_hm_matrix(pk_selected_rr) %>%
  t %>%
  as.data.frame() %>%
    rownames_to_column(var = "experimentalCondition") %>%
    filter(!experimentalCondition %in% c("ctrl_0", "PBS_24", "PBS_48")) 

df_as_matrix <- df_no_ctrl_PBS %>%
  dplyr::select(-experimentalCondition) %>%
  # drop_na() %>%
  as.matrix() %>%
  scale %>%
  t
colnames(df_as_matrix) <- df_no_ctrl_PBS$experimentalCondition

ann_df <-  pk_selected_rr %>%
  filter(!experimentalCondition %in% c("ctrl_0", "PBS_24", "PBS_48")) %>%
 prep_hm_annotations()

clust_num <- 13
cluster_method = "kmeans"
hm <- get_iheatmap(df_as_matrix, assay_name = "Prior knowledge, rrscale-transformed") %>%
  format_hm(k = clust_num, cluster_method = cluster_method)
hm

```

### heatmap no ctrl/PBS with callouts 

```{r staticHeatmapnoCtrlPBS, fig.width=14}
#create a second, static heatmap to have larger labels
y_order <- hm@yaxes@listData[["y"]]@order %>%
  rev()

row_nums <- which(rownames(df_as_matrix) %in% c(hallmark_pathways$feature, RPPA_pathways$feature))
ra <- rowAnnotation(foo = anno_mark(at = c(row_nums), labels = rownames(df_as_matrix)[row_nums]))
#Create the heatmap
shm <- Heatmap(matrix = df_as_matrix,
               name = "",
               #         column_title = paste0("Correlation of ",assay_name, title_suffix),
               #         top_annotation = create_top_annotations(df, md),
               #         left_annotation = haRow,
               right_annotation = ra,
               show_row_names = FALSE,
               row_order = y_order,
               show_column_names = TRUE,
               cluster_rows = FALSE,
               cluster_columns = FALSE,
               #col = circlize::colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
               na_col = "grey")
shm

```


### method  

This dataset combines MCF10A Molecular Deep Dive (MDD) RNAseq, Reverse Phase Protein Array(RPPA), Global Chromatin Profiling (GCP) histone modifications, immunofluorescence (IF), cyclic immunofluorescence (cycIF) and ATACseq chromatin configuration data with prior knowledge of signaling pathways, DNA motifs and transcription factors. All data is collected at 0, 24 and 48 hours. Most of the data is median summarized from three replicate experiments run in sequential weeks. It is then rrscale transformed and then zscored by feature.   

RRscale  
Use rrscale https://CRAN.R-project.org/package=rrscale to transform features in the heatmap before clustering.  This method performs a box-cox gaussianizing transformation on each feature, zscores the results then eliminates outliers with absolute values greater than `r zscore_cutoff`. All values from each assay was rrscaled as a matrix with features in the columns and conditions in the rows. Because of the differences in scales in the morphology, spatial and statistical values in the IF dataset, only the intensity values are included.      

This analysis is based on the LINCS Molecular Deep Dive data files on Synapse at  https://www.synapse.org/#!Synapse:syn2862345/wiki/588244.  


RNAseq Values  
A gene list representative of the largest gene expression changes across all ligand treatments was produced from differential expression analyses of each experimental condition compared to the control condition. For each experimental condition, a differential expression analysis was performed using the R package DESeq2 (1.24.0), with shrunken log2 fold change estimates calculated using the apeglm method. Each list was filtered to include only genes with an adjusted p value < 0.05, then ranked by the absolute value of their log2 fold change estimate. The 25 most differentially expressed genes per condition were selected. Redundant genes were removed, yielding a list of `r length(unique(RNAseq_selected$feature))` unique genes included in the heatmap.  

RPPA Values  
Median summarize the values across the replicates  
Select the RPPA antibodies with a variance above the `r RPPA_variance_probs_thresh*100`th percentile. 
This yields `r length(unique(RPPA_selected$feature))` high variance antibodies in the heatmap.  

GCP Values  
Median summarize the values across the replicates  
Select the GCP histones with a variance above the `r GCP_variance_probs_thresh*100`th percentile. This yields `r length(unique(GCP_selected$feature))` high variance histones in the heatmap.  

IF Values  
Filter out the CellMask intensities and use the remaining `r length(unique(IF_values$feature))` IF features  
Median summarize the values across the replicates  

cycIF Values  
Select `r length(unique(cycIF_values$feature))` cycIF features based on their biological relevance  
Median summarize the values across the replicates  

ATACseq Values  
Median summarize the values across the replicates  
Select the genes with a variance above the `r ATACseq_variance_probs_thresh*100`th percentile. This yields `r length(unique(ATACseq_selected$feature))` high variance genes in the heatmap.  

RPPA pathways   
Use the MDACC algorithm and OHSU's antibody-to-pathway assignments to create pathway scores for each , 0, 24 and 48 hour sample.  
Median summarize the pathway scores to the treatment level.  
There are `r length(unique(RPPA_pathways$feature))` RPPA pathways in the heatmap.    

Hallmark pathways    
Hallmark pathways are based on the RNAseq data.  
There are `r length(unique(hallmark_pathways$feature))` Hallmark pathways in the heatmap.    

ATACseq motifs  
Read in supplied ATACseq motif values. 
There are `r length(unique(ATACseq_motif_selected$feature))` ATACseq motifs in the heatmap.  

ChEA3 TF Values  
There are `r length(unique(TFs_selected$feature))` ChEA3 TFs in the heatmap.    

Clustering  
The hierarchical clustering uses the complete linkage method to find similar clusters. The number of clusters as shown in the side annotation bar. It is chosen as `r clust_num` based on gap statistic analysis. The scatter plot above the heatmap shows the mean value for each cluster within each condition.  


```{r featureTypes, fig.width=8, fig.height=4}
#create a bar chart of the feature types
df <- pk_selected_rr %>%
  drop_na() %>%
  select(feature, Type) %>%
  distinct()
p <- ggplot(df, aes(x = Type, fill = Type)) +
  geom_bar() +
  labs(title = paste("Type distribution of the",length(unique(df$feature)), "features in the MDD integrated dataset"),
       x ="Feature type") +
  theme(axis.text = element_text(angle = 90))
p

```

### rrscale diagnostics

```{r beforeAfterrr, fig.height=4}

set.seed(42)
df <- pk_selected_rr %>%
  dplyr::select(Type, feature) %>%
  distinct() %>%
  group_by(Type) %>%
  sample_n(4) %>%
  ungroup() %>%
  drop_na() %>%
  inner_join(pk_selected_rr, by = c("feature", "Type")) %>%
  arrange(Type) %>%
  rename(value_rr = value) %>%
  left_join(pk_selected, by = c("feature", "Type"))

p <- ggplot(df, aes(x = value, fill = Type)) +
  geom_density() +
  #guides(fill = FALSE) +
  facet_wrap(~feature, scales = "free", ncol = 10) +
  theme(strip.text.x = element_text(size = 4))
p

p <- ggplot(df, aes(x = value_rr, fill = Type)) +
  geom_density() +
  facet_wrap(~feature, scales = "free", ncol = 10) +
  theme(strip.text.x = element_text(size = 4))
p

winsorizeValues <- function(x, limits = c(-4, 4)){
  x[x<limits[1]] <- limits[1]
  x[x > limits[2]] <- limits[2]
  return(x)
}

df <- pk_selected_rr %>%
  mutate(value = winsorizeValues(value))


p <- ggplot(pk_selected_rr, aes(x = value, colour = Type)) +
  geom_density(alpha = .2)+ facet_wrap(~Type, scales = "free") +
  labs(title = "rrscale densities")
p

p <- ggplot(df, aes(x = value, colour = Type)) +
  geom_density(alpha = .2)+ facet_wrap(~Type, scales = "free") +
  labs(title = "Winsorized rrscale densities")
p

```

### cluster analysis

```{r gapStatistics, fig.width=4, fig.height=4, eval = TRUE}
#implement gap statistic analysis

hafun <- function(x, k){
    hm <- get_iheatmap(x, assay_name = "Prior knowledge, rrscale-transformed") %>%
  format_hm(k = k, cluster_method = cluster_method)
    Cluster <- hm@plots@listData$score@data %>%
      as.data.frame() %>%
      mutate(cluster = hm@plots@listData$Cluster@data) %>%
      drop_na() %>%
      dplyr::select(cluster) %>%
      as.list()
}

df_as_matrix <- pk_selected_rr %>% 
spread(experimentalCondition, value) %>%
  drop_na() %>%
  gather(experimentalCondition, value, -Type, -feature) %>%
  prep_hm_matrix
if(file.exists(paste0("R/gss_",cluster_method,".rda"))){
  load(paste0("R/gss_",cluster_method,".rda"))
} else {
  gss <- clusGap(df_as_matrix, FUN = hafun, K.max = 20, B = 100, verbose = FALSE)
  save(gss, file = paste0("R/gss_",cluster_method,".rda"))
}

plot_gap = function(x) {
  gstab = data.frame(x$Tab, k = seq_len(nrow(x$Tab)))
  ggplot(gstab, aes(k, gap)) + 
    geom_line() +
    geom_errorbar(aes(ymax = gap + SE.sim,
                      ymin = gap - SE.sim), width=0.1) +
    geom_point(size = 3, col=  "red") +
    labs(title = paste("Gap analysis to determine cluster number,",cluster_method))
}
plot_gap(gss)

```


This analysis can be interpreted to select the number of clusters in the dataset. It is based on a method described in Modern Statistics for Modern Biology by Susan Holmes and Wolfgang Huber http://web.stanford.edu/class/bios221/book/Chap-Clustering.html and is excerpted below.  

Taking the logarithm of the within-sum-of-squares (log(WSSk)) and comparing it to averages from simulated data with less structure can be a good way of choosing k. This is the basic idea of the gap statistic introduced by Tibshirani, Walther, and Hastie (2001). We compute log(WSSk) for a range of values of k, the number of clusters, and compare it to that obtained on reference data of similar dimensions with various possible ‘non-clustered’ distributions. We can use uniformly distributed data as we did above or data simulated with the same covariance structure as our original data.

The default choice for the number of clusters, k1, is the first value of k for which the gap is not larger than the first local maximum minus a standard error (see the manual page of the clusGap function). ... the choice recommended by Tibshirani, Walther, and Hastie (2001) is the smallest k such that gap(k)≥gap(k+1)−s′k+1...

The first local maximum is k = 3 which does not yield informative clusters. The second local maximum is k = 11 and then k = `r clust_num-1`. This analysis is choosing to use k = `r clust_num` which is is the first value where the gap value minus its standard error is less than this local maximum at `r clust_num-1`.  


<bksp>

<bksp>

###

#### Correlations  

```{r, include=TRUE, fig.align="center", fig.cap=c("Correlation heatmap"), echo=FALSE, eval = FALSE}

knitr::include_graphics("./Correlation_heatmap.pdf")

```

The interactive table below can be used to mine correlation values between the selected features in the columns and the entire set of features. Each of the `r length(unique(pk_rr$feature))` features is represented by a vector of 15 rrscaled values. Spearman correlations are calculated using all complete observations.  

```{r specificCorrelations}
df_sp <- pk_rr %>%
  spread(experimentalCondition, value) 

df_mat <- df_sp %>%
  select_if(is.numeric) %>%
            as.matrix() %>%
            t %>%
            scale(scale = FALSE)

selected_cor_features <- df_sp %>%
  filter(str_detect(feature, "Apop|Caspase")) %>%
  gather(condition, value, -feature, -Type) %>%
  dplyr::select(-Type) %>%
  spread(feature, value) %>%
  select_if(is.numeric) %>%
  as.matrix()

selectedCor <- cor(df_mat, selected_cor_features, use = "complete.obs",method = "spearman") %>%
  as_tibble() %>%
  mutate_if(is.numeric, signif, digits = 3) %>%
  mutate(feature = df_sp$feature) %>%
  dplyr::select(feature, everything())

datatable(selectedCor)
```
