---
title: "MDD standard figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE,  message=FALSE, warning = FALSE, fig.height=10, fig.width=10)

suppressMessages(library(tidyverse))
suppressMessages(library(ComplexHeatmap))
library(iheatmapr)
suppressMessages(library(circlize))
library(umap)
```


## Overview{.tabset .tabset-fade}
These figures are derived from the level 4 data downloaded from the LINCS Molecular Deep Dive files on Synapse https://www.synapse.org/#!Synapse:syn2862345/wiki/588244.  

The level 4 data have been filtered for QC issues and normalized to the EGF timepoints.  

```{r circosFunctions}

parseCorMatrix <- function(x){
  alpha_factor <- 255
  from_condition <- x$Condition
  to_condition <- setdiff(names(x), c("Condition", "Ligand","Time"))
  #remove metadata from correlations
  cors <- x[to_condition] %>%
    unlist()
  #remove correlations within the ligand conditions
  cors[str_detect(names(cors), x[["Ligand"]])] <- 0
  #Remove negative correlations
  cors[cors<0] <- 0
  #scale alpha based on correlations
  alpha <- (cors * alpha_factor) %% 256
  alpha <- as.integer(alpha) %>%
    as.character.hexmode()
  col <- paste0(ligand_cols[x$Ligand],alpha)
  df <- tibble(From_Condition = x$Condition,
               To_Condition =  to_condition,
               From_sector = str_remove(from_condition, "_.*"),
               To_sector = str_remove(to_condition, "_.*"),
               From_timepoint = str_remove(from_condition, ".*_"),
               To_timepoint = str_remove(to_condition, ".*_"),
               Col = col,
               Cors = as.numeric(cors)) %>%
    mutate(From_timepoint = str_replace(From_timepoint,"^4$",".5"),
           From_timepoint = str_replace(From_timepoint,"^8$","1.5"),
           From_timepoint = str_replace(From_timepoint,"^24$","2.5"),
           From_timepoint = str_replace(From_timepoint,"^48$","3.5"),
           From_timepoint = as.numeric(From_timepoint),
           To_timepoint = str_replace(To_timepoint,"^4$",".5"),
           To_timepoint = str_replace(To_timepoint,"^8$","1.5"),
           To_timepoint = str_replace(To_timepoint,"^24$","2.5"),
           To_timepoint = str_replace(To_timepoint,"^48$","3.5"),
           To_timepoint = as.numeric(To_timepoint),
           From_timepoint_st = From_timepoint - .45,
           From_timepoint_end = From_timepoint + .45,
           To_timepoint_st = To_timepoint - .45,
           To_timepoint_end = To_timepoint + .45)
}


addLink <- function(x){
  cor_thresh <- .45
  if(x["Cors"] > cor_thresh&!is.na(x["Cors"])) {
    lwd <- max(0, as.numeric(x["Cors"])-cor_thresh)*30
    #cat("lwd ",lwd, "col ", x["Col"])
    circos.link(x["From_sector"], as.numeric(x["From_timepoint"]), x["To_sector"], as.numeric(x["To_timepoint"]), h = 0.4, lwd = lwd, col =x["Col"])
  }
}

prepare_data <- function(x){
  df <- x %>%
    data.frame() %>%
    rownames_to_column(var ="Condition") %>%
    mutate(Time = str_remove(Condition, ".*_"),
           Time = str_replace_all(Time, "^4$","1"),
           Time = str_replace_all(Time, "^8$","2"),
           Time = str_replace_all(Time, "^24$","3"),
           Time = str_replace_all(Time, "^48$","4"),
           Ligand = str_remove(Condition, "_.*")
           ) %>%
    rename(Replicate = replicate)
}

draw_one_track <- function(x, title_text){
  circos.par("track.height" = 0.2,
             "points.overflow.warning" = FALSE)
  circos.initialize(factors = as.factor(x), xlim =c(0,4), sector.width = 20)

  circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
    xlim = CELL_META$xlim
    ylim = CELL_META$ylim
    breaks = seq(xlim[1], xlim[2], by = 1)
    n_breaks = length(breaks)
    circos.rect(breaks[-n_breaks], rep(ylim[1], n_breaks - 1),
                breaks[-1], rep(ylim[2], n_breaks - 1),
                col = track_col_names,
                border = NA)
    circos.text(CELL_META$xcenter,
                CELL_META$cell.ylim[2] + uy(3, "mm"),
                CELL_META$sector.index,
                col = c("black","red","black", "red", "red"))
  }) 
  
  title(title_text,
        outer = TRUE,
        cex.main = 1,
        line = -1)
  
  legend(-1.1,
         -.9,
         "TGFB, BMP2 and IFNG\n are paired with EGF",
         bty = "n",
         cex = .6)
}

plot_circos <- function(df, md, assay_name, ligand_cols, ligand_2_cols = c("EGF" = "#9E79A3", "none" = "#BE7249")) {

      ####DEBUG refactor to match new data structure######
  
           # Collection = str_remove(specimenName, "[[:alnum:]]*_{1}[[:digit:]]*_{1}"),
           # Collection = str_remove(Collection, "_.*"),
           # Replicate = str_remove(specimenName, ".*_")
           # Time = str_replace_all(Time, "^4$","1"),
           # Time = str_replace_all(Time, "^8$","2"),
           # Time = str_replace_all(Time, "^24$","3"),
           # Time = str_replace_all(Time, "^48$","4")
  #convert specimenID column names to specimanName, 
  #reorder, median summarizes replicates
  df_sum <- prep_data(df, md) %>%
    gather(specimenName, value = value, -feature) %>%
    mutate(Ligand = str_remove(specimenName, "_.*"),
           Time = str_remove(specimenName, "[[:alnum:]]*_{1}"),
           Time = str_remove(Time, "_.*")) %>%
    group_by(feature, Ligand, Time) %>%
    summarise_if(is.numeric, median, na.rm = TRUE) %>%
    ungroup() %>%
        mutate(Condition = paste(Ligand, Time, sep = "_"))%>%
  select(-Ligand, -Time)  %>%
    spread(key = Condition, value = value)
  
  df_cors <- df_sum %>%
    select(-feature) %>%
    as.matrix() %>%
    cor(use = "complete.obs") %>%
    as.data.frame %>%
    rownames_to_column("Condition") %>%
    mutate(Ligand = str_remove(Condition, "_.*"),
           Time = str_remove(Condition, ".*_")) %>%
  ####DEBUG
    select(-matches("_1|_0")) %>%
    filter(!Time %in% c("0","1"))
  ######
    
  #Assign colors to conditions in the track
  track_col_names <- c("4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black")
   Time = c( "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black")
   
  # res <- draw_one_track(unique(md$ligand), title_text = paste(assay_name, "Correlations"))

  for(ligand in unique(df_cors$Ligand)){
    res <- draw_one_track(ligand, title_text = "Correlations")
    df <-df_cors %>%
      filter(Ligand == ligand)
    for(i in 1:nrow(df)){
      foo <- parseCorMatrix(df[i,]) #
      res <- apply(foo, 1, addLink)
    }
    circos.clear()
  }
}

```



```{r standardFigureFunctions}

#' Read an assay's csv or tsv data file 
load_data <- function(assay_name, level = 4){
  file_name <- dir(pattern = paste0("MDD_", assay_name,"_","Level",level),recursive = TRUE, full.names = TRUE)
  if(!length(file_name) == 1) stop("There was not one ", assay_name, " data file found")
    
  if(str_detect(file_name, "csv")) {
    df <- read_csv(file_name)
  } else if (str_detect(file_name, "tsv")){
    df <- read_tsv(file_name)
  } else stop("The data file ",file_name, " must be a csv or tsv file type")
} #end of function

get_win_probs <- function(df, low = .02, hi = .98){
  windsorize_probs <- df %>%
    select(-feature) %>%
    unlist %>%
    quantile(probs = c(low, hi), na.rm = TRUE)
}
  
prep_data <- function(df, md){
    df <- df %>%
    gather(specimenID, value, -feature) %>%
    inner_join(md, by = "specimenID") %>%
    select(feature, specimenName, value) %>%
    spread(specimenName, value) %>%
    mutate(feature = factor(feature, levels = unique(df$feature))) %>%
      arrange(feature,)
  #Set the order of the columns
  condition_order <-  paste(rep(c("ctrl","PBS", "HGF", "OSM", "EGF","BMP2", "IFNG", "TGFB"), each = 30, times = 3), rep(c(0, 1, 4 ,8, 24, 48), each = 5, times = 3), rep(c("C1", "C2", "C3"), each = 240), rep(c("A", "B", "C", "D", "E"), times = 109),  sep = "_") %>%
    intersect(colnames(df))
  
  #Arrange columns by standard condition order
  df <- df %>%
    select(feature, condition_order)
  return(df)
}

prep_annotations <- function(df, md){
      df_ann <- df %>%
      gather("specimenName", "value", -feature) %>%
      inner_join(md, by = "specimenName") %>%
      select(ligand, secondLigand, experimentalTimePoint, replicate, collection) %>%
      distinct() %>%
      rename(Time = experimentalTimePoint,
             Ligand = ligand,
             Ligand2 = secondLigand,
             Replicate = replicate,
             Collection = collection) %>%
      as.data.frame()
      return(df_ann)
}
 #Create the heatmap column annotations
create_top_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md)
  haTop <- HeatmapAnnotation(df = df_ann,
                             col = list(Ligand = c("ctrl" = "#7A4A2A",
                                                   "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072",
                                                    "BMP2" = "#b3de69",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#ffd92f"),
                                        Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haTop)
}

prep_top_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md)
  haTop <- HeatmapAnnotation(df = df_ann,
                             col = list(Ligand = c("ctrl" = "#7A4A2A",
                                                   "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072",
                                                    "BMP2" = "#b3de69",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#ffd92f"),
                                        Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haTop)
}

create_row_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md) %>%
    select(Collection, Replicate, Time, Ligand2, Ligand)
  ####ha = rowAnnotation(foo = anno_mark(at = c(1:4, 20, 60, 97:100), labels = month.name[1:10])
  ###
  haRow <- rowAnnotation(df = df_ann, show_legend = FALSE,
                             col = list(Ligand = c( "ctrl" = "#7A4A2A",
                                                    "PBS" = "#8dd3c7",
                                                    "HGF" = "#80b1d3",
                                                    "OSM" = "#fdb462",
                                                    "EGF" = "#fb8072",
                                                    "BMP2" = "#b3de69",
                                                    "IFNG" = "#bebada",
                                                    "TGFB" = "#ffd92f"),
                                        Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                        Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haRow)
}

  
#####
plot_heatmap <- function(df, md, assay_name, EGF_normed = TRUE, var_thresh_percentile = 0, show_row_names = FALSE, mark_rows = NULL, cluster_rows = TRUE){
  
  df_vars <- df %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(df_vars, probs = var_thresh_percentile)
df <- df[df_vars > var_thresh,]

  windsorize_probs <- get_win_probs(df, .02, .98)
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md)
title_suffix <- ""
  if(EGF_normed){
    title_suffix = ", EGF timecourse normalized"
    df <- df %>%
    select(-matches("EGF"))
  }
  if(!is.null(mark_rows)) {
    row_nums <- which(df$feature %in% mark_rows)
    ra <- rowAnnotation(foo = anno_mark(at = c(row_nums), labels = df$feature[row_nums]))
  } else {
    ra <- NULL
  }
  df_as_matrix <- select(df, -feature) %>%
            as.matrix()
  rownames(df_as_matrix) <- df$feature
   #Create the heatmap
  Heatmap(matrix = df_as_matrix,
          name = "abundance",
          column_title = paste0(assay_name, title_suffix),
          top_annotation = create_top_annotations(df, md),
          right_annotation = ra,
          show_row_names = show_row_names,
          show_column_names = FALSE,
          cluster_columns = FALSE,
          cluster_rows = cluster_rows,
          col = colorRamp2(c(windsorize_probs[1], 0, windsorize_probs[2]), c("blue", "white", "red")),
          na_col = "grey")
}

plot_iheatmap <- function(df, md, assay_name, EGF_normed = TRUE, var_thresh_percentile = 0, show_row_names = FALSE, mark_rows = NULL, cluster_rows = TRUE){
  
  df_vars <- df %>%
    select(-feature) %>%
    as.matrix() %>%
    apply(1, var, na.rm = TRUE)
  
  var_thresh <- quantile(df_vars, probs = var_thresh_percentile)
  df <- df[df_vars > var_thresh,]
  
  windsorize_probs <- get_win_probs(df, .02, .98)
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md)
  
  #Create annotation values
  df_ann <-prep_annotations(df, md) %>%
    select(Collection, Replicate, Time, Ligand2, Ligand)
  
  
  title_suffix <- ""
  if(EGF_normed){
    title_suffix = ", EGF timecourse normalized"
    df <- df %>%
      select(-matches("EGF"))
  }
  if(!is.null(mark_rows)) {
    row_nums <- which(df$feature %in% mark_rows)
    ra <- rowAnnotation(foo = anno_mark(at = c(row_nums), labels = df$feature[row_nums]))
  } else {
    ra <- NULL
  }
  df_as_matrix <- select(df, -feature) %>%
    as.matrix()
  rownames(df_as_matrix) <- df$feature
  #Create the heatmap
  hm <- main_heatmap(data = df_as_matrix,
                     name = "abundance",
                     zmid = 0,
                     zmin = windsorize_probs[1],  
                     zmax = windsorize_probs[2]) %>%
    add_row_clustering() %>%
    add_col_annotation(df_ann,
                       colors = list(Ligand = c( "ctrl" = "#7A4A2A",
                                                 "PBS" = "#8dd3c7",
                                                 "HGF" = "#80b1d3",
                                                 "OSM" = "#fdb462",
                                                 "EGF" = "#fb8072",
                                                 "BMP2" = "#b3de69",
                                                 "IFNG" = "#bebada",
                                                 "TGFB" = "#ffd92f"),
                                     Time = c("0" = "white", "1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                     Ligand2 = c("EGF" = "#9E79A3", "none" = "#BE7249"),
                                     Replicate = c("A" = '#80b1d3',"B" = '#fdb462',"C" = '#ffd92f',"D" = '#8dd3c7',"E" = '#bebada'),
                                     Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")),
                       size = 0.02)
  hm
}


plot_umap <- function(df, md, ligand_cols = ligand_cols, assay_name,
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #Generate a 2d scatterplot of the UMAP dimensions
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md) %>%
    select(-matches("EGF"))
  
   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature

  df_UMAP <- umap(df_mat, na.rm = TRUE)$layout %>%
  data.frame() %>%
  rename(UMAP_1 = X1,
         UMAP_2 = X2) %>%
    cbind(df_ann) %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
p <- ggplot(df_UMAP, aes(x = UMAP_1,
                         y = UMAP_2,
                         size = Time,
                         colour = Ligand,
                         fill = Ligand2)) +
    geom_point(alpha = .8) +
  geom_point(shape = 21, stroke = 4, alpha = .8) +
  scale_color_manual( values = ligand_cols) +
  scale_fill_manual( values = ligand_2_cols) +
  scale_size(breaks = c(1,4,8,24,48)) +
  labs(title = paste("UMAP embedding of ", assay_name, " data")) +
  theme_bw()
p
  
}

plot_pca <- function(df, md, ligand_cols = ligand_cols, assay_name,
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #Generate a 2d scatterplot of the first two princpal components
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimenName and reorder
  df <- prep_data(df, md) %>%
    select(-matches("EGF"))
  
   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature

    df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann)
    
  df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann) %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
p <- ggplot(df_pca, aes(x = PC1,
                         y = PC2,
                         size = Time,
                         colour = Ligand,
                         fill = Ligand2)) +
  geom_point(alpha = .8) +
  geom_point(shape = 21, stroke = 4, alpha = .8) +
  scale_color_manual( values = ligand_cols) +
  scale_fill_manual( values = ligand_2_cols) +
  scale_size_area(max_size = 10, breaks = c(1,4,8,24,48)) +
  labs(title = paste("PCA of ", assay_name, " data")) + 
  theme_bw()
p
}


plot_correlation <- function(df, md, assay_name, EGF_normed = TRUE,
                      ligand_cols = c( "PBS" = "#8dd3c7",
                                      "HGF" = "#80b1d3",
                                      "OSM" = "#fdb462",
                                      "EGF" = "#fb8072",
                                      "BMP2" = "#b3de69",
                                      "IFNG" = "#bebada",
                                      "TGFB" = "#ffd92f"),
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #calculate correlations across the conditions and show in a heatmap

  windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimanName, reorder delete EGF samples
  df <- prep_data(df, md) 
  
  title_suffix <- ""
  if(EGF_normed){
    title_suffix <- ", EGF Normalized"
    df <- df %>%
    select(-matches("EGF"))
  }
  #Create annotation values
  #df_ann <-prep_annotations(df, md)
  
  haRow <- create_row_annotations(df, md)
  
  #Create the heatmap
  Heatmap(matrix = select(df, -feature) %>%
            as.matrix() %>%
            cor(use = "complete.obs"),
          name = "correlation",
          column_title = paste0("Correlation of ",assay_name, title_suffix),
          top_annotation = create_top_annotations(df, md),
          left_annotation = haRow,
          show_row_names = FALSE,
          show_column_names = FALSE,
          cluster_rows = FALSE,
          cluster_columns = FALSE,
          col = colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")
}


##########
```

```{r globalVariables}
ligand_cols <- c("ctrl" = "#7A4A2A",
                 "PBS" = "#8dd3c7",
                  "HGF" = "#80b1d3",
                  "OSM" = "#fdb462",
                  "EGF" = "#fb8072",
                  "BMP2" = "#b3de69",
                  "IFNG" = "#bebada",
                  "TGFB" = "#ffd92f")

md <- read_csv("metadata/MDD_sample_annotations.csv")
```

### RPPA{.tabset .tabset-fade}

#### Heatmap

```{r rppa_figures}

RPPA <- load_data("RPPA") %>%
  rename(feature = antibody)

RPPA_l1 <- load_data(assay_name = "RPPA", level = 1) %>%
  rename(feature = antibody) %>%
  gather(key = specimenID, value = value, -feature) %>%
  group_by(feature) %>%
  mutate(value = value - median(value, na.rm = TRUE)) %>%
  ungroup() %>%
    spread(key = specimenID, value = value)

```


```{r}
var_thresh_percentile <- 0

plot_heatmap(RPPA_l1, md, assay_name = "RPPA", EGF_normed = FALSE, var_thresh_percentile = var_thresh_percentile, mark_rows = c("Caspase-7-cleaved", "c-Myc",  "Connexin-43", "DUSP4", "EGFR", "EMA", "HER2", "IRF-1", "PD-L1", "Stat3_pY705"))

plot_iheatmap(RPPA_l1, md, assay_name = "RPPA", EGF_normed = FALSE, var_thresh_percentile = var_thresh_percentile)

plot_heatmap(RPPA, md, assay_name = "RPPA", var_thresh_percentile = var_thresh_percentile, mark_rows = c("Caspase-7-cleaved", "c-Myc",  "Connexin-43", "DUSP4", "EGFR", "EMA", "HER2", "IRF-1", "PD-L1", "Stat3_pY705"))

plot_iheatmap(RPPA, md, assay_name = "RPPA", var_thresh_percentile = var_thresh_percentile)

```

Antibodies with variances above the `r var_thresh_percentile*100` percentile are shown in the heatmaps.  

#### Line graphs

```{r, fig.height=30, fig.width=15}

plot_line_graphs <- function(df, md, ligand_cols, EGF_normed = TRUE, fc_thresh = 1, assay_name){
  
  windsorize_probs <- get_win_probs(df, .01, .99)
  
  df_med <- df %>%
    prep_data(md = md) %>%
    gather("specimenName", "value", -feature) %>%
    inner_join(md, by = "specimenName") %>%
    select(ligand, secondLigand, experimentalTimePoint, replicate, collection, feature, value) %>%
    filter(!ligand == "EGF") %>%
    rename(Time = experimentalTimePoint,
           Ligand = ligand,
           Ligand2 = secondLigand,
           Replicate = replicate,
           Collection = collection) %>%
    group_by(feature, Collection, Time, Ligand) %>%
    summarise(value = median(value)) %>%
    ungroup() 
  
  fc_set <- df_med %>%
    group_by(Time, Ligand) %>%
    filter(abs(value) > fc_thresh) %>%
    ungroup() %>%
    select(feature) %>%
    distinct() %>%
    inner_join(df_med, by = "feature") %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
  title_suffix <- ""
  if(EGF_normed) title_suffix <- ", EGF normalized"
  p <- ggplot(fc_set, aes(x=Time, y=value, colour=Ligand))+
    geom_line()+
    labs(title=paste0("Line Graphs",title_suffix),
         x = "Time (hours)",
         y="Intensity (AU)") +
    scale_x_continuous(breaks = c(0,8,24,48)) +
    scale_color_manual(values = ligand_cols) +
    theme(axis.text.x = element_text(angle=90),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          strip.text = element_text(size=10, hjust=0),
            panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "gray95"))+
    facet_wrap(~feature, ncol = 10,scales = "free_y")
  p
  
}
fc_thresh = 0

plot_line_graphs(df = RPPA_l1, md, ligand_cols = ligand_cols, EGF_normed = FALSE, fc_thresh = fc_thresh, assay_name = "RPPA")

plot_line_graphs(df = RPPA, md, ligand_cols = ligand_cols, fc_thresh = fc_thresh, assay_name = "RPPA")

```

The line graphs show the antibodies that have at least one response greater than `r fc_thresh` the EGF response. All values are the median summary of the biological replicates.  

#### UMAP

```{r}

plot_umap(RPPA, md, ligand_cols = ligand_cols, assay_name = "RPPA")

```

#### PCA 

```{r}

plot_pca(RPPA, md, ligand_cols = ligand_cols, assay_name = "RPPA")

```

#### Correlation

```{r}

#plot_correlation(RPPA, md, assay_name = "RPPA")

var_thresh_percentile <- .5

RPPA_l1_vars <- RPPA_l1 %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(RPPA_l1_vars, probs = var_thresh_percentile)
RPPA_l1_hi_vars <- RPPA_l1[RPPA_l1_vars > var_thresh,]

plot_correlation(RPPA_l1_hi_vars, md, EGF_normed = FALSE, assay_name = "RPPA")

RPPA_vars <- RPPA %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(RPPA_vars, probs = var_thresh_percentile)
RPPA_hi_vars <- RPPA[RPPA_vars > var_thresh,]

plot_correlation(RPPA_hi_vars, md, EGF_normed = TRUE, assay_name = "RPPA")

```


The heatmap shows the pearson correlation values of the top `r var_thresh_percentile*100`th percentile varying features.


###  

### IF{.tabset .tabset-fade}

#### Heatmap  

```{r IF_figures, eval = TRUE}

boundedLog2 <- function(x) log2(x + .0001)

immunofluoresence <- load_data(assay_name = "IF") %>%
  mutate_if(is.numeric, .funs = boundedLog2) %>%
  filter(!feature == "EdUPositiveProportion")
  
plot_heatmap(immunofluoresence, md, assay_name = "IF", mark_rows = immunofluoresence$feature)

plot_iheatmap(immunofluoresence, md, assay_name = "IF", mark_rows = immunofluoresence$feature)

```

The row features in the heatmap are `r knitr::combine_words(immunofluoresence$feature)`

#### UMAP

```{r}

plot_umap(immunofluoresence, md, ligand_cols = ligand_cols, assay_name = "IF")

```

#### PCA 

```{r}

plot_pca(immunofluoresence, md, ligand_cols = ligand_cols, assay_name = "IF")

```

#### Correlation

```{r}

plot_correlation(immunofluoresence, md, assay_name = "IF")

```

### cycIF{.tabset .tabset-fade}

#### Heatmap 

```{r cycIF_figures, eval = TRUE}

cycIF <- load_data(assay_name = "cycIF") %>%
  filter(str_detect(feature, "_int_")) %>%
  mutate_if(is.numeric, log2)

cycIF_vars <- cycIF %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh <- quantile(cycIF_vars, probs = var_thresh_percentile)
cycIF_hi_vars <- cycIF[cycIF_vars > var_thresh,]

plot_heatmap(cycIF_hi_vars, md, assay_name = "cycIF", mark_rows = c("egfr_2_af488fitc_int_mean_nuc","stat3_4_af488fitc_int_med_cytoplasm","met_6_af488fitc_int_mean_nuc"))

plot_iheatmap(cycIF_hi_vars, md, assay_name = "cycIF")

```


Intensity features included in the heatmap have a variance above the `r var_thresh_percentile` percentile . 

#### Line graphs

```{r, fig.height=15, fig.width=15}

plot_line_graphs(cycIF, md, ligand_cols = ligand_cols, assay_name = "cycIF")

```


#### UMAP

```{r}

plot_umap(cycIF, md, ligand_cols = ligand_cols, assay_name = "cycIF")

```

#### PCA 

```{r}

plot_pca(cycIF, md, ligand_cols = ligand_cols, assay_name = "cycIF")

```

#### Correlation

```{r}

plot_correlation(cycIF, md, assay_name = "cycIF")

```

### RNAseq{.tabset .tabset-fade}

#### Heatmap  

```{r rnaSeq_figures, eval = TRUE}
RNAseq <- load_data(assay_name = "RNAseq") %>%
  rename(feature = ensembl_gene_id)

RNAseq_vars <- RNAseq %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh_percentile <- .95
var_thresh <- quantile(RNAseq_vars, probs = var_thresh_percentile)
RNAseq_hi_vars <- RNAseq[RNAseq_vars > var_thresh,]
  
plot_heatmap(RNAseq_hi_vars, md, assay_name = "RNAseq")
plot_iheatmap(RNAseq_hi_vars, md, assay_name = "RNAseq")

```


Genes included in the heatmap have a variance above the `r var_thresh_percentile` percentile . 

#### UMAP

```{r}

plot_umap(RNAseq, md, ligand_cols = ligand_cols, assay_name = "RNAseq")

```

#### PCA 

```{r}

plot_pca(RNAseq, md, ligand_cols = ligand_cols, assay_name = "RNAseq")

```

#### Correlation

```{r}

plot_correlation(RNAseq_hi_vars, md, assay_name = "RNAseq")

```

### GCP{.tabset .tabset-fade}

#### Heatmap  

```{r GCP_figures}

var_thresh_percentile <- 0

GCP <- load_data(assay_name = "GCP") %>%
  rename(feature = histone)

GCP_level2 <- load_data(assay_name = "GCP", level = 2) %>%
  rename(feature = histone) %>%
  gather(key = specimenID, value = value, -feature) %>%
  group_by(feature) %>%
  mutate(value = value - median(value, na.rm = TRUE)) %>%
  ungroup() %>%
    spread(key = specimenID, value = value) %>%
  mutate(feature = factor(feature, levels = GCP$feature)) %>%
  arrange(feature)
  

plot_heatmap(GCP_level2, md, assay_name = "GCP", EGF_normed = FALSE, var_thresh_percentile = var_thresh_percentile)
plot_iheatmap(GCP_level2, md, assay_name = "GCP", EGF_normed = FALSE, var_thresh_percentile = var_thresh_percentile)

plot_heatmap(GCP_level2, md, assay_name = "GCP", EGF_normed = FALSE, var_thresh_percentile = var_thresh_percentile, show_row_names = TRUE, cluster_rows = FALSE)

plot_heatmap(GCP, md, assay_name = "GCP", var_thresh_percentile = var_thresh_percentile)
plot_iheatmap(GCP, md, assay_name = "GCP", var_thresh_percentile = var_thresh_percentile)

plot_heatmap(GCP, md, assay_name = "GCP", show_row_names = TRUE, cluster_rows = FALSE, var_thresh_percentile = var_thresh_percentile)

```


#### Line graphs

```{r, fig.height=15, fig.width=15}

plot_line_graphs(GCP, md, ligand_cols = ligand_cols, assay_name = "GCP", fc_thresh = 0)

```

#### UMAP

```{r}
df <- GCP %>%
  drop_na()
plot_umap(df, md, ligand_cols = ligand_cols, assay_name = "GCP")

```

#### PCA 

```{r}
df <- GCP %>%
  drop_na()
plot_pca(df, md, ligand_cols = ligand_cols, assay_name = "GCP")

```

#### Correlation

```{r}

var_thresh_percentile <- .5

GCP_vars <- GCP %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(GCP_vars, probs = var_thresh_percentile)
GCP_hi_vars <- GCP[GCP_vars > var_thresh,]

plot_correlation(GCP_hi_vars, md, assay_name = "GCP")

GCP_level2_vars <- GCP_level2 %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(GCP_level2_vars, probs = var_thresh_percentile)
GCP_level2_hi_vars <- GCP_level2[GCP_level2_vars > var_thresh,]

plot_correlation(GCP_level2_hi_vars, md, assay_name = "GCP")

```


The heatmap shows the pearson correlation values of the top `r var_thresh_percentile*100`th percentile varying features.  

### ATACseq{.tabset .tabset-fade}

#### Heatmap  

```{r ATACseq_figures, eval = TRUE}

ATACseq <- load_data(assay_name = "ATACseq") %>%
  rename(feature = peak)

ATACseq_vars <- ATACseq %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh_percentile <- .98
var_thresh <- quantile(ATACseq_vars, probs = var_thresh_percentile)
ATACseq_hi_vars <- ATACseq[ATACseq_vars > var_thresh,]
  
plot_heatmap(ATACseq_hi_vars, md, assay_name = "ATACseq")
plot_iheatmap(ATACseq_hi_vars, md, assay_name = "ATACseq")

```


Peaks included in the heatmap have a variance above the `r var_thresh_percentile` percentile . 


#### UMAP

```{r}

plot_umap(ATACseq, md, ligand_cols = ligand_cols, assay_name = "ATACseq")

```

#### PCA 

```{r}

plot_pca(ATACseq, md, ligand_cols = ligand_cols, assay_name = "ATACseq")

```

#### Correlation

```{r}

plot_correlation(ATACseq_hi_vars, md, assay_name = "ATACseq")

```
 

### L1000{.tabset .tabset-fade}

#### Heatmap  
 
```{r l1000_figures}

L1000 <- load_data("L1000") %>%
  rename(feature = probeset)

```


Genes included in the heatmap have a variance above the `r var_thresh_percentile` percentile . 

```{r}

var_thresh_percentile <- .5
plot_heatmap(L1000, md, "L1000",var_thresh_percentile = var_thresh_percentile)
plot_iheatmap(L1000, md, "L1000",var_thresh_percentile = var_thresh_percentile)

```

#### UMAP

```{r}

plot_umap(L1000, md, ligand_cols = ligand_cols, assay_name = "L1000")

```

#### PCA 

```{r}

plot_pca(L1000, md, ligand_cols = ligand_cols, assay_name = "L1000")

```

#### Correlation

```{r}

plot_correlation(L1000, md, assay_name = "L1000")

```
 
##