---
title: "MCF10A live cell image analysis"
output: 
  html_document:
  code_folding: hide
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(ggbeeswarm)
library(tidyverse)
library(RColorBrewer)
library(ggtree)

ligand_cols <- c("CTRL" = "#7A4A2A",
                 "PBS" = "#8dd3c7",
                 "HGF" = "#80b1d3",
                 "OSM" = "#fdb462",
                 "EGF" = "#fb8072",
                 "BMP2+EGF" = "#b3de69",
                 "IFNG+EGF" = "#bebada",
                 "TGFB+EGF" = "#ffd92f",
                  "EGF+TGFB" = "#ffd92f",
                  "OSM+EGF" = "#fdb462",
                  "OSM+EGF+TGFB" = "#ff0000",
                  "OSM+TGFB" = "#00ff00",
                  "TGFB" = "#0000ff"
                 )

create_pdfs <- FALSE

tmin_period = 30
###DEBUG
plateID <- "LI802303"

```

This report is for `r plateID`.

```{r}

df_input <- dir(paste0("Data/",plateID,"/Data"), pattern = paste0(plateID,".*_lineages.csv$"), full.names = TRUE) %>%
  map(read_csv,
      col_types = cols(
        .default = col_double(),
        D2Rpixel = col_number(),
        field = col_character(),
        well = col_character(),
        treatment = col_character()
      )) %>%
  map(tibble) %>%
  bind_rows() %>%
  mutate(key = paste(well, field, TID, sep = "_"),
         lineage_key= paste0(well, "_", field, "_", lineage),
       well_field = paste0(well, "_",field),
         thour = tmin/60)

df_orphans <-  dir(paste0("Data/",plateID,"/Data"), pattern = paste0(plateID,".*_orphans.csv$"), full.names = TRUE) %>%
  map(read_csv,
      col_types = cols(
        tmin = col_character(),
        TID = col_character(),
        xpixel = col_character(),
        ypixel = col_character(),
        PID = col_character(),
        Ival = col_character(),
        well = col_character(),
        field = col_character()
      )) %>%
  map(tibble) %>%
  bind_rows()
write_csv(df_orphans, paste0(plateID,"_orphans.csv"))

```


```{r TIDdistributions, fig.width=3, fig.height=3}

#count and display the number of cells in each lineage
df <- df_input %>%
  select(key, treatment) %>%
  distinct() 

p <- ggplot(df, aes(x = treatment)) +
  geom_bar() +
  labs(title = "Count of cells in dataset") +
  theme(axis.text.x = element_text(angle = 90))
p

#count and display the number of lineages in each lineage

df <- df_input %>%
  select(lineage, well_field, treatment) %>%
  distinct() 

p <- ggplot(df, aes(x = treatment)) +
  geom_bar() +
  labs(title = "Count of lineages in dataset") +
  theme(axis.text.x = element_text(angle = 90))
p

df <- df_input %>%
  select(key, treatment, average_velocity) %>%
  group_by(treatment) %>%
  summarise(average_velocity = mean(average_velocity, na.rm = TRUE), .groups = "drop_last")

p <- ggplot(df, aes(x = treatment, y = average_velocity)) +
  geom_col() +
  labs(title = "Average cell velocity by treatment",
       subtitle = "",
       y = "velocity(pixels/hour)") +
  theme(axis.text.x = element_text(angle = 90))
p
```

```{r beeswarm, fig.height=3, fig.width=5}
df <- df_input %>%
  select(key, treatment, average_velocity) %>%
  group_by(key, treatment) %>%
  summarise(average_velocity = mean(average_velocity, na.rm = TRUE), .groups = "drop_last") 

p <- ggplot(df, aes(x = treatment, y = average_velocity, colour = treatment)) +
  geom_quasirandom(cex = .5, alpha = .5) +
  scale_color_manual(values = ligand_cols) +
  labs(title = "Average cell velocity by treatment",
       subtitle = "",
       y = "velocity(pixels/hour)") +
  theme(axis.text.x = element_text(angle = 90))
p

```

```{r velocityTracks, fig.height=8, fig.width=12}
p <- ggplot(df_input, aes(x = thour, y = 2*distance, color = factor(TID))) +
  geom_path(size = .5, alpha = .8) +
  guides(color = FALSE)+
  labs(y = "velocity(pixels/hour)") +
  facet_wrap(~well_field+treatment) +
    theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 4,margin = margin(1)))
p

df <- df_input %>%
  filter(treatment %in% c("EGF", "OSM"))

p <- ggplot(df, aes(x = thour, y = 2*distance, color = factor(key), group = field)) +
  geom_path(size = .5, alpha = .8) +
geom_smooth(method = "loess", formula = 'y ~ x', alpha = .8, se = FALSE
 )+
  guides(color = FALSE)+
  labs(y = "velocity(pixels/hour)") +
  coord_cartesian(ylim = c(0,150)) +
  facet_wrap(~well_field+treatment) +
    theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 4,margin = margin(1)))
p

df <- df_input %>%
  filter(treatment %in% c("EGF", "OSM")) %>%
  select(treatment, average_velocity, well_field, TID) %>%
  distinct()

p <- ggplot(df, aes(x = treatment, y = average_velocity, colour = treatment)) +
  geom_quasirandom(cex = 5, alpha = .8) +
 scale_color_manual(values = ligand_cols) +
  labs(title = "Average cell velocity by treatment",
       subtitle = "",
       y = "velocity(pixels/hour)")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) 
p

EGF_vel <- df %>%
  filter(treatment == "EGF") %>%
  pull(average_velocity)

OSM_vel <- df %>%
  filter(treatment == "OSM") %>%
  pull(average_velocity)

egf_osm_ttest_obj <- t.test(x = EGF_vel, y = OSM_vel)

p <- ggplot(df, aes(x = well_field, y = average_velocity, colour = treatment)) +
  geom_quasirandom(cex = 5, alpha = .8) +
  scale_color_manual(values = ligand_cols) +
  labs(title = "Cell velocity by treatment",
       subtitle = "each dot is one cell's velocity at one time point",
       y = "velocity(pixels/hour)") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme_bw()
p


```

```{r velocity_across_time, fig.width=8, fig.height=5}

df <- df_input %>%
  group_by(treatment, thour) %>%
  summarise(average_velocity = 60*mean(distance)/tmin_period) %>%
  ungroup()

p <- ggplot(df, aes(x = thour, y = average_velocity, colour = treatment)) +
  geom_path(size = .8, alpha = .6) +
  geom_point(size = .8, alpha = .4)+
  geom_smooth(method = "loess", formula = 'y ~ x', alpha = .8, se = FALSE
  )+
  coord_cartesian(xlim = c(0,48)) +
  scale_x_continuous(breaks = c(0,12,24,36,48)) +
  labs(title = "Migration velocity over time",
       x = "time (hours)",
       y = "velocity(pixels/hour)") +
  theme_bw()
p

if(create_pdfs){
  pdf("MDD_migration_velocity_summarized.pdf", useDingbats = FALSE, width = 8, height = 5)
  print(p)
  res <- dev.off()
}

```

```{r migration_facets, fig.width=8, fig.height=7}

df <- df_input %>%
  group_by(treatment, thour, well, field) %>%
    summarise(average_velocity = 60*mean(distance)/tmin_period, .groups = "drop_last") 

p <- ggplot(df, aes(x = thour, y = average_velocity, colour = treatment, group = field)) +
  geom_path(size = .4, alpha = .8) +
  geom_point(size = .4, alpha = .4)+
  geom_smooth(method = "loess", formula = 'y ~ x', alpha = .8, se = FALSE
  )+
  coord_cartesian(xlim = c(0,48),
                  ylim = c(0,60)) +
  scale_x_continuous(breaks = c(0,12,24,36,48)) +
  labs(title = "Migration velocity over time",
       x = "time (hours)",
       y = "velocity (pixels/hour)") +
  theme_bw() +
  facet_wrap(~treatment, ncol = 4)
p

if(create_pdfs){
  pdf("MDD_migration_distance.pdf", useDingbats = FALSE, width = 8, height = 4)
  print(p)
  res <- dev.off()
}

```



```{r plot_migration, fig.width=6, fig.height=8}

cols <- RColorBrewer::brewer.pal(9, name = "Set1")[sample(c(1:5, 7:9), length(unique(df_input$TID)), replace = TRUE)]

p <- ggplot(df_input, aes(xpixel, ypixel, colour = factor(TID))) +
  geom_point(size = .1, alpha = .8) +
  geom_path(size = .1, alpha = .8) +
  scale_colour_manual(values = cols) +
  guides(colour = FALSE) +
  labs(title = "absolute positions, all data") +
  theme_bw() +
  theme(strip.text = element_text(size = 5, margin = margin()),
        strip.background = element_rect(fill = "transparent")) +
  facet_wrap(~well+field+treatment)
p

```


```{r relative_migration, fig.width=8, fig.height=8}

T0_rel <- df_input %>%
  filter(PID == 1) %>%
  select(key, xpixel, ypixel) %>%
  rename(x_T0 = xpixel,
         y_T0 = ypixel) %>%
  right_join(df_input, by = c("key") )%>%
  mutate(x_rel = xpixel-x_T0,
         y_rel = ypixel-y_T0) %>%
  select(key, lineage_key, TID, motherID, x_rel, y_rel,  well, field, treatment, lineage, generation, tmin) %>%
  drop_na()

set.seed(42)
cols <- RColorBrewer::brewer.pal(9, name = "Set1")[sample(c(1:5, 7:9), length(unique(T0_rel$TID)), replace = TRUE)]

p <- ggplot(T0_rel, aes(x_rel, y_rel, colour = factor(TID))) +
  geom_point(size = .1, alpha = .8) +
  geom_path(size = .1, alpha = .8) +
  scale_colour_manual(values = cols) +
  guides(colour = FALSE) +
    labs(title = "relative tracks",
         subtitle = "all tracks start at the origin")+
  theme_bw()+
    theme(strip.text = element_text(size = 5, margin = margin()),
        strip.background = element_rect(fill = "transparent")) +
facet_wrap(~well+field+treatment)
p
```

```{r relative_migration_linked, fig.width=8, fig.height=5}
#balance the number of lineages in each treatment
#get the number of lineages per treatment and return minimum
min_treatment_lineages <- T0_rel %>%
  select(treatment, lineage) %>%
  distinct() %>%
  count(treatment) %>%
  pull() %>%
  min()

set.seed(42)
lineages <- T0_rel %>%
  select(lineage_key, treatment) %>%
  distinct() %>%
  group_by(treatment) %>%
  sample_n(size = min_treatment_lineages, replace = FALSE) %>%
  ungroup()

#given a dataframe of one lineage
#select one track from the last time period
#if the track does not start at time 0, select it's mother
#continue this until a track starts at time 0
#return a list of the TIDs

select_lineage_tracks <- function(x, ...){
  #browser()
  last_track <- x %>%
    slice_tail(tmin, n = 1)
  first_time_period <- x %>%
    filter(TID == last_track$TID) %>%
    pull(tmin) %>%
    min()
  if(first_time_period == 0){
    selected_tracks <- x %>%
      filter(TID == last_track$TID)
  } else {
    pgen1_track <- x %>%
      filter(TID == last_track$motherID) %>%
      slice_head(tmin, n = 1)
    if(nrow(pgen1_track) ==  0)  browser()
    if(pgen1_track$tmin == 0) {
      selected_tracks <- x %>%
        filter(TID %in% c(last_track$TID, pgen1_track$TID))
    } else {
      pgen2_track <- x %>%
        filter(TID == pgen1_track$motherID) %>%
        slice_head(tmin, n = 1)
      if(pgen2_track$tmin == 0) {
        selected_tracks <- x %>%
          filter(TID %in% c(last_track$TID, pgen1_track$TID, pgen2_track$TID))
      } else {
        pgen3_track <- x %>%
          filter(TID == pgen2_track$motherID) %>%
          slice_head(tmin, n = 1)
        if(pgen3_track$tmin == 0) {
          selected_tracks <- x %>%
            filter(TID %in% c(last_track$TID, pgen1_track$TID, pgen2_track$TID, pgen3_track$TID))
        } else {
          pgen4_track <- x %>%
            filter(TID == pgen3_track$motherID) %>%
            slice_head(tmin, n = 1)
          if(pgen4_track$tmin == 0) {
            selected_tracks <- x %>%
              filter(TID %in% c(last_track$TID,
                                pgen1_track$TID,
                                pgen2_track$TID,
                                pgen3_track$TID,
                                pgen4_track$TID))
          } else {
            stop(paste("generations do not span entire time period for ", unique(x$linage_key)))
          }
        }
      }
    }
  }
  return(selected_tracks)
}

#average distance moved
calc_distance <- function(x,y){
  df <-data.frame(x,y)
  mat <- dist(df) %>%
    as.matrix()
  res <- c(0, mat[col(mat)==row(mat)+1])
  return(res)
}

#select an equal number of lineages for each treatement
#select tracks within each lineage to create one full 
#set of data across all timepoints
single_cell_lineage <- df_input %>%
  filter(lineage_key %in% lineages$lineage_key) %>%
  group_by(lineage_key) %>%
  group_modify(select_lineage_tracks) %>%
  ungroup()

#accumulate the sum of x

single_cell_lineage_T0_rel <- single_cell_lineage %>%
  filter(tmin == 0) %>%
  mutate(xpixel_T0 = xpixel,
         ypixel_T0 = ypixel) %>%
  select(lineage_key, xpixel_T0, ypixel_T0) %>%
  right_join(single_cell_lineage,by = "lineage_key") %>%
  mutate(xpixel_rel = xpixel - xpixel_T0,
         ypixel_rel = ypixel - ypixel_T0,
         treatment = factor(treatment, levels = names(ligand_cols)))  %>%
  group_by(lineage_key) %>%
  mutate(lineage_D2P = calc_distance(xpixel_rel, ypixel_rel),
         lineage_length = cumsum(lineage_D2P)) %>%
  ungroup()  %>%
  group_by(treatment, thour) %>%
  mutate(lineage_length_mean = mean(lineage_length)) %>%
  ungroup() %>%
  group_by(treatment, lineage) %>%
  mutate(lineage_col = sample(1:9, size = 1)) %>%
  ungroup()
  

mitotic_events <- single_cell_lineage_T0_rel %>%
  group_by(key) %>%
  slice_tail(tmin, n = 1) %>%
  ungroup() %>%
  filter(!tmin == max(single_cell_lineage_T0_rel$tmin))

p <- ggplot(single_cell_lineage_T0_rel, aes(x = tmin, y = lineage_key)) +
  geom_point(size = 1.2) +
  labs(title = "Lineages vs time to show missing data") +
  theme_bw()+
  theme(axis.text.y = element_text(size = 4))
p

p <- ggplot(single_cell_lineage_T0_rel, aes(xpixel_rel, ypixel_rel, colour = factor(lineage_key))) +
  geom_path(size = .4, alpha = .8) +
  scale_colour_manual(values = cols) +
  guides(colour = FALSE) +
  labs(title = "Relative motion of equal number of lineages")+
  theme_bw()+
  theme(strip.text = element_text(size = 12),
        strip.background = element_rect(fill = "transparent"),
        axis.title = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line =element_blank()) +
  facet_wrap(~treatment, ncol = 4)
p

df <- single_cell_lineage_T0_rel %>%
  filter(treatment %in% c("TGFB+EGF", "EGF", "OSM", "PBS")) 

p_subset <- ggplot(df, aes(xpixel_rel, ypixel_rel, colour = factor(lineage_key))) +
  geom_path(size = .4, alpha = .8) +
  scale_colour_manual(values = cols) +
  guides(colour = FALSE) +
  labs(title = "Relative motion of equal number of lineages")+
  theme_bw()+
  theme(strip.text = element_text(size = 12),
        strip.background = element_rect(fill = "transparent"),
        axis.title = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line =element_blank()) +
  facet_wrap(~treatment, ncol = 4)
p_subset

if(create_pdfs){
  pdf("MDD_migration_relative_tracks.pdf", useDingbats = FALSE, width = 8, height = 5)
  print(p)
  res <- dev.off()
    pdf("MDD_migration_relative_tracks_subset.pdf", useDingbats = FALSE, width = 8, height = 3)
    print(p_subset)
  res <- dev.off()
  
}

df <- single_cell_lineage_T0_rel %>%
  select(thour, lineage_length_mean, treatment) %>%
  distinct() %>%
  arrange(thour)

p <- ggplot(single_cell_lineage_T0_rel, aes(x = thour, y = lineage_length,  group = lineage_key, colour = factor(lineage_col))) +
  geom_point(data = mitotic_events, size = .6, alpha = .8) +
  geom_path(size = .6, alpha = .8) +
  scale_x_continuous(breaks = c(0,12,24, 36, 48)) +
  scale_color_brewer(palette = "Paired") +
    labs(title = "Distance traveled by lineage",
         subtitle = "equal number of lineages in each treatment, distance measured in pixels") +
  guides(color = FALSE) +
  theme_bw()+
  theme(strip.text = element_text(size = 12),
        strip.background = element_rect(fill = "transparent",linetype = "blank"),
        axis.title = element_blank(),
        axis.text = element_text(colour = "black"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line()) +
  facet_wrap(~treatment, nrow = 1)
#p <- p + 
p <- p + geom_path(data = df, aes(x = thour, y = lineage_length_mean), color = "black", group = 1, alpha = .8, size = 1)
p

if(create_pdfs){
  pdf("MDD_migration_distance.pdf", useDingbats = FALSE, width = 8, height = 2)
  print(p)
  res <- dev.off()
}

```

Method  

Distance is cumulative measured along the path of a cell.  
Velocity is calculated per time period as the distance moved in pixels per time in hours.  
Mean distance per treatment is shown by the black line.  
Lineages are a subset of a lineage tree with one daughter cell chosen after each mitotic event. The subset of the tree is chosen to span the experiment duration where possible.  
Mitotic events are shown with a vertical line.  

```{r treeExploration, eval = FALSE}
treeData <- "((t2:0.04,t1:0.34):0.89,(t5:0.37,(t4:0.03,t3:0.67):0.9):0.59);"

library("treeio")
library("ggtree")

nwk <- system.file("extdata", "sample.nwk", package="treeio")
tree <- read.tree(nwk)

treeObj <- read.tree(text = treeData)

ggplot(treeObj, aes(x, y)) + geom_tree() + theme_tree()

df_one_tree <- df_input %>%
  filter(treatment == "OSM",
         field =="A7_3",
         lineage == 1) %>%
  group_by(TID, motherID) %>%
  summarise(lifetime = max(tmin) - min(tmin)) %>%
  ungroup() %>%
  select(motherID, TID)

one_tree_obj <- as.phylo(df_one_tree)

ggplot(one_tree_obj, aes(x, y)) + 
  geom_tree() + 
  geom_tiplab()+
  theme_tree()

p <- ggtree(one_tree_obj) + geom_nodepoint(color="#b5e521", alpha=1/4, size=10)
p + geom_tippoint(color="#FDAC4F", shape=8, size=3)
p <- p + geom_tiplab(size=3, color="purple")
p


```
