---
title: "Reactome analysis of the MDD combined14 dataset"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, cache.lazy = FALSE, message=FALSE, warning = FALSE, fig.height=10, fig.width=12)

suppressMessages(library(tidyverse))
library(readxl)
library(httr)
library(jsonlite)
library(ComplexHeatmap)
library(circlize)


create_pdfs <- FALSE
write_csvs <- FALSE

#dotplot filter values 
fdr_thresh <- .1
p_value_thresh <- .05
min_gene_nbr <- 1
max_gene_nbr <- 500
gene_ratio_thresh <- .1
keep_gene_nbr <- 5
pathway_nbr <- 5

```

Read in the cluster data from the integrated analysis, filter to RNA features and write out the gene symbols.  


```{r reactome_API_pathways}

 get_pathway_data <- function(genes) {
   #browser()
   #submit file of gene names to Reactome
   res <- POST(url = "https://reactome.org/AnalysisService/identifiers/projection/?interactors=false&species=Homo%20sapiens&pageSize=200&page=1&sortBy=ENTITIES_FDR&order=ASC&resource=TOTAL&pValue=1&includeDisease=true",
               body = genes,
               accept_json(),
               content_type("text/plain"))
   
   if(!status_code(res) == 200) stop("received API code ", status_code(res))
   
   #get res token
   res_token <- fromJSON(rawToChar(res$content))$summary$token
   
     #Identify Reactome  not found genes
     found <- GET(paste0("https://reactome.org/AnalysisService/download/",res_token,"/entities/found/TOTAL/result.csv"),
                      accept("text/plain"))
     reactome_genes <- rawToChar(found$content) %>%
       str_split("\n") %>%
       unlist() %>%
       tibble() %>%
       janitor::clean_names() %>%
       slice(n =-1) %>%
       mutate(Submitted_identifier = str_remove(x, ",.*")) %>%
       pull(Submitted_identifier) %>%
       unique()
     #Check that some of the genes are in reactome
     if(!identical(reactome_genes, "")){
        #submit file of found genes to Reactome
   pathway_res <- POST(url = "https://reactome.org/AnalysisService/identifiers/projection/?interactors=false&species=Homo%20sapiens&pageSize=200&page=1&sortBy=ENTITIES_FDR&order=ASC&resource=TOTAL&pValue=1&includeDisease=true",
             body = reactome_genes,
           accept_json(),
               content_type("text/plain"))
   
   if(!status_code(pathway_res) == 200) stop("received API code ", status_code(pathway_res))
   
   pathways <- fromJSON(rawToChar(pathway_res$content))$pathways 
   
   pathway_data <- pathways %>%
     select(-entities, -reactions, -species) %>%
     bind_cols(pathways$entities) %>%
     select(-exp) %>%
     mutate(geneRatio = found/total)
     } else {
       pathway_data <- NULL
     }
  
   return(pathway_data)
 }
 
#make a list with a character vector of gene names from each module
 get_genes_from_excel <- function(sheet_name, file_name){
     genes <- read_excel(file_name, sheet = sheet_name) %>%
    filter(Type == "RNAseq") %>%
       mutate(module = sheet_name) %>%
       select(module, feature)
     if(nrow(genes) == 0) return(NULL)
     return(genes$feature)
 }
 
 #manually combine some modules based on biological review
  if(!file.exists("../tables/MDD_Reactome_pathways_combined14.csv")){
    file_name <- "../tables/MDD_int_rr_combined18_tables.xlsx"
    gene_list <- map(excel_sheets(file_name), get_genes_from_excel, file_name = file_name)
    names(gene_list) <- excel_sheets(file_name)
    gene_list[["module_3+4"]] <- c(gene_list[[3]], gene_list[[4]])
    gene_list[["module_8+17"]] <- c(gene_list[[8]], gene_list[[17]])
    gene_list[["module_10+15"]] <- c(gene_list[[10]], gene_list[[15]])
    gene_list[["module_6+16"]] <- c(gene_list[[6]], gene_list[[16]])
    gene_list[c("module_3", "module_4", "module_6","module_8", "module_10", "module_15", "module_16", "module_17")] <- NULL
          gene_list <- compact(gene_list)
   pathways_combined14 <- map_dfr(gene_list, get_pathway_data, .id = "module") %>%
         mutate(module = factor(module, ordered = TRUE, levels = gtools::mixedsort(unique(module))))
   write_csv(pathways_combined14, "../tables/MDD_Reactome_pathways_combined14.csv")
 } else {
   pathways_combined14 <- read_csv("../tables/MDD_Reactome_pathways_combined14.csv") %>%
         mutate(module = factor(module, ordered = TRUE, levels = gtools::mixedsort(unique(module))))
 }

```


```{r create_dotplots}

make_dotplots_API <- function(df, fdr_thresh = .2, gene_ratio_thresh = 0.1, pathway_nbr = 3, min_pathway_genes = 10, max_pathway_genes = 500, pathway_method = NULL, axis.text.x.angle = 0){
  #browser()
  selected_pathways <- df %>%
    filter(fdr <= fdr_thresh,
           geneRatio >= gene_ratio_thresh,
           total >= min_pathway_genes,
           total <= max_pathway_genes) %>%
        group_by(module) %>%
    arrange(fdr, desc(geneRatio), .by_group = TRUE) %>%
    slice_head(n = pathway_nbr) %>%
    ungroup() %>%
    dplyr::select(name) %>%
    distinct()

  df_selected <- df %>%
    mutate(module = str_remove(module, "module_"),
           module = factor(module, ordered = TRUE, levels = gtools::mixedsort(decreasing = TRUE,unique(module)))) %>%
    right_join(selected_pathways, by = "name") %>%
    filter(fdr <= fdr_thresh,
           geneRatio >= gene_ratio_thresh) %>%
    group_by(name) %>%
    mutate(mean_fdr = mean(fdr)) %>%
    ungroup() %>%
    arrange(module, mean_fdr) %>%
    mutate(name = str_wrap(name, 42),
           name = fct_inorder(name, ordered = TRUE))
  #df_selected$name[df_selected$name %in% silence_pathway_names] <- ""
 #Use the Reactome hierarchy to reduce the labels used in the dotplot
#get parent pathway names
reactome_pathways <- read_delim("../integrated_analysis/Data/ReactomePathways.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE) %>%
  rename(stId = X1,
         name = X2,
         species = X3) %>%
  filter(species == "Homo sapiens") %>%
  select(-species)

#Get the child to parent relationships
#Some child_stIds have more that one parent ID
#Many parents have mor ethan one child
# reactome_parents <- read_table2("../integrated_analysis/Data/ReactomePathwaysRelation.txt",
#     col_names = FALSE) %>%
#   rename(parent_stId = X1,
#          child_stId = X2) %>%
#   filter(str_detect(child_stId, "-HSA-")) %>%
#   left_join(reactome_pathways, by = c("child_stId" = "stId")) %>%
#   rename(child_name = name) %>%
#   left_join(reactome_pathways, by = c("parent_stId" = "stId")) %>%
#   rename(parent_name = name)

# %>%
#     left_join(reactome_pathways, by = c("parent_stId" = "stId")) %>%
#   rename(grandparent_name = name)

#use reactome hierarchy to determine next level up
reactome_two_levels <- read_table2("../integrated_analysis/Data/ReactomePathwaysRelation.txt",
    col_names = FALSE) %>%
  rename(parent_stId = X1,
         child_stId = X2) %>%
  filter(str_detect(child_stId, "-HSA-"))

 reactome_three_levels <- reactome_two_levels %>%
   left_join(reactome_two_levels, by = c("parent_stId" = "child_stId")) %>%
   rename(grandparent_stId = parent_stId.y) %>%
   #copy parent into grandparent if the parent was the terminal pathway
   mutate(grandparent_stId = case_when(is.na(grandparent_stId) ~ parent_stId,
                                       TRUE ~grandparent_stId)) %>%
   left_join(reactome_pathways, by = c("child_stId" = "stId")) %>%
   rename(child_name = name) %>%
   left_join(reactome_pathways, by = c("parent_stId" = "stId")) %>%
   rename(parent_name = name)   %>%
   left_join(reactome_pathways, by = c("grandparent_stId" = "stId")) %>%
   rename(grandparent_name = name)
 
df_selected_three_levels <- df_selected %>%
  left_join(reactome_three_levels, by = c("stId" = "child_stId")) %>%
  mutate(child_name = case_when(is.na(child_name) ~ as.character(name),
                                       TRUE ~child_name),
         parent_name = case_when(is.na(parent_name) ~ child_name,
                                       TRUE ~parent_name),
         grandparent_name = case_when(is.na(grandparent_name) ~ parent_name,
                                       TRUE ~grandparent_name)) %>%
  select(module, child_name, parent_name, grandparent_name, fdr, geneRatio)

res <- write_csv(df_selected_three_levels, "../tables/MDD_reactome_three_levels.csv")
p <- ggplot(df_selected, aes(x = name,
                               y = module,
                               colour = fdr,
                               size = geneRatio)) +
    geom_point() +
    scale_color_gradient(low = "red",high = "blue") +
    scale_size(range = c(1,5.5))+
    scale_x_discrete(drop = FALSE) +
    scale_y_discrete(drop = FALSE) +
    labs(title = paste(str_replace_all(pathway_method, "_|_API", " "), "dataset pathways"),
         #subtitle = paste("FDR max:", fdr_thresh, "Gene ratio min:", gene_ratio_thresh, "members min:", min_pathway_genes, "members max:", max_pathway_genes, "keep pathways:",pathway_nbr),
          x = "") +
    theme_bw() +
    theme(axis.text.y = element_text(size = 6,
                                     color = "black"),
          axis.text.x = element_text(size = 4, angle = axis.text.x.angle))
  print(p)
  
  pdf(paste0("../plots/MDD_reactome_dotplot",pathway_method,"_FDR_", fdr_thresh, "_GRmin_", gene_ratio_thresh, "_membersMin_", min_pathway_genes, "_membersMax_", max_pathway_genes, "_keep_",pathway_nbr,".pdf"),
      width = 7,
      height = 4,
      useDingbats = FALSE,)
  print(p)
  dev_off_res <- dev.off()
  
return(df_selected)
}
###end of function

pathway_method <- "_combined14_API"
fdr_thresh <- 0.15
gene_ratio_thresh <- 0.05
min_pathway_genes <- 10
max_pathway_genes <- 500
pathway_nbr <- 3
                         
dotplot_pathways <- make_dotplots_API(pathways_combined14,
                         pathway_method = pathway_method,
                         fdr_thresh = fdr_thresh,
                         gene_ratio_thresh = gene_ratio_thresh,
                         min_pathway_genes = min_pathway_genes,
                         max_pathway_genes = max_pathway_genes,
                         pathway_nbr = pathway_nbr,
                         axis.text.x.angle = 90)

```

#### Methods
The Reactome hierarchy has three levels, child, parent and grandparent. These are defined from the specific (child) up through the more all encompassing (grandparent). Parent and child are terms used in Reactome. Grandparent is not but it follows as a natural extension. Reactome also uses the term subpathway to for a lower pathway and superpathway for a higher one.    
A pathway may have 0, 1 or more than one parent pathway. In this analyis, if there is no parent pathway the child pathway is copied into the parent pathway slot. If there is no grandparent pathway, the parent pathway is copied into the grandparent slot.  

```{r reactome_pathways, eval = TRUE}

#functions
#for each pathway, return a tibble with columns of gene and stId and each row is a gene

parse_entities <- function(stId, entities){
  df <- tibble(gene = str_split(entities, ";") %>%
                 unlist(),
               stId = stId) 
}
####

# pathways_super<- dotplot_pathways %>%
#   filter(module == "9") %>%
#   mutate(superpathway = case_when(stId == "R-HSA-69190" ~"DNA Replication",
#                                   stId == "R-HSA-68962" ~"DNA Replication",
#                                   stId == "R-HSA-69205" ~"Cell Cycle",
#                                   stId == "R-HSA-176974" ~"DNA Replication",
#                                   stId == "R-HSA-69206" ~"Cell Cycle",
#                                   stId == "R-HSA-69278"  ~"Cell Cycle",
#                                   stId == "R-HSA-1538133" ~"Cell Cycle",
#                                   stId == "R-HSA-453279" ~"Cell Cycle",
#                                   stId == "R-HSA-176187" ~"Cell Cycle",
#                                   stId == "R-HSA-141444" ~"Cell Cycle",
#                                   stId == "R-HSA-141424" ~"Cell Cycle")) 
pathways_super <- read_csv("../integrated_analysis/Data/MDD_Reactome_pathways_superpathways.csv")

#Read in reactome file from 
module9_reactome_web <- read_csv("../integrated_analysis/Data/reactome_module9_noProject.csv") %>%
  janitor::clean_names() %>%
  filter(str_detect(pathway_identifier, "-HSA-"))

#filter to pathways in the dotplot
module9_dotplot_genes <- dotplot_pathways %>%
  filter(module == 9) %>%
  left_join(module9_reactome_web, by = c("stId" ="pathway_identifier")) %>%
    select(stId, submitted_entities_found)

gene_pathways <- map2(module9_dotplot_genes$stId, module9_dotplot_genes$submitted_entities_found, parse_entities) %>%
  bind_rows() %>%
  distinct() %>%
  left_join(pathways_super, by = c("stId" = "stId")) %>%
  select(gene, stId, superpathway)

res <- write_csv(gene_pathways, "../tables/MDD_integrated_reactome_module9_genes.csv")

#write out pathways for all genes in module 9
# pathway_names <- pathways_combined14 %>%
#   select(stId, name) %>%
#   distinct()  %>%
#   mutate(superpathway = case_when(stId == "R-HSA-69190" ~"DNA Replication",
#                                   stId == "R-HSA-68962" ~"DNA Replication",
#                                   stId == "R-HSA-69205" ~"Cell Cycle",
#                                   stId == "R-HSA-176974" ~"DNA Replication",
#                                   stId == "R-HSA-69206" ~"Cell Cycle",
#                                   stId == "R-HSA-69278"  ~"Cell Cycle",
#                                   stId == "R-HSA-1538133" ~"Cell Cycle",
#                                   stId == "R-HSA-453279" ~"Cell Cycle",
#                                   stId == "R-HSA-176187" ~"Cell Cycle",
#                                   stId == "R-HSA-141444" ~"Cell Cycle",
#                                   stId == "R-HSA-141424" ~"Cell Cycle")) 

gene_pathways_module9 <- map2(module9_reactome_web$pathway_identifier, module9_reactome_web$submitted_entities_found, parse_entities) %>%
  bind_rows() %>%
  distinct()  %>%
  left_join(pathways_super, by = "stId") %>%
  filter(str_detect(stId, "-HSA-")) %>%
  mutate(superpathway = case_when(is.na(superpathway) ~"not_annotated",
                                  TRUE ~superpathway))
# 
# df <- gene_pathways_module9 %>%
#   mutate(name = make.names(name)) %>%
#   select(-stId) %>%
#   distinct() %>%
#   pivot_wider(names_from = name, values_from = superpathway, values_fill = "notMember")
# res <- write_csv(df, "../tables/MDD_integrated_reactome_module9_genes_pathways.csv")

#heatmap (0 = white, 1 = red) for the attached module 9 gene list clustering both the rows and columns

gene_pathways <- read_xlsx("../integrated_analysis/Data/REACTOME_SUPERPATHWAYS_MODULE9.xlsx")

gene_pathways_dm <- gene_pathways %>%
  select(-gene) %>%
  as.matrix()
rownames(gene_pathways_dm) <- gene_pathways$gene
gp_hm <- Heatmap(gene_pathways_dm,
                  name = "Member",
                  row_names_gp = gpar(fontsize = 4),
                  column_names_gp = gpar(fontsize = 10),
                  col = colorRamp2(c(-1, 0, 1), c("#2166AC", "white", "#B2182B")))
  pdf("../plots/MDD_reactome_gene_pathways.pdf",
      width = 5,
      height =20,
      useDingbats = FALSE,)
  draw(gp_hm)
  dev_off_res <- dev.off()

```

```{r module9_patways_vs_superpathways}

#Start with module 9 pathways from Reactome
#Filter to significant pathways
#Associate pathwayd with superpathway
#plot superpathwya vs pathway, color by FDR, size by geneRatio
FDR_thresh_mod9 <- .1
gene_ratio_thresh_mod9 <- .1

# module9_pathways_selected <- module9_reactome_web %>%
#   mutate(geneRatio = number_entities_found/number_entities_total) %>%
#   filter(entities_fdr < FDR_thresh_mod9,
#          geneRatio > gene_ratio_thresh_mod9) %>%
#   arrange(entities_fdr) %>%
#     mutate(pathway_name = str_wrap(pathway_name, 32),
#            pathway_name = fct_inorder(pathway_name, ordered = TRUE)) %>%
#   left_join(pathways_super, by = c("pathway_identifier" = "stId"))

module9_pathways_selected <- module9_reactome_web %>%
  mutate(geneRatio = number_entities_found/number_entities_total) %>%
    mutate(pathway_name = str_wrap(pathway_name, 38),
           pathway_name = fct_inorder(pathway_name, ordered = TRUE)) %>%
  left_join(pathways_super, by = c("pathway_identifier" = "stId")) %>%
  group_by(superpathway) %>%
  select(-mapped_entities) %>%
  drop_na() %>%
  slice_min(order_by = c(entities_p_value), n = 5, with_ties = FALSE) %>%
  ungroup()

p <- ggplot(module9_pathways_selected, aes(x = pathway_name, y = -log10(entities_fdr), fill = superpathway)) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 20, angle = 90))
p

p <- ggplot(module9_pathways_selected, aes(x = pathway_name, y = superpathway, size = geneRatio, color = entities_fdr)) +
  geom_point() +
  scale_color_gradient(low = "red",high = "blue") +
  scale_size(range = c(1,20))+
  theme_bw() +
  theme(axis.text.x = element_text(size = 20, angle = 90),
        axis.text.y = element_text(size = 20))
p

df <- module9_pathways_selected %>%
  group_by(superpathway) %>%
  arrange(geneRatio, .by_group = TRUE) %>%
  ungroup() %>%
  mutate(pathway_name =  as.character(pathway_name),
         pathway_name = fct_inorder(pathway_name))

p <- ggplot(df, aes(x = geneRatio, y =  pathway_name, size = number_entities_found, color = entities_fdr)) +
  geom_point() +
  scale_color_gradient(low = "red",high = "blue") +
  scale_size(range = c(.8, 3))+
  theme_bw() +
  labs(title = "Module 9 Reactome Pathway Enrichment",
       x = "Gene Ratio",
       y = "",
       size = "Entities Found",
       color ="Entities FDR") +
  theme(title = element_text(size = 4),
        axis.title = element_text(size = 2),
        axis.text.x = element_text(size = 2.5, angle = 90),
        axis.text.y = element_text(size = 4, color = "black"),
        legend.title=element_text(size=4, color = "black"), 
        legend.text=element_text(size=2, color = "black"),
        legend.key.size = unit(4,"points"))
p

  pdf("../plots/MDD_reactome_module9_selected_pathways.pdf",
      width = 3.1,
      height =2.5,
      useDingbats = FALSE,)
  print(p)
  dev_off_res <- dev.off()

```

###

