---
title: "Reactome analysis of the MDD combined14 dataset"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, cache.lazy = FALSE, message=FALSE, warning = FALSE, fig.height=10, fig.width=12)

suppressMessages(library(tidyverse))
library(readxl)
library(httr)
library(jsonlite)


create_pdfs <- TRUE
write_csvs <- TRUE

#dotplot filter values 
fdr_thresh <- .1
p_value_thresh <- .05
min_gene_nbr <- 1
max_gene_nbr <- 500
gene_ratio_thresh <- .1
keep_gene_nbr <- 5
pathway_nbr <- 5

```

Read in the cluster data from the integrated analysis, filter to RNA features and write out the gene symbols.  


```{r reactome_API_pathways}

 get_pathway_data <- function(genes) {
   #browser()
   #submit file of gene names to Reactome
   res <- POST(url = "https://reactome.org/AnalysisService/identifiers/projection/?interactors=false&species=Homo%20sapiens&pageSize=200&page=1&sortBy=ENTITIES_FDR&order=ASC&resource=TOTAL&pValue=1&includeDisease=true",
               body = genes,
               accept_json(),
               content_type("text/plain"))
   
   if(!status_code(res) == 200) stop("received API code ", status_code(res))
   
   #get res token
   res_token <- fromJSON(rawToChar(res$content))$summary$token
   
     #Identify Reactome  not found genes
     found <- GET(paste0("https://reactome.org/AnalysisService/download/",res_token,"/entities/found/TOTAL/result.csv"),
                      accept("text/plain"))
     reactome_genes <- rawToChar(found$content) %>%
       str_split("\n") %>%
       unlist() %>%
       tibble() %>%
       janitor::clean_names() %>%
       slice(n =-1) %>%
       mutate(Submitted_identifier = str_remove(x, ",.*")) %>%
       pull(Submitted_identifier) %>%
       unique()
     #Check that some of the genes are in reactome
     if(!identical(reactome_genes, "")){
        #submit file of found genes to Reactome
   pathway_res <- POST(url = "https://reactome.org/AnalysisService/identifiers/projection/?interactors=false&species=Homo%20sapiens&pageSize=200&page=1&sortBy=ENTITIES_FDR&order=ASC&resource=TOTAL&pValue=1&includeDisease=true",
             body = reactome_genes,
           accept_json(),
               content_type("text/plain"))
   
   if(!status_code(pathway_res) == 200) stop("received API code ", status_code(pathway_res))
   
   pathways <- fromJSON(rawToChar(pathway_res$content))$pathways 
   
   pathway_data <- pathways %>%
     select(-entities, -reactions, -species) %>%
     bind_cols(pathways$entities) %>%
     select(-exp) %>%
     mutate(geneRatio = found/total)
     } else {
       pathway_data <- NULL
     }
  
   return(pathway_data)
 }
 
#make a list with a character vector of gene names from each module
 get_genes_from_excel <- function(sheet_name, file_name){
     genes <- read_excel(file_name, sheet = sheet_name) %>%
    filter(Type == "RNAseq") %>%
       mutate(module = sheet_name) %>%
       select(module, feature)
     if(nrow(genes) == 0) return(NULL)
     return(genes$feature)
 }
 
 # #make a list with a character vector of gene names from each module
 # get_genes_from_previous_excel <- function(sheet_name, file_name){
 #     genes <- read_excel(file_name, sheet = sheet_name) %>%
 #    filter(assay == "RNA") %>%
 #       mutate(module = sheet_name) %>%
 #       select(module, symbol)
 #     return(genes$symbol)
 # }
 # # 
 # if(!file.exists("MDD_Reactome_pathways_combined18.csv")){
 #   file_name <- "MDD_int_rr_combined18_tables.xlsx"
 #   gene_list <- map(excel_sheets(file_name), get_genes_from_excel, file_name = file_name)
 #   names(gene_list) <- excel_sheets(file_name)
 #       gene_list <- compact(gene_list)
 #   pathways_combined <- map_dfr(gene_list, get_pathway_data, .id = "module") %>%
 #         mutate(module = fct_inorder(module,ordered = TRUE))
 #   write_csv(pathways_combined, "MDD_Reactome_pathways_combined.csv")
 # } else {
 #   pathways_combined <- read_csv("MDD_Reactome_pathways_combined18.csv") %>%
 #         mutate(module = fct_inorder(module, ordered = TRUE))
 # 
 # }
 
 #manually combine some modules based on biological review
  if(!file.exists("../tables/MDD_Reactome_pathways_combined14.csv")){
    file_name <- "../tables/MDD_int_rr_combined18_tables.xlsx"
    gene_list <- map(excel_sheets(file_name), get_genes_from_excel, file_name = file_name)
    names(gene_list) <- excel_sheets(file_name)
    gene_list[["module_3+4"]] <- c(gene_list[[3]], gene_list[[4]])
    gene_list[["module_8+17"]] <- c(gene_list[[8]], gene_list[[17]])
    gene_list[["module_10+15"]] <- c(gene_list[[10]], gene_list[[15]])
    gene_list[["module_6+16"]] <- c(gene_list[[6]], gene_list[[16]])
    gene_list[c("module_3", "module_4", "module_6","module_8", "module_10", "module_15", "module_16", "module_17")] <- NULL
          gene_list <- compact(gene_list)
   pathways_combined14 <- map_dfr(gene_list, get_pathway_data, .id = "module") %>%
         mutate(module = factor(module, ordered = TRUE, levels = gtools::mixedsort(unique(module))))
   write_csv(pathways_combined14, "../tables/MDD_Reactome_pathways_combined14.csv")
 } else {
   pathways_combined14 <- read_csv("../tables/MDD_Reactome_pathways_combined14.csv") %>%
         mutate(module = factor(module, ordered = TRUE, levels = gtools::mixedsort(unique(module))))
 }
 
 #Get pathways of newly induced gene sets
   if(!file.exists("MDD_Reactome_pathways_newly_induced_genes_ligands.csv")){
    file_name <- "MDD_RNAseq_inducedGeneList_ligands.csv"
    newly_induced_genes_ligands_df <- read_csv(paste0("../integrated_analysis/Data/",file_name)) %>%
      group_by(Ligand) %>%
      select(hgnc_symbol)
    
    newly_induced_genes_ligands <- newly_induced_genes_ligands_df %>%
      group_split()
    newly_induced_genes_ligands_list <- map(newly_induced_genes_ligands, function(x){
      return(x$hgnc_symbol)
      })
    names(newly_induced_genes_ligands_list)<- newly_induced_genes_ligands_df %>%
      group_keys() %>%
      unlist()
    
   pathways_newly_induced_genes_ligands <- map_dfr(newly_induced_genes_ligands_list, get_pathway_data, .id = "module") %>%
         mutate(module = fct_inorder(module, ordered = TRUE))
   write_csv(pathways_newly_induced_genes_ligands, "../tables/MDD_Reactome_pathways_newly_induced_genes_ligands.csv")
 } else {
   pathways_newly_induced_genes_ligands <- read_csv("MDD_Reactome_pathways_newly_induced_genes_ligands.csv") %>%
         mutate(module = fct_inorder(module, ordered = TRUE))
 }
 #Pathway analaysis of newly induced genes by condition
   if(!file.exists("MDD_Reactome_pathways_newly_induced_genes_conditions.csv")){
    file_name <- "MDD_RNAseq_inducedGeneList_experimentalConditions.csv"
    newly_induced_genes_conditions_df <- read_csv(paste0("../integrated_analysis/Data/",file_name)) %>%
      group_by(experimentalCondition) %>%
      select(hgnc_symbol)
    
    newly_induced_genes_conditions <- newly_induced_genes_conditions_df %>%
      group_split()
    newly_induced_genes_conditions_list <- map(newly_induced_genes_conditions, function(x){
      return(x$hgnc_symbol)
      })
    names(newly_induced_genes_conditions_list)<- newly_induced_genes_conditions_df %>%
      group_keys() %>%
      unlist()
    
   pathways_newly_induced_genes_conditions <- map_dfr(newly_induced_genes_conditions_list, get_pathway_data, .id = "module") %>%
         mutate(module = fct_inorder(module, ordered = TRUE))
   write_csv(pathways_newly_induced_genes_conditions, "../tables/MDD_Reactome_pathways_newly_induced_genes_conditions.csv")
 } else {
   pathways_newly_induced_genes_conditions <- read_csv("MDD_Reactome_pathways_newly_induced_genes_conditions.csv") %>%
         mutate(module = fct_inorder(module, ordered = TRUE))
 }
```


```{r create_dotplots}

make_dotplots_API <- function(df, fdr_thresh = .2, gene_ratio_thresh = 0.1, pathway_nbr = 3, min_pathway_genes = 10, max_pathway_genes = 500, pathway_method = NULL, axis.text.x.angle = 0){
  #browser()
  selected_pathways <- df %>%
    group_by(module) %>%
    filter(fdr <= fdr_thresh,
           geneRatio >= gene_ratio_thresh,
           total >= min_pathway_genes,
           total <= max_pathway_genes) %>%
    slice_min(order_by = c(fdr), n = pathway_nbr, with_ties = TRUE) %>%
    slice_max(geneRatio, n = pathway_nbr, with_ties = TRUE) %>%
    ungroup() %>%
    dplyr::select(name) %>%
    distinct()

  df_selected <- df %>%
    mutate(module = str_remove(module, "module_"),
           module = factor(module, ordered = TRUE, levels = gtools::mixedsort(decreasing = TRUE,unique(module)))) %>%
    right_join(selected_pathways, by = "name") %>%
    filter(fdr <= fdr_thresh,
           geneRatio >= gene_ratio_thresh) %>%
    group_by(name) %>%
    mutate(mean_fdr = mean(fdr)) %>%
    ungroup() %>%
    arrange(module, mean_fdr) %>%
    mutate(name = str_wrap(name, 32),
           name = fct_inorder(name, ordered = TRUE))
  #df_selected$name[df_selected$name %in% silence_pathway_names] <- ""
 #Use the Reactome hierarchy to reduce the labels used in the dotplot
#get parent pathway names
reactome_pathways <- read_delim("../integrated_analysis/Data/ReactomePathways.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE) %>%
  rename(stId = X1,
         name = X2,
         species = X3) %>%
  filter(species == "Homo sapiens") %>%
  select(-species)

#Get the child to parent relationships
#Some child_stIds have more that one parent ID
#Many parents have mor ethan one child
# reactome_parents <- read_table2("../integrated_analysis/Data/ReactomePathwaysRelation.txt",
#     col_names = FALSE) %>%
#   rename(parent_stId = X1,
#          child_stId = X2) %>%
#   filter(str_detect(child_stId, "-HSA-")) %>%
#   left_join(reactome_pathways, by = c("child_stId" = "stId")) %>%
#   rename(child_name = name) %>%
#   left_join(reactome_pathways, by = c("parent_stId" = "stId")) %>%
#   rename(parent_name = name)

# %>%
#     left_join(reactome_pathways, by = c("parent_stId" = "stId")) %>%
#   rename(grandparent_name = name)

#use reactome hierarchy to determine next level up
reactome_two_levels <- read_table2("../integrated_analysis/Data/ReactomePathwaysRelation.txt",
    col_names = FALSE) %>%
  rename(parent_stId = X1,
         child_stId = X2) %>%
  filter(str_detect(child_stId, "-HSA-"))

 reactome_three_levels <- reactome_two_levels %>%
   left_join(reactome_two_levels, by = c("parent_stId" = "child_stId")) %>%
   rename(grandparent_stId = parent_stId.y) %>%
   #copy parent into grandparent if the parent was the terminal pathway
   mutate(grandparent_stId = case_when(is.na(grandparent_stId) ~ parent_stId,
                                       TRUE ~grandparent_stId)) %>%
   left_join(reactome_pathways, by = c("child_stId" = "stId")) %>%
   rename(child_name = name) %>%
   left_join(reactome_pathways, by = c("parent_stId" = "stId")) %>%
   rename(parent_name = name)   %>%
   left_join(reactome_pathways, by = c("grandparent_stId" = "stId")) %>%
   rename(grandparent_name = name)
 
df_selected_three_levels <- df_selected %>%
  left_join(reactome_three_levels, by = c("stId" = "child_stId")) %>%
  mutate(child_name = case_when(is.na(child_name) ~ as.character(name),
                                       TRUE ~child_name),
         parent_name = case_when(is.na(parent_name) ~ child_name,
                                       TRUE ~parent_name),
         grandparent_name = case_when(is.na(grandparent_name) ~ parent_name,
                                       TRUE ~grandparent_name)) %>%
  select(module, child_name, parent_name, grandparent_name, fdr, geneRatio)

res <- write_csv(df_selected_three_levels, "../tables/MDD_reactome_three_levels.csv")
p <- ggplot(df_selected, aes(x = name,
                               y = module,
                               colour = fdr,
                               size = geneRatio)) +
    geom_point() +
    scale_color_gradient(low = "red",high = "blue") +
    scale_size(range = c(0,4))+
    scale_x_discrete(drop = FALSE) +
    scale_y_discrete(drop = FALSE) +
    labs(title = paste(str_replace_all(pathway_method, "_|_API", " "), "dataset pathways"),
          x = "") +
    theme_bw() +
    theme(axis.text.y = element_text(size = 6,
                                     color = "black"),
          axis.text.x = element_text(size = 4, angle = axis.text.x.angle))
  print(p)
  
  
  
  pdf(paste0("../plots/MDD_reactome_dotplot",pathway_method,".pdf"),
      width = 8,
      useDingbats = FALSE,)
  print(p)
  dev_off_res <- dev.off()
}

res <- make_dotplots_API(pathways_combined14, pathway_method = "_combined14_API",fdr_thresh = 0.15, gene_ratio_thresh = 0.1, axis.text.x.angle = 90, pathway_nbr = 5)

res <- make_dotplots_API(pathways_newly_induced_genes_ligands, pathway_method = "_newly_induced_genes_ligands_API",fdr_thresh = 0.1, gene_ratio_thresh = 0.05, axis.text.x.angle = 90)

res <- make_dotplots_API(pathways_newly_induced_genes_conditions, pathway_method = "_newly_induced_genes_conditions_API",fdr_thresh = 0.1, gene_ratio_thresh = 0.05, axis.text.x.angle = 90)

####
# Create bubble plot of Reactome enrichments for combined14 modules with the following filters (Mark: let's brainstorm this a bit):
# Remove gene sets with > 500 features or < 15 features
# Require FDR < 0.1
# Require GeneRatio > 0.1
# Plot: 1) up to the top 5 pathways for each module that meet these thresholds; 2) up to the top 3 pathways for each module that meet these threshold

####
# 
###
#   #####NULL dataset analysis
#   
#   selected_pathways_null <- enriched_pathways_null %>%
#   group_by(module) %>%
#   slice_min(order_by = c(entities_p_value), n = keep_gene_nbr) %>%
#   slice_max(geneRatio, n = keep_gene_nbr) %>%
#   ungroup() %>%
#   dplyr::select(pathway_identifier)
# 
# df <- enriched_pathways_null %>%
#   right_join(selected_pathways_null) %>%
#   filter(entities_p_value <= p_value_thresh,
#          geneRatio > gene_ratio_thresh) %>%
#   group_by(pathway_name) %>%
#   mutate(mean_fdr = mean(entities_fdr),
#          mean_p_value = mean(entities_p_value)) %>%
#   ungroup() %>%
#   arrange(desc(module), mean_p_value) %>%
#   mutate(pathway_name = fct_inorder(pathway_name))
# 
# p <- ggplot(df, aes(x = module,
#                     y = pathway_name,
#                     colour = entities_p_value,
#                     size = geneRatio)) +
#   geom_point() +
#   scale_x_discrete(drop = FALSE) +
#   scale_color_gradient(low = "red",high = "blue") +
#   scale_size(range = c(0,4))+
#   labs(title = "Null dataset") +
#   theme_bw() +
#   theme(axis.text.y = element_text(size = 6,
#                                    color = "black"))
# p
# 
#  pdf(paste0("MDD_null_dataset_reactome_dotplot.pdf"),
#      width = 8,
#      useDingbats = FALSE)
#   print(p)
#   dev_off_res <- dev.off()

foo <- tibble(a = factor(c(1,3,4), levels = 1:4, ordered = TRUE),
              b = factor(letters[c(2,3,5)], levels = letters[1:6], ordered = TRUE),
              c = 1:3)

p <-  ggplot(foo, aes(x = a, y = b, color = c, size = c)) +
  geom_point() +
  scale_y_discrete(drop = FALSE)
p

```

####Methods
The Reactome hierarchy has three levels, child, parent and grandparent. These are defined from the specific (child) up through the more all encompassing (grandparent). Parent and child are terms used in Reactome. Grandparent is not but it follows as a natural extension.  
A pathway may have 0, 1 or more than one parent pathway. In this analyis, if there is no parent pathway the child pathway is copied into the parent pathway slot. If there is no grandparent pathway, the parent pathway is copied into the grandparent slot.  

###

