---
title: "MDD integrated analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, cache.lazy = FALSE, message=FALSE, warning = FALSE, fig.height=10, fig.width=14)

suppressMessages(library(tidyverse))
library(ComplexHeatmap)
library(circlize)
library(umap)
library(cluster)
library(RColorBrewer)
library(writexl)

source("../R/MDD_functions.R")
create_pdfs <- TRUE
write_csvs <- TRUE

```



```{r setup_parameters_and_metadata}
keep_PBS <- FALSE
PBS_pointer <- "_noPBS2"
if(keep_PBS) PBS_pointer <- ""

data_path <- paste0("../integrated_analysis/integrated_matrix_lfc_rr",PBS_pointer, ".csv")
p_values_path <- paste0("../integrated_analysis/integrated_adj_p_values", PBS_pointer, ".csv") 
shared_features_path <- paste0("../integrated_analysis/integrated_shared_features_rr", PBS_pointer, ".csv") 
unique_features_path <- paste0("../integrated_analysis/integrated_unique_features_rr", PBS_pointer, ".csv") 

clust_num <- 10
cluster_method <- "pam"

experimentalTimePoints <- c("_24", "_48")
ligand_order <-  c("PBS", "HGF", "OSM", "EGF","BMP2+EGF", "IFNG+EGF", "TGFB+EGF")
ligand_order <-  c("HGF", "OSM", "EGF","BMP2+EGF", "IFNG+EGF", "TGFB+EGF")
condition_order <- paste0(rep(ligand_order, each = length(experimentalTimePoints)), rep(c("_24", "_48"), times = length(ligand_order)))

#create a named vector of color values for the Heatmap clusters
cluster_cols <- c(structure(RColorBrewer::brewer.pal(12, "Paired")[c(1:12)], names = 1:12), c("13" = "black", "14" = "gray"))

ligand_cols <- c("CTRL" = "#7A4A2A",
                 "PBS" = "#8dd3c7",
                 "HGF" = "#80b1d3",
                 "OSM" = "#fdb462",
                 "EGF" = "#fb8072",
                 "BMP2+EGF" = "#b3de69",
                 "IFNG+EGF" = "#bebada",
                 "TGFB+EGF" = "#ffd92f")

#create factors for Up/Down unique features
up_down_cluster_levels <- paste(names(ligand_cols), rep(c("Negative", "Mixed", "Positive"), each = length(ligand_cols)), sep = "_")
 
up_down_cluster_colors <- rep(ligand_cols, times = 3)
names(up_down_cluster_colors) <- up_down_cluster_levels

```

This analysis is run with keep_PBS flag set to `r keep_PBS`.  
The cluster method is `r cluster_method`.  
Input files are `r data_path`, `r shared_features_path` and `r unique_features_path`.  


```{r prepData}

#Read in the integrated data, add EGF to some names, set order
if(!file.exists(data_path)) stop("cannot read ",data_path)
all_features_data <- read_csv(data_path) %>%
  pivot_longer(cols = matches("_24|_48"), names_to = "condition") %>%
  mutate(condition = str_replace(condition, "BMP2","BMP2+EGF"),
         condition = str_replace(condition, "IFNG","IFNG+EGF"),
         condition = str_replace(condition, "TGFB","TGFB+EGF"),
         feature_type =  paste(feature, Type, sep = "_")) %>%
  pivot_wider(names_from = condition, values_from = value) %>%
  select(feature, Type, feature_type, all_of(condition_order)) %>%
  drop_na

#Read in the integrated data, add EGF to some names, set order
if(!file.exists(unique_features_path)) stop("cannot read ",unique_features_path)
unique_features <- read_csv(unique_features_path)
unique_features_data <- unique_features%>%
  mutate(ligand = str_replace(ligand, "BMP2","BMP2+EGF"),
         ligand = str_replace(ligand, "IFNG","IFNG+EGF"),
         ligand = str_replace(ligand, "TGFB","TGFB+EGF"),
         feature_type =  paste(feature, Type, sep = "_")) %>%
  select(ligand, feature_type, Direction) %>%
  left_join(all_features_data, by = "feature_type") %>%
  rename(Cluster = ligand) %>%
  mutate(Cluster = paste(Cluster, Direction, sep = "_")) %>%
  drop_na

#Read in the integrated data, add EGF to some names, set order
if(!file.exists(shared_features_path)) stop("cannot read ",shared_features_path)
shared_features <- read_csv(shared_features_path)
shared_features_data <- shared_features %>%
  mutate(ligand = str_replace(ligand, "BMP2","BMP2+EGF"),
         ligand = str_replace(ligand, "IFNG","IFNG+EGF"),
         ligand = str_replace(ligand, "TGFB","TGFB+EGF"),
         feature_type =  paste(feature, Type, sep = "_")) %>%
  select(feature_type) %>%
  distinct() %>%
  left_join(all_features_data, by = c("feature_type" ))  %>%
  drop_na

set_features_data <- bind_rows(shared_features_data, select(unique_features_data, -Direction, -Cluster))
assay_cols <- structure(RColorBrewer::brewer.pal(8, "Paired")[c(6, 2:3,5, 1)], names = unique(set_features_data$Type))

```
## {.tabset .tabset-fade}

### heatmap

```{r heatmaps}

#create a numeric matrix with feature row names
set_features_dm <- set_features_data %>%
  select(all_of(condition_order)) %>%
  as.matrix() 
rownames(set_features_dm) <- set_features_data$feature_type
set_features_dm <- set_features_dm[!is.nan(set_features_dm[,1]),]

#cluster all features 
if(cluster_method == "kmeans"){
#seed the clustering with a fixed, random set of points
set.seed(42)
centers <- set_features_dm[sample(1:nrow(set_features_dm), size = clust_num, replace = FALSE),]
clusters <- kmeans(x = set_features_dm, iter.max = 20, centers = centers)$cluster
} else if(cluster_method == "HA") {
  clust_obj <- hclust(dist(set_features_dm))
  clusters <- cutree(clust_obj, k = clust_num)
} else if(cluster_method == "pam") {
  set.seed(42)
medoids <- sample(1:nrow(set_features_dm), size = clust_num, replace = FALSE)
  clusters  <- pam(x = set_features_dm, k = clust_num, medoids = medoids, cluster.only=TRUE)
} else stop("unsupported cluster method chosen")

#Add the cluster assignements to the data
set_features_data_ann <- set_features_ann <- tibble(feature_type = names(clusters),
                           Cluster = clusters) %>%
  right_join(set_features_data, by = "feature_type") %>%
    mutate(Cluster = factor(Cluster, ordered = TRUE))

set_features_ann <- set_features_data_ann %>%
  select(Cluster, Type) %>%
  as.data.frame()

haRow_set <- HeatmapAnnotation(df = set_features_ann,
                               which = "row",
                               col = list(Cluster = factor(cluster_cols, levels = cluster_cols, ordered = TRUE),
                                          Type = assay_cols))

hm_set <- Heatmap(set_features_dm,
               name = "lfc rr",
               column_title = "Set features",
               column_title_gp = gpar(fontsize = 12),
               col = colorRamp2(c(-2, 0, 2), c("#2166AC", "white", "#B2182B")),
               cluster_rows = FALSE,
               cluster_row_slices = FALSE, 
               cluster_columns = FALSE,
               row_split = set_features_ann$Cluster,
               row_gap = unit(2, "mm"),
               row_title = " ",
               show_row_names = FALSE,
               row_names_gp = gpar(fontsize = 8),
               column_names_gp = gpar(fontsize = 10),
               left_annotation = haRow_set,
               use_raster = FALSE)
hm_set

if(create_pdfs){
  pdf("MDD_integrated_rr_set_heatmap.pdf") 
  draw(hm_set)
  res <- dev.off()
} 
#######
#Shared features

#Create data and metadata matrices and annotations
#cluster and order by cluster value

shared_features_dm <- shared_features_data %>%
  select(all_of(condition_order)) %>%
  as.matrix() 

rownames(shared_features_dm) <- shared_features_data$feature_type
shared_features_dm <- shared_features_dm[!is.nan(shared_features_dm[,1]),]

#cluster shared matrix
if(cluster_method == "kmeans"){
  #seed the kmeans clustering with a fixed, random set of points
set.seed(42)
centers <- shared_features_dm[sample(1:nrow(shared_features_dm), size = clust_num, replace = FALSE),]
clusters <- kmeans(x = shared_features_dm, iter.max = 20, centers = centers)$cluster
} else if(cluster_method == "HA") {
  clust_obj <- hclust(dist(shared_features_dm))
  clusters <- cutree(clust_obj, k = clust_num)
} else if(cluster_method == "pam") {
   set.seed(42)
medoids <- sample(1:nrow(shared_features_dm), size = clust_num, replace = FALSE)
  clusters  <- pam(x = shared_features_dm, k = clust_num, medoids = medoids, cluster.only=TRUE)
} else stop("unsupported cluster method chosen")

shared_features_data_ann <-tibble(feature_type = names(clusters),
                 Cluster = clusters) %>%
  right_join(shared_features_data, by = "feature_type") %>%
  mutate(Cluster = factor(Cluster, levels = 1:clust_num, ordered = TRUE))

shared_features_ann <-shared_features_data_ann %>%
    select(Cluster, Type) %>%
  as.data.frame()

haRow_shared <- HeatmapAnnotation(df = shared_features_ann,
                           which = "row",
                           col = list(Cluster = factor(cluster_cols, levels = cluster_cols, ordered = TRUE),
                                      Type = assay_cols))

hm_shared <- Heatmap(shared_features_dm,
               name = "lfc rr",
               column_title = "Shared features",
               column_title_gp = gpar(fontsize = 12),
               col = colorRamp2(c(-2, 0, 2), c("#2166AC", "white", "#B2182B")),
               cluster_rows = FALSE,
               cluster_row_slices = TRUE, 
               cluster_columns = FALSE,
               row_split = shared_features_ann$Cluster,
               row_gap = unit(2, "mm"),
               row_title = " ",
               show_row_names = FALSE,
               row_names_gp = gpar(fontsize = 8),
               column_names_gp = gpar(fontsize = 10),
               left_annotation = haRow_shared,
               use_raster = FALSE)
hm_shared
if(create_pdfs){
  pdf("MDD_integrated_rr_shared_heatmap.pdf") 
  draw(hm_shared)
  res <- dev.off()
} 

# #Create a matrix and heatmap of the mean cluster values by condition
 shared_features_data_mean <- shared_features_data_ann %>%
  group_by(Cluster) %>%
  summarise(across(.cols = matches("_24|48"), .fns = mean))

shared_features_data_mean_dm <- shared_features_data_mean %>%
  select(-Cluster) %>%
  as.matrix()
rownames(shared_features_data_mean_dm) <- paste0("Module_",shared_features_data_mean$Cluster)

hm_mean <- Heatmap(shared_features_data_mean_dm,
               name = "lfc ",
               column_title = "Shared cluster averages",
               column_title_gp = gpar(fontsize = 12),
               col = colorRamp2(c(-2, 0, 2), c("#2166AC", "white", "#B2182B")),
               cluster_rows = FALSE,
               cluster_row_slices = FALSE,
               cluster_columns = FALSE,
               row_gap = unit(2, "mm"),
               row_title = " ",
               show_row_names = TRUE,
               row_names_gp = gpar(fontsize = 8),
               column_names_gp = gpar(fontsize = 10),
               use_raster = FALSE)

#make a heatmap of the unique features
#Create data and metadata matrices and annotations
#k means cluster and order by cluster value

unique_features_data_ann <- unique_features_data %>%
  mutate(Ligand = str_remove(Cluster, "_.*"),
         Ligand = factor(Ligand,  ordered = TRUE),
         Direction = factor(Direction, ordered = TRUE),
    Cluster = factor(Cluster, ordered = TRUE)) 

unique_features_ann <- unique_features_data_ann %>%
    select(Ligand, Direction, Type) %>%
  as.data.frame()

haRow_unique <- HeatmapAnnotation(df = unique_features_ann,
                                  which = "row",
                                  col = list(Ligand = factor(ligand_cols, levels = ligand_cols, ordered = TRUE),
                                             Type = assay_cols,
                                             Direction = factor(c("Negative" = "blue","Mixed" = "gray","Positive" = "red"))))

unique_features_data_dm<- unique_features_data_ann %>%
    select(all_of(condition_order)) %>%
  as.matrix()

hm_unique <- Heatmap(unique_features_data_dm,
               name = "lfc rr",
               column_title = "Ligand-unique features",
               column_title_gp = gpar(fontsize = 12),
               col = colorRamp2(c(-2, 0, 2), c("#2166AC", "white", "#B2182B")),
               cluster_rows = FALSE,
               cluster_row_slices = TRUE, 
               cluster_columns = FALSE,
               row_split = unique_features_data_ann$Cluster,
               row_gap = unit(2, "mm"),
               row_title = " ",
               show_row_names = FALSE,
               row_names_gp = gpar(fontsize = 8),
               column_names_gp = gpar(fontsize = 10),
               left_annotation = haRow_unique,
               use_raster = FALSE)
hm_unique

#hm_mean

if(create_pdfs){
  pdf("MDD_integrated_rr_unique_heatmap.pdf") 
  draw(hm_unique)
  res <- dev.off()
} 

# #create a mean matrix of the unique features
#  integrated_unique_mean <- integrated_unique %>%
#   group_by(Cluster) %>%
#   summarise(across(.cols = matches("_24|48"), .fns = mean))
# 
# dm_unique_mean <- integrated_unique_mean %>%
#   select(-Cluster) %>%
#   as.matrix()
# rownames(dm_unique_mean) <- paste0("Module_",integrated_unique_mean$Cluster)

```

```{r writeTables}
selectFeature <- function(df){
  #browser()
  res <- df %>%
    select(feature, Type) %>%
    mutate(analyte = str_remove(feature, "_.*")) %>%
    rename(assay = Type) %>%
    select(feature, analyte, assay, time, symbol)
return(res)
}

selectDerivedFeature <- function(df){
  #browser()
  res <- df %>%
    select(feature, FDR, Type)
return(res)
}

module_features <- shared_features_data_ann %>%
  arrange(Cluster, Type) %>%
  mutate(Cluster = paste0("module_", Cluster),
         Cluster = factor(Cluster, levels = unique(Cluster))) %>%
  split(.$Cluster) 

res <- write_xlsx(module_features,
                  path = "MDD_int_rr_shared_tables.xlsx")

module_features <- unique_features_data %>%
  arrange(Cluster, Type) %>%
  mutate(Cluster = paste0("module_", Cluster),
         Cluster = factor(Cluster, levels = unique(Cluster))) %>%
  split(.$Cluster) 

res <- write_xlsx(module_features,
                  path = "MDD_int_rr_unique_tables.xlsx")
```


### UMAPs

```{r UMAPAnalysis, fig.width=5.75, fig.height=4.75}

#Generate a 2d scatterplot of the UMAP dimensions and color by clusters, assay type

#Create annotation values
dm <- shared_features_data_ann %>%
  dplyr::select(-feature, -Cluster, -Type, -feature_type) %>%
  as.matrix()
rownames(dm) <- shared_features_data_ann$feature_type

custom_settings = umap.defaults
custom_settings$random_state <- 42
custom_settings$n_epochs = 200
custom_settings$n_neighbors = 15

if(file.exists(paste0("../integrated_analysis/MDD_",cluster_method,"_UMAP.rda"))){
  load(paste0("../integrated_analysis/MDD_",cluster_method,"_UMAP.rda"))
} else {
df_UMAP <- umap(dm, config = custom_settings, na.rm = TRUE)$layout %>%
  data.frame(feature_type = rownames(dm)) %>%
  rename(UMAP_1 = X1,
         UMAP_2 = X2) %>%
  left_join(shared_features_data_ann, by = "feature_type") 
save(df_UMAP, file = paste0("../integrated_analysis/MDD_",cluster_method,"_UMAP.rda"))
}

plot_UMAP_3d <- FALSE
if(plot_UMAP_3d){
  custom_settings$n_components = 3
  
  df_UMAP_3d <-  umap(dm, config = custom_settings, na.rm = TRUE)$layout %>%
    data.frame(feature = rownames(dm)) %>%
    rename(UMAP_1 = X1,
           UMAP_2 = X2,
           UMAP_3 = X3) %>%
    left_join(integrated_shared, by = "feature")
  df_UMAP_3d$Cluster_color <- cluster_cols[df_UMAP_3d$Cluster]
  save(df_UMAP_3d, file = "MDD_UMAP_3d")
  
  library(rgl)
  plot3d(df_UMAP_3d$UMAP_1,
         df_UMAP_3d$UMAP_2,
         df_UMAP_3d$UMAP_3,
         col = factor(df_UMAP_3d$Cluster_color))
}

  p <- ggplot(df_UMAP, aes(x = UMAP_1,
                           y = UMAP_2,
                           colour = factor(Cluster))) +
    geom_point(size = 1, alpha = .6, shape = 16) +
    scale_color_manual(values = cluster_cols) +
    labs(title = paste("UMAP embedding colored by cluster"),
         colour = "Cluster") +
    guides(colour = guide_legend(override.aes = list(alpha = 1, size = 4))) +
    theme_bw()+
    theme(axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line =element_blank())
  p
  
  if(create_pdfs) {
    pdf("MDD_conditions_umap_main.pdf")    
    print(p)
    res <- dev.off()
  } 
```

There are `r dim(dm)[1]` dots in the 2d UMAP that represent the significant differentially expressed features relative to T0 in the shared-features integrated matrix.  


```{r UMAPAnalysis2, fig.width=7.5, fig.height=4}
p <- ggplot(data = df_UMAP[df_UMAP$Type=="RNA",], aes(x = UMAP_1,
                                                      y = UMAP_2,
                                                      colour = factor(Type))) +
  geom_point(size = .8, alpha = .6) +
  scale_colour_manual(values = assay_cols)+
  labs(title = paste("UMAP embedding colored by assay"),
       colour = "Assay") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 4))) +
  theme_bw()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line =element_blank())
p <- p + geom_point(data = df_UMAP[!df_UMAP$Type=="RNA",],size = .8, alpha = .8)
p

if(create_pdfs) {
  pdf("MDD_ligands_umap_type.pdf") 
      print(p)
  res <- dev.off()
}

```

```{r highlightAssay, fig.width=2.5, fig.height=2.5}

p_list <- lapply(unique(df_UMAP$Type), function(assay){
  
  df <- df_UMAP
  
  p <- ggplot(data = df[!df$Type==assay,], 
              aes(x = UMAP_1,
                  y = UMAP_2,
                  color = Type)) +
    geom_point(color = "gray",size = 1, alpha = .4, shape = 16) +
    scale_colour_manual(values = assay_cols) +
    annotate("text",x = -2, y = -6, label = assay) +
    scale_size_area(limits = c(0, 30)) +
    labs()+
    guides(colour = FALSE,
           size = FALSE)+
    theme_bw()+
    theme(axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line =element_blank())
  p <- p + geom_point(data = df[df$Type==assay,], 
                      size = 1,
                      alpha = .9,
                      shape = 16) 
  return(p)
  
})

res <- map(p_list, print)

if(create_pdfs) {
  pdf("MDD_ligands_umap_type_highlights.pdf") 
  res <- map(p_list, print)
  res <- dev.off()
}
```


```{r UMAPbyCondition, fig.width=2.5, fig.height=2.5, eval = TRUE}


df <- df_UMAP %>%
  rename("BMP2_EGF_24" = "BMP2+EGF_24",
         "IFNG_EGF_24" =  "IFNG+EGF_24",
         "TGFB_EGF_24" = "TGFB+EGF_24",
         "BMP2_EGF_48" = "BMP2+EGF_48",
         "IFNG_EGF_48" =  "IFNG+EGF_48",
         "TGFB_EGF_48" = "TGFB+EGF_48")
df_UMAP_conditions <- colnames(df)[str_detect(colnames(df_UMAP),"PBS|HGF|OSM|EGF|BMP2|IFNG|TGFB")] %>%
  str_replace("[+]","_")

p_list <- lapply(df_UMAP_conditions, function(condition){
  p <- ggplot(df, aes_string(x = "UMAP_1",
                             y = "UMAP_2",
                             colour = condition)) +
    geom_point(size = 1.5, alpha = .8, shape = 16) +
    annotate("text",x = -2, y = -6, label = condition)+
    scale_color_gradientn(colours = colorRampPalette(rev(brewer.pal(n = 7, name =
                                                                      "RdBu")))(100)) +
    labs()+
    guides(colour = FALSE)+
    theme_bw()+
    theme(text = element_text(size = 1),
          axis.title = element_blank(),
          panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(colour = "black"))
  return(p)
})
res <- map(p_list, print)
if(create_pdfs) {
  pdf("MDD_integrated_umap_by_condition.pdf")
  res <- map(p_list, print)
  res <- dev.off()
}

```

### phenotype analysis

```{r correlate_to_phenotypes}

level3File <- paste0("../IF/Data/MDD_IF_Level3.csv")

raw_df <- read_csv(level3File)
colnames(raw_df) <- raw_df[raw_df$X1 == "specimenName",]

if_df <- raw_df %>%
  select(matches("specimenName|24_C2|48_C2")) %>%
  rename(feature = specimenName) %>%
  filter(str_detect(feature, "_")) %>%
  mutate(across(-feature, as.numeric)) %>%
  pivot_longer(cols = matches("_C2_")) %>%
  mutate(condition = str_remove(name, "_C2_.$")) %>%
    mutate(condition = str_replace(condition, "BMP2","BMP2+EGF"),
         condition = str_replace(condition, "IFNG","IFNG+EGF"),
         condition = str_replace(condition, "TGFB","TGFB+EGF")) %>%
  group_by(feature, condition) %>%
  summarise(value = mean(value), .groups = "drop") %>%
  ungroup() %>%
  pivot_wider(names_from = condition, values_from = value) %>%
  select(feature, all_of(condition_order))

if_dm <- if_df %>%
  select(-feature) %>%
  as.matrix()
  rownames(if_dm) <- if_df$feature
  
unique_features_data_mean_df <- unique_features_data_ann %>%
  filter(!Direction == "Mixed") %>%
  group_by(Cluster) %>%
      summarise(across(.cols = matches("_24|48"), .fns = mean)) %>%
    ungroup()
unique_features_data_mean_dm <- unique_features_data_mean_df %>%
  select(-Cluster) %>%
  as.matrix()
rownames(unique_features_data_mean_dm) <- unique_features_data_mean_df$Cluster
  
if_module_dm <- rbind(shared_features_data_mean_dm, unique_features_data_mean_dm, if_dm) %>%
  t
  
cor_dm <- cor(if_module_dm)

cor_hm <- Heatmap(cor_dm,
                   name = "correlation ",
                  row_names_gp = gpar(fontsize = 6),
                  column_names_gp = gpar(fontsize = 6),)
cor_hm

if(create_pdfs) {
  pdf("MDD_integrated_module_IF_correlations.pdf")
  res <- print(cor_hm)
  res <- dev.off()
}

#Reduce phenotype set
if_module_dm_reduced <- if_module_dm[,str_detect(colnames(if_module_dm),"Module|Well_Cell_Count|Mean_Cells_per_Cluster|Normalized_Second_Neighbor_Dist|DNA2n_Proportion|Edu_Positive_Proportion|MeanIntensity_KRT5|StdIntensity_KRT5|MinIntensity_KRT5|_Negative|_Positive")]
cor_dm <- cor(if_module_dm_reduced)

cor_hm <- Heatmap(cor_dm,
                   name = "correlation",
                  row_names_gp = gpar(fontsize = 10),
                  column_names_gp = gpar(fontsize = 10),)
cor_hm

if(create_pdfs) {
  pdf("MDD_integrated_module_IF_correlations_reduced.pdf",
      width = 9, height = 8)
  res <- print(cor_hm)
  res <- dev.off()
}


#Modules only set
if_module_dm_reduced <- if_module_dm[,str_detect(colnames(if_module_dm),"Module|_Negative|_Positive")]
cor_dm <- cor(if_module_dm_reduced)

cor_hm <- Heatmap(cor_dm,
                   name = "correlation",
                  row_names_gp = gpar(fontsize = 10),
                  column_names_gp = gpar(fontsize = 10),)
cor_hm

if(create_pdfs) {
  pdf("MDD_integrated_module_module_correlations_reduced.pdf",
      width = 9, height = 8)
  res <- print(cor_hm)
  res <- dev.off()
}

#Modules vs IF  set
if_modules_dm <- if_module_dm[,str_detect(colnames(if_module_dm),"Module|_Negative|_Positive")]

if_phenotypes_dm <- if_module_dm[,str_detect(colnames(if_module_dm), "Well_Cell_Count|Mean_Cells_per_Cluster|Normalized_Second_Neighbor_Dist|DNA2n_Proportion|Edu_Positive_Proportion|MeanIntensity_KRT5|MinIntensity_KRT5")]
cor_dm <- cor(if_modules_dm, if_phenotypes_dm)

cor_hm <- Heatmap(cor_dm,
                   name = "correlation",
                  row_names_gp = gpar(fontsize = 10),
                  column_names_gp = gpar(fontsize = 10),)
cor_hm

if(create_pdfs) {
  pdf("MDD_integrated_module_phenotype_correlations_reduced.pdf",
      width = 9, height = 8)
  res <- print(cor_hm)
  res <- dev.off()
}

```


### cluster analysis


```{r featureTypes, fig.width=8, fig.height=4}
#create a bar chart of the feature types
df <- shared_features_data %>%
  #drop_na() %>%
  select(feature, Type) %>%
  distinct()

p <- ggplot(df, aes(x = Type, fill = Type)) +
  geom_bar() +
  coord_cartesian(ylim = c(0, 10000)) +
  scale_fill_manual(values = assay_cols)+
  labs(title = paste("Type distribution of the",length(unique(paste0(df$feature))), "features at unique timepoints in the shared-features MDD integrated dataset"),
       x ="Feature type") +
  theme_bw()+
  theme(axis.text.x =  element_text(angle = 90),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line =element_blank())

p

p_log <- ggplot(df, aes(x = Type, fill = Type)) +
  geom_bar() +
  scale_y_log10() +
  scale_fill_manual(values = assay_cols)+
  labs(title = paste("Type distribution of the",length(unique(paste0(df$feature))), "shared features at unique timepoints in the MDD integrated dataset"),
       x ="Feature type") +
  theme_bw()+
  theme(axis.text.x =  element_text(angle = 90),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line =element_blank())

p_log
if(create_pdfs){
  pdf("MDD_ligands_feature_type_distribution.pdf")
  print(p)  
  res <- dev.off()
  pdf("MDD_ligands_feature_type_distribution_log_scale.pdf")
  print(p_log)
  res <- dev.off()
} 

```


```{r gapStatistics, fig.width=4, fig.height=4, eval = TRUE}

plot_gap_custom <- function(x) {
  gstab = data.frame(x$Tab, k = seq_len(nrow(x$Tab)))
  p <- ggplot(gstab, aes(k, gap)) + 
    geom_line() +
    geom_errorbar(aes(ymax = gap + SE.sim,
                      ymin = gap - SE.sim), width=0.1) +
    geom_point(size = .5, col=  "red") +
    labs(title = paste("Gap analysis to determine shared-features cluster number,",cluster_method)) +
     theme_bw()+
  theme(axis.ticks = element_blank(),
        panel.border = element_blank(),
        axis.line =element_blank())
  return(p)
}

hclusCut <- function(x, k, d.meth = "euclidean", ...)
   list(cluster = cutree(hclust(dist(x, method=d.meth), ...), k=k))

dm <- shared_features_data %>%
    select(all_of(condition_order)) %>%
  as.matrix()

if(file.exists(paste0("../integrated_analysis/gss_",cluster_method,"_T0_DE.rda"))){
  load(paste0("../integrated_analysis/gss_",cluster_method,"_T0_DE.rda"))
} else {
  gss <- clusGap(dm, FUN = pam, K.max = 15, B = 50, verbose = FALSE)
  save(gss, file = paste0("../integrated_analysis/gss_",cluster_method,"_T0_DE.rda"))
}

  plot_gap_custom(gss)
if(create_pdfs){
  pdf("MDD_ligands_gap_cluster.pdf")
  p <- plot_gap_custom(gss)
  print(p)
  res <- dev.off()
}
# 
# p <- ggplot(integrated_shared, aes(x = as.factor(Cluster), fill = Type)) +
#   geom_bar() +
#   scale_fill_manual(values = assay_cols)+
#   labs(x = "Cluster",
#        fill = "Assay") +
#   theme_bw()+
#   theme(axis.ticks = element_blank(),
#         panel.border = element_blank(),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line =element_blank())
# p
# if(create_pdfs){
#   pdf("MDD_ligands_assay_cluster.pdf")
#   print(p)
#   res <- dev.off()
# }
```

```{r assayClusterdistribution, fig.width=8, fig.height=4, eval = TRUE}

p <- ggplot(shared_features_data_ann, aes(x = as.factor(Cluster), fill = Type)) +
  geom_bar(position = "dodge") +  
  scale_fill_manual(values = assay_cols) +
  labs(x = "Cluster",
       fill = "Assay") +
  facet_wrap(~Type, scales = "free_y") +
  theme_bw()+
  theme(#axis.title = element_blank(),
    #axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line =element_blank())
p

if(create_pdfs){
  pdf("MDD_ligands_cluster_feature_counts.pdf")
  print(p)
  res <- dev.off()
} 

```


This analysis can be interpreted to select the number of clusters in the dataset. It is based on a method described in Modern Statistics for Modern Biology by Susan Holmes and Wolfgang Huber http://web.stanford.edu/class/bios221/book/Chap-Clustering.html and is excerpted below.  

Taking the logarithm of the within-sum-of-squares (log(WSSk)) and comparing it to averages from simulated data with less structure can be a good way of choosing k. This is the basic idea of the gap statistic introduced by Tibshirani, Walther, and Hastie (2001). We compute log(WSSk) for a range of values of k, the number of clusters, and compare it to that obtained on reference data of similar dimensions with various possible ‘non-clustered’ distributions. We can use uniformly distributed data as we did above or data simulated with the same covariance structure as our original data.

The default choice for the number of clusters, k1, is the first value of k for which the gap is not larger than the first local maximum minus a standard error (see the manual page of the clusGap function). ... the choice recommended by Tibshirani, Walther, and Hastie (2001) is the smallest k such that gap(k)≥gap(k+1)−s′k+1...

The first local maximum is k = 3 and the second one is at 9. This analysis is choosing to use k = `r clust_num` which is is the first value where the gap value is less than the local maximum at 9 minus its standard error.  


<bksp>

<bksp>

###

