---
title: "MDD standard figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE,  message=FALSE, warning = FALSE, fig.height=10, fig.width=10)

suppressMessages(library(tidyverse))
suppressMessages(library(ComplexHeatmap))
suppressMessages(library(circlize))
library(umap)
```


## Overview{.tabset .tabset-fade}
These figures are derived from the level 3 data downloaded from the LINCS Molecular Deep Dive files on Synapse https://www.synapse.org/#!Synapse:syn21577710/wiki/601042.  


```{r standardFigureFunctions}

#' Read an assay's csv or tsv data file 
load_data <- function(assay_name, level = 4){
  file_name <- dir(path = paste0("../",assay_name,"/Data"), pattern = paste0("MDD_", assay_name,"_","Level",level,".csv"), recursive = TRUE, full.names = TRUE)
  if(!length(file_name) == 1) stop("There was not one ", assay_name, " data file found")
    
  if(str_detect(file_name, "csv")) {
    df <- read_csv(file_name)
  } else if (str_detect(file_name, "tsv")){
    df <- read_tsv(file_name)
  } else stop("The data file ",file_name, " must be a csv or tsv file type")
} #end of function

get_win_probs <- function(df, low = .02, hi = .98){
  windsorize_probs <- df %>%
    select(-feature) %>%
    unlist %>%
    quantile(probs = c(low, hi), na.rm = TRUE)
}
  
prep_data <- function(df, md){
    df <- df %>%
    gather(specimenID, value, -feature) %>%
    inner_join(md, by = "specimenID") %>%
    select(feature, specimenName, value) %>%
    spread(specimenName, value) %>%
    mutate(feature = factor(feature, levels = unique(df$feature))) %>%
      arrange(feature,)
  #Set the order of the columns
  condition_order <-  paste(rep(c("CTRL","PBS", "HGF", "OSM", "EGF","BMP2", "IFNG", "TGFB"), each = 30, times = 3), rep(c(0, 1, 4 ,8, 24, 48), each = 5, times = 3), rep(c("C1", "C2", "C3"), each = 240), rep(c("A", "B", "C", "D", "E"), times = 105),  sep = "_") %>%
    intersect(colnames(df))
  
  #Arrange columns by standard condition order
  df <- df %>%
    select(feature, condition_order)
  return(df)
}

prep_annotations <- function(df, md){
      df_ann <- df %>%
      gather("specimenName", "value", -feature) %>%
      inner_join(md, by = "specimenName") %>%
      select(ligand, secondLigand, experimentalTimePoint, replicate, collection) %>%
      distinct() %>%
      rename(Time = experimentalTimePoint,
             Ligand = ligand,
             EGF = secondLigand,
             Replicate = replicate,
             Collection = collection) %>%
        mutate(Time = factor(Time,levels = c(0,1,4,8,24,48)),
               Ligand = str_replace(Ligand, "BMP2", "BMP2+EGF"),
               Ligand = str_replace(Ligand, "IFNG", "IFNG+EGF"),
               Ligand = str_replace(Ligand, "TGFB", "TGFB+EGF"),
               Ligand = factor(Ligand, levels = names(ligand_cols))) %>%
      as.data.frame()
      return(df_ann)
}
 #Create the heatmap column annotations
create_top_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md)
  haTop <- HeatmapAnnotation(df = df_ann,
                             col = list(Ligand = ligand_cols,
                                        Time = c("0" = "azure","1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                        EGF = c("EGF" = "#fb8072", "none" = "#8dd3c7"),
                                        Replicate = c("A" = "#a6cee3","B" = "#1f78b4", "C" = "#b2df8a", "D" = "#33a02c", "E" = 
"#fb9a99"),
                                        Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haTop)
}

create_row_annotations <- function(df, md){
  df_ann <- prep_annotations(df, md) %>%
    select(Collection, Replicate, Time, EGF, Ligand)
  ####ha = rowAnnotation(foo = anno_mark(at = c(1:4, 20, 60, 97:100), labels = month.name[1:10])
  ###
  haRow <- rowAnnotation(df = df_ann, show_legend = FALSE,
                         col = list(Ligand = ligand_cols,
                                    Time = c("0" = "azure","1" = "azure1", "4" = "azure2", "8" = "azure3", "24" = "azure4", "48" = "black"), 
                                    EGF = c("EGF" = "#fb8072", "none" = "#8dd3c7"),
                                    Replicate = c("A" = "#a6cee3","B" = "#1f78b4", "C" = "#b2df8a", "D" = "#33a02c", "E" = "#fb9a99"),
                                    Collection = c("C1" = "#D79FD4", "C2" = "#3CC3B9", "C3" = "#B9B34C")))
  return(haRow)
}

  
#####
plot_heatmap <- function(df, md, zscore_rows = FALSE,  assay_name, var_thresh_percentile = 0, show_row_names = FALSE, mark_rows = NULL, cluster_rows = TRUE){
  
  df_vars <- df %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(df_vars, probs = var_thresh_percentile)
df <- df[df_vars >= var_thresh,]

  #convert specimenID column names to specimanName and reorder
  df <- prep_data(df, md) 
title_suffix <- ""
  if(!is.null(mark_rows)) {
    row_nums <- which(df$feature %in% mark_rows)
    ra <- rowAnnotation(foo = anno_mark(at = c(row_nums), labels = df$feature[row_nums]))
  } else {
    ra <- NULL
  }

df_as_matrix <- select(df, -feature) %>%
  as.matrix()
rownames(df_as_matrix) <- df$feature

if(zscore_rows){
  df_as_matrix <- df_as_matrix %>%
    t() %>%
    scale() %>%
    t()
}

windsorize_probs <- quantile(df_as_matrix, probs = c(.02, .98), na.rm = TRUE)
   
#Create the heatmap
  hm <- Heatmap(matrix = df_as_matrix,
          name = "abundance",
          column_title = paste0(assay_name, title_suffix),
          top_annotation = create_top_annotations(df, md),
          right_annotation = ra,
          show_row_names = show_row_names,
          show_column_names = FALSE,
          cluster_columns = FALSE,
          cluster_rows = cluster_rows,
          col = colorRamp2(c(windsorize_probs[1], 0, windsorize_probs[2]), c("blue", "white", "red")),
          na_col = "grey")
  pdf(paste0("MDD_standard_heatmap_",assay_name, ".pdf"))
  draw(hm)
  res <- dev.off()
  draw(hm)
}

plot_umap <- function(df, md, ligand_cols = ligand_cols, assay_name){
  #Generate a 2d scatterplot of the UMAP dimensions
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimanName and reorder
 df <- prep_data(df, md) %>%
    select(-matches("EdU"))
  
   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature
custom_settings = umap.defaults
custom_settings$random_state <- 42
custom_settings$n_epochs = 500
custom_settings$n_neighbors = 25

  df_UMAP <- umap(df_mat,config = custom_settings, na.rm = TRUE)$layout %>%
    data.frame() %>%
    rename(UMAP_1 = X1,
           UMAP_2 = X2) %>%
    cbind(df_ann) %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)),
           Time = as.character(Time),
           Time = as.numeric(Time))
  
  p <- ggplot(df_UMAP, aes(x = UMAP_1,
                           y = UMAP_2,
                           size = Time,
                           colour = Ligand,
                           fill = Ligand)) +
    geom_point(shape = 21, alpha = .9) +
    scale_color_manual(values = ligand_EGF_cols) +
    scale_fill_manual(values = ligand_cols) +
    scale_radius(breaks = c(0,1,4, 8, 24, 48), range = c(2, 6)) +
    labs(title = paste("UMAP embedding of ", assay_name, " data")) +
    guides(colour = guide_legend(override.aes = list(size = 4)),
           size = guide_legend(override.aes = list(shape = 19))) +
    theme_bw() +
    theme(axis.title = element_blank(),
          #axis.ticks = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())

  pdf(paste0("MDD_standard_UMAP_",assay_name, ".pdf"), width = 6, height = 5)
  print(p)
  res <- dev.off()
  print(p)
}

plot_pca <- function(df, md, ligand_cols = ligand_cols, assay_name,
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #Generate a 2d scatterplot of the first two princpal components
  #Use size to represent timepoint, dot in center to represent ligand_2 == EGF

  windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimenName and reorder
  df <- prep_data(df, md) 
  
   #Create annotation values
  df_ann <-prep_annotations(df, md)

   df_mat <- df %>%
    select(-feature) %>%
     t() %>%
  as.matrix()
  colnames(df_mat) <- df$feature

    df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann)
    
  df_pca <- prcomp(df_mat)$x %>%
  data.frame() %>%
    cbind(df_ann) %>%
   mutate(Ligand = factor(Ligand, levels = names(ligand_cols)),
           Time = as.character(Time),
           Time = as.numeric(Time))
  p <- ggplot(df_pca, aes(x = PC1,
                          y = PC2,
                          size = Time,
                          colour = Ligand,
                          fill = Ligand)) +
    geom_point(shape = 21, alpha = .8) +
    scale_color_manual(values = ligand_EGF_cols) +
    scale_fill_manual(values = ligand_cols) +
    scale_radius(breaks = c(0,1,4, 8, 24, 48), range = c(2, 6)) +
    labs(title = paste("UMAP embedding of ", assay_name, " data")) +
    guides(colour = guide_legend(override.aes = list(size = 4)),
           size = guide_legend(override.aes = list(shape = 19))) +
    theme_bw() +
    theme(axis.title = element_blank(),
          #axis.ticks = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
  
p
}


plot_correlation <- function(df, md, assay_name,
                      ligand_cols = ligand_cols,
                   ligand_2_cols = c("EGF" = "#ff0000",
                     "None" = "#00000000")){
  #calculate correlations across the conditions and show in a heatmap

  windsorize_probs <- get_win_probs(df, .01, .99)
  
  #convert specimenID column names to specimanName, reorder delete EGF samples
  df <- prep_data(df, md) 
  
  title_suffix <- ""
 
  #Create annotation values
  haRow <- create_row_annotations(df, md)
  
  #Create the heatmap
  Heatmap(matrix = select(df, -feature) %>%
            as.matrix() %>%
            t %>%
            scale(scale = FALSE) %>%
            t %>%
            cor(use = "complete.obs",method = "spearman"),
          name = "correlation",
          column_title = paste0("Correlation of ",assay_name, title_suffix),
          top_annotation = create_top_annotations(df, md),
          left_annotation = haRow,
          show_row_names = FALSE,
          show_column_names = FALSE,
          cluster_rows = FALSE,
          cluster_columns = FALSE,
          col = colorRamp2(c(-1, 0, 1), c("blue", "white", "red")),
          na_col = "grey")
}

```

```{r globalVariables}

ligand_cols <- c("CTRL" = "#7A4A2A",
                 "PBS" = "#8dd3c7",
                 "HGF" = "#80b1d3",
                 "OSM" = "#fdb462",
                 "EGF" = "#fb8072",
                 "BMP2+EGF" = "#b3de69",
                 "IFNG+EGF" = "#bebada",
                 "TGFB+EGF" = "#ffd92f")

ligand_EGF_cols <- c("CTRL" = "#7A4A2A",
                      "PBS" = "#8dd3c7",
                      "HGF" = "#80b1d3",
                      "OSM" = "#fdb462",
                      "EGF" = "#ff0000",
                      "BMP2+EGF" = "#ff0000",
                      "IFNG+EGF" = "#ff0000",
                      "TGFB+EGF" = "#ff0000")

md <- read_csv("../metadata/MDD_sample_annotations.csv") %>%
  mutate(specimenName = str_replace(specimenName, "ctrl", "CTRL"),
         experimentalCondition = str_replace(experimentalCondition, "ctrl", "CTRL"),
         ligand= str_replace(ligand, "ctrl", "CTRL"))

```

### RPPA{.tabset .tabset-fade}

#### Heatmap

```{r rppa_figures}

RPPA <- load_data("RPPA", level = 3) %>%
  rename(feature = antibody)

```


```{r}
var_thresh_percentile <- 0

plot_heatmap(RPPA,
             zscore_rows = TRUE,
             md,
             assay_name = "RPPA",
             var_thresh_percentile = var_thresh_percentile,
             mark_rows = c("Rb_pS807_S811", "c-Myc"	, "Caspase-7-cleaved","Stat3_pY705", "IRF-1", "EGFR", "HER2_pY1248", "HSP27", "TSC1", "S6_pS235_S236", "PAI-1","P-Met_pY1234_Y1235"))

```


#### Line graphs

```{r, fig.height=15, fig.width=15}

plot_line_graphs <- function(df, md, ligand_cols, fc_thresh = 1, assay_name){
  
  windsorize_probs <- get_win_probs(df, .01, .99)
  
  df_med <- df %>%
    prep_data(md = md) %>%
    gather("specimenName", "value", -feature) %>%
    inner_join(md, by = "specimenName") %>%
    select(ligand, secondLigand, experimentalTimePoint, replicate, collection, feature, value) %>%
    filter(!ligand == "EGF") %>%
    rename(Time = experimentalTimePoint,
           Ligand = ligand,
           EGF = secondLigand,
           Replicate = replicate,
           Collection = collection) %>%
    mutate( Ligand = str_replace(Ligand, "BMP2", "BMP2+EGF"),
            Ligand = str_replace(Ligand, "IFNG", "IFNG+EGF"),
            Ligand = str_replace(Ligand, "TGFB", "TGFB+EGF"),
            Ligand = factor(Ligand, levels = names(ligand_cols))) %>%
    group_by(feature, Collection, Time, Ligand) %>%
    summarise(value = median(value)) %>%
    ungroup() 
  
  fc_set <- df_med %>%
    group_by(Time, Ligand) %>%
    filter(abs(value) >= fc_thresh) %>%
    ungroup() %>%
    select(feature) %>%
    distinct() %>%
    inner_join(df_med, by = "feature") %>%
    mutate(Ligand = factor(Ligand, levels = names(ligand_cols)))
  
  title_suffix <- ""
  p <- ggplot(fc_set, aes(x=Time, y=value, colour=Ligand))+
    geom_line()+
    labs(title=paste0("Line Graphs",title_suffix),
         x = "Time (hours)",
         y="Intensity (AU)") +
    scale_x_continuous(breaks = c(0,8,24,48)) +
    scale_color_manual(values = ligand_cols) +
    theme(axis.text.x = element_text(angle=90),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          strip.text = element_text(size=10, hjust=0),
            panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "gray95"))+
    facet_wrap(~feature, scales = "free_y")
  p
  
}
fc_thresh = 0


plot_line_graphs(df = RPPA, md, ligand_cols = ligand_cols, fc_thresh = fc_thresh, assay_name = "RPPA")

```

The line graphs show the antibodies that have at least one response greater than `r fc_thresh` the EGF response. All values are the median summary of the biological replicates.  

#### UMAP

```{r}

plot_umap(RPPA, md, ligand_cols = ligand_cols, assay_name = "RPPA")

```

#### PCA 

```{r}

plot_pca(RPPA, md, ligand_cols = ligand_cols, assay_name = "RPPA")

```

#### Correlation

```{r}

#plot_correlation(RPPA, md, assay_name = "RPPA")

var_thresh_percentile <- .5

RPPA_vars <- RPPA %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(RPPA_vars, probs = var_thresh_percentile)
RPPA_hi_vars <- RPPA[RPPA_vars > var_thresh,]

plot_correlation(RPPA_hi_vars, md, assay_name = "RPPA")

```


The heatmap shows the pearson correlation values of the top `r 100-var_thresh_percentile*100`th percentile varying features.


###  

### IF{.tabset .tabset-fade}

#### Heatmap  

```{r IF_figures, eval = TRUE}

boundedLog2 <- function(x) log2(x + .0001)

immunofluoresence <- load_data(assay_name = "IF", level = 3) %>%
  rename(feature = "X1") %>%
  mutate(feature = as.character(feature)) %>%
  filter(!feature %in% c("specimenName", "experimentalTimePoint", "collection", "ligand", "replicate")) %>%
  mutate_at(vars(-feature), .funs = as.numeric) %>%
  mutate_if(is.numeric, .funs = boundedLog2)

plot_heatmap(immunofluoresence,
             zscore_rows = TRUE,
             md,
             show_row_names = TRUE,
             assay_name = "IF")

```

The row features in the heatmap are r knitr::combine_words(immunofluoresence$feature)

#### UMAP

```{r, eval = TRUE}
df <- immunofluoresence %>%
  drop_na()
plot_umap(df, md, ligand_cols = ligand_cols, assay_name = "IF")

```

#### PCA 

```{r, eval = TRUE}

plot_pca(df, md, ligand_cols = ligand_cols, assay_name = "IF")

```

#### Correlation

```{r, eval = TRUE}

plot_correlation(df, md, assay_name = "IF")

```

### cycIF{.tabset .tabset-fade}

#### Heatmap 

```{r cycIF_figures, eval = TRUE}

cycIF <-  load_data(assay_name = "cycIF", level = 3) %>%
  filter(str_detect(feature, "_int_")) %>%
  mutate_if(is.numeric, log2) %>%
  prep_data(md = md) %>%
  gather(specimenName, value, -feature) %>%
  mutate(Replicate = str_remove(specimenName, ".*_"),
         feature = as.character(feature)) %>%
  group_by(Replicate, feature) %>%
  mutate(MedCenter = value-median(value, na.rm = TRUE),
         ZScore = MedCenter/sd(value, na.rm = TRUE)) %>%
  ungroup() %>%
    left_join(md) %>%
  dplyr::select(feature, specimenID, ZScore) %>%
  spread(key = specimenID, value = ZScore)

cycIF_vars <- cycIF %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh_value <- 0
var_thresh <- quantile(cycIF_vars, probs = var_thresh_value)
cycIF_hi_vars <- cycIF[cycIF_vars >= var_thresh,]

plot_heatmap(cycIF_hi_vars,
             zscore_rows = FALSE,
             md,
             assay_name = "cycIF",
             mark_rows = c("egfr_2_af488fitc_int_mean_nuc","stat3_4_af488fitc_int_med_cytoplasm","met_6_af488fitc_int_mean_nuc"))

```


Intensity features included in the heatmap have a variance above the `r names(var_thresh) %>% str_remove("%")` percentile. Each feature (row) was z-scaled within each replicate set.  

#### Line graphs

```{r, fig.height=15, fig.width=15}

plot_line_graphs(cycIF, md, ligand_cols = ligand_cols, assay_name = "cycIF")

```


#### UMAP

```{r}

plot_umap(cycIF, md, ligand_cols = ligand_cols, assay_name = "cycIF")

```

#### PCA 

```{r}

plot_pca(cycIF, md, ligand_cols = ligand_cols, assay_name = "cycIF")

```

#### Correlation

```{r}

plot_correlation(cycIF, md, assay_name = "cycIF")

```

### RNAseq{.tabset .tabset-fade}

#### Heatmap  

```{r rnaSeq_figures, eval = TRUE}
RNAseq <- load_data(assay_name = "RNAseq", level = 3) %>%
  rename(feature = ensembl_gene_id)

RNAseq_vars <- RNAseq %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh <- quantile(RNAseq_vars, probs = .9)
RNAseq_hi_vars <- RNAseq[RNAseq_vars > var_thresh,]
  
plot_heatmap(RNAseq_hi_vars,
             zscore_rows = TRUE,
             md,
             assay_name = "RNAseq")

```


Genes included in the heatmap have a variance above the `r names(var_thresh) %>% str_remove("%")` percentile . 

#### UMAP

```{r}

plot_umap(RNAseq, md, ligand_cols = ligand_cols, assay_name = "RNAseq")

```

#### PCA 

```{r}

plot_pca(RNAseq, md, ligand_cols = ligand_cols, assay_name = "RNAseq")

```

#### Correlation

```{r}

plot_correlation(RNAseq_hi_vars, md, assay_name = "RNAseq")

```

### GCP{.tabset .tabset-fade}

#### Heatmap  

```{r GCP_figures}

var_thresh_percentile <- 0

GCP <- load_data(assay_name = "GCP", level = 3) %>%
  rename(feature = histone)

GCP_level2 <- load_data(assay_name = "GCP", level = 2) %>%
  rename(feature = histone) %>%
  gather(key = specimenID, value = value, -feature) %>%
  group_by(feature) %>%
  mutate(value = value - median(value, na.rm = TRUE)) %>%
  ungroup() %>%
    spread(key = specimenID, value = value) %>%
  mutate(feature = factor(feature, levels = GCP$feature)) %>%
  arrange(feature)
  

plot_heatmap(GCP,
             zscore_rows = TRUE,
             md,
             assay_name = "GCP",
             var_thresh_percentile = var_thresh_percentile)

plot_heatmap(GCP,
             zscore_rows = TRUE,
             md,
             assay_name = "GCP",
             var_thresh_percentile = var_thresh_percentile,
             show_row_names = TRUE,
             cluster_rows = FALSE)

```


#### Line graphs

```{r, fig.height=15, fig.width=15}

plot_line_graphs(GCP, md, ligand_cols = ligand_cols, assay_name = "GCP")

```

#### UMAP

```{r}
df <- GCP %>%
  drop_na()
plot_umap(df, md, ligand_cols = ligand_cols, assay_name = "GCP")

```

#### PCA 

```{r}
df <- GCP %>%
  drop_na()
plot_pca(df, md, ligand_cols = ligand_cols, assay_name = "GCP")

```

#### Correlation

```{r}

var_thresh_percentile <- .5

GCP_vars <- GCP %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(GCP_vars, probs = var_thresh_percentile)
GCP_hi_vars <- GCP[GCP_vars > var_thresh,]

plot_correlation(GCP_hi_vars, md, assay_name = "GCP")

GCP_level2_vars <- GCP_level2 %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var, na.rm = TRUE)
  
var_thresh <- quantile(GCP_level2_vars, probs = var_thresh_percentile)
GCP_level2_hi_vars <- GCP_level2[GCP_level2_vars > var_thresh,]

plot_correlation(GCP_level2_hi_vars, md, assay_name = "GCP")

```


The heatmap shows the pearson correlation values of the top `r var_thresh_percentile*100`th percentile varying features.  

### ATACseq{.tabset .tabset-fade}

#### Heatmap  

```{r ATACseq_figures, eval = TRUE}

ATACseq <- load_data(assay_name = "ATACseq", level = 3) %>%
  rename(feature = peak)

ATACseq_vars <- ATACseq %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh <- quantile(ATACseq_vars, probs = .95)
ATACseq_hi_vars <- ATACseq[ATACseq_vars > var_thresh,]
  
plot_heatmap(ATACseq_hi_vars,
             md,
             zscore_rows = TRUE,
             assay_name = "ATACseq")

```


Peaks included in the heatmap have a variance above the `r names(var_thresh) %>% str_remove("%")` percentile . 


#### UMAP

```{r}

plot_umap(ATACseq, md, ligand_cols = ligand_cols, assay_name = "ATACseq")

```

#### PCA 

```{r}

plot_pca(ATACseq, md, ligand_cols = ligand_cols, assay_name = "ATACseq")

```

#### Correlation

```{r}

plot_correlation(ATACseq_hi_vars, md, assay_name = "ATACseq")

```
 

### ATACseq motifs{.tabset .tabset-fade}

#### Heatmap  

```{r ATACseq_motifs_figures, eval = TRUE}
ATACseq_motifs <- load_data(assay_name = "motifs", level = 3) %>%
  rename(feature = family)

ATACseq_vars <- ATACseq %>%
  select(-feature) %>%
  as.matrix() %>%
  apply(1, var)
  
var_thresh <- quantile(ATACseq_vars, probs = .95)
ATACseq_hi_vars <- ATACseq[ATACseq_vars > var_thresh,]
  
plot_heatmap(ATACseq_motifs,
             md,
             show_row_names = TRUE,
             zscore_rows = TRUE,
             assay_name = "motifs")

```



#### UMAP

```{r}

plot_umap(ATACseq_motifs, md, ligand_cols = ligand_cols, assay_name = "motifs")

```

#### PCA 

```{r}

plot_pca(ATACseq_motifs, md, ligand_cols = ligand_cols, assay_name = "motifs")

```

#### Correlation

```{r}

plot_correlation(ATACseq_motifs, md, assay_name = "motifs")

```
 

### L1000{.tabset .tabset-fade}

#### Heatmap  
 
```{r l1000_figures}

L1000_all <- load_data("L1000",
                   level = 3) %>%
  rename(feature = probeset)

C2 <- md %>%
  filter(collection == "C2")

L1000 <- L1000_all %>%
  select(one_of("feature", C2$specimenID))

```



```{r}

plot_heatmap(L1000,
             md,
             zscore_rows = TRUE,
             assay_name = "L1000")

```

#### UMAP

```{r}

plot_umap(L1000, md, ligand_cols = ligand_cols, assay_name = "L1000")

```

#### PCA 

```{r}

plot_pca(L1000, md, ligand_cols = ligand_cols, assay_name = "L1000")

```

#### Correlation

```{r}

plot_correlation(L1000, md, assay_name = "L1000")

```
 
##